### 2012-05-30 doc/ms.tex
Added paper, GPL

\documentclass[useAMS,usenatbib, a4paper]{mn2e} \usepackage{natbib} 
\addtolength{\voffset}{-0.5in}
\usepackage{graphicx} \usepackage{amssymb}
\usepackage{amsmath}

\input{macros.tex}

% -----------------------------------------------------------------------------

\title[Time Delays with Gaussian Processes]
{Inference of Gravitational Lens Time Delays Using Gaussian Processes} 
\author[B. J. Brewer]{Brendon J. Brewer$^1$, Philip J. Marshall$^2$, Gregory Dobler$^3$ \\
$^1$Department of Physics, University of California, Santa Barbara, CA, USA \\
$^2$Department of Physics, University of Oxford, Keble Road, Oxford, OX1 3RH, UK \\
$^3$KITP}

\begin{document}

\date{\today} 
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2011} 
\maketitle \label{firstpage}

% -----------------------------------------------------------------------------

\begin{abstract}

Observations of gravitationally lensed quasars provide a unique method for
measuring cosmological parameters, particularly the Hubble constant $H_0$.
A key part of
this process is the extraction of the time delays from the photometric
monitoring data of the multiple images of the quasar. Quantifying the
uncertainty in this process is challenging, particularly when the time delays
are obscured by the presence of microlensing in the light curves. We present a
fully probabilistic model for the process of inferring the time delays from
the light curve data. In this model, the probability distribution over light
curves is described by a Gaussian Process (GP) whose covariance function
depends on the time delays as well as the properties of the microlensing.
Inferring the parameters of the covariance function from data yields
the time delays, while allowing for a wide range of intrinsic, microlensing, and 
noise-related variability. We test this method extensively on simulated data, 
and find the inferences (including parameter uncertainties) to be robust.
We apply our method to radio and optical monitoring data of the well
studied lenses B1608+656 and RXJ1131$-$123, deriving new time delay estimates complete 
with covariances. The effect on the $H_0$ precision from B1608 is ...

\end{abstract}

\begin{keywords} gravitational lensing --- methods: statistical --- 
\end{keywords}

% -----------------------------------------------------------------------------

\section{Introduction}

Gravitational lensing has long been recognised as a unique tool for measuring
the cosmological parameters \citep{schechter}. The method makes use of the
intrinsic time variability of a quasar that happens to be multiply imaged. As
the quasar's luminosity varies, each of the multiple images presents the same
fluctuations, but with a time delay $\tau$ that depends on the cosmological
parameters. Therefore, measuring the time delays between the images and
comparing them to those predicted from a model of the gravitational lens
potential allows the cosmological parameters to be estimated.

Historical estimates and problems

The promise of this technique has led to dedicated monitoring programs such as
COSMOGrail \citep[][]{2005A&A...436...25E, 2008A&A...488..481V}. Recently,
thorough investigation of systematic effects and robust lens modelling
techniques have allowed $H_0$ to be measured to a precision of X \% using a
single gravitational lens system \citep{2010ApJ...711..201S}. Specific
systematic effects that were robustly modelled include the uncertainty in the
lens potential reconstruction \citep{2009ApJ...691..277S} and the effects of
external convergence from matter along the line of sight
\citep{2006ApJ...642...30F, 2010ApJ...711..201S}. With large numbers of
suitable gravitational lens systems expected to be found in upcoming surveys
\citep{2010MNRAS.405.2579O}, combining the constraints from many systems can
be expected to yield high precision cosmological measurements
\citep{2010ApJ...712.1378P}, provided sources of systematic error are
understood and well modelled.

However, any inference of $H_0$ from the gravitational lensing method inherits
the uncertainty from the measurements of the time delays themselves. A
thorough investigation of all known sources of uncertainty is an integral part
of measuring cosmological parameters, and thus the purpose of this paper is to
develop and present a reliable method for estimating the uncertainty on the
values of the time delays themselves. The method is Bayesian
\citep{2004kats.book.....O} so the results obtained are probability
distributions for the time delays, and the uncertainties in the time delays
are indicated by the widths of these probability distributions. To describe
the underlying time variation of the quasar and the effects of microlensing on
the light curves, we use Gaussian Processes (GPs), a family of probability
distributions over function space that has computationally convenient
properties \citep{rasmussen}.

Our method for extracting time delays from light curve data is similar in
principle to that presented by \citet{2008ApJ...676...80M}, who described the
effect of microlensing by using simulated microlensing light curves. However,
in practical applications their method becomes computationally prohibitive at
a faster rate than the Gaussian Process approach, because the probability of
finding simulated microlensing tracks that match the observed data decreases
approximately exponentially as the length of the data set grows. In contrast,
GP calculations tend to scale as $O(N^3)$, although this can be improved with
the use of approximations such as conjugate gradient techniques
\citep{gibbsmackay}.

Our method is also related to the curve-shifting approach of
\citet{1996A&A...305...97P}, which aims to find those time delays that
minimise the scatter in the shifted light curves. The primary differences
between the GP approach and standard curve shifting are that the GP approach
is fully Bayesian and allows for exploration of the time delay parameter
space, rather than providing a point estimate. Secondly, it is the likelihood
function, rather than the scatter, which guides the inference, making it clear
how uncertainties are to be obtained.

An introduction to Gaussian Processes and their applications can be found in
\citet{rasmussen}. however, in their applications, GPs are usually used as a
prior distribution over the space of functions. Here, GPs describe the
sampling distributions, or the probability of the data given the model
parameters. This is a common situation in astrophysics in fields where the
signals can be modelled by a stochastic process, for example quasar
variability \citep{2009ApJ...698..895K, 2010ApJ...721.1014M, 2011ApJ...735...80Z} and solar-like
oscillations of stars \citep{2009MNRAS.395.2226B}.

% -----------------------------------------------------------------------------

\section{The Model}

We begin by defining our model for the prior knowledge about how the time
delays give rise rise to the observed data. This results in a probability
distribution for the data given some unknown parameters, i.e. sampling
distributions $p(D|\theta)$. Once specific data is in hand, the sampling
distributions become a likelihood function for the parameters which can be
combined with a prior distribution $p(\theta)$ to give inferences. By Bayes'
rule, the posterior distribution for some parameters $\theta$ given data $D$
is given by:
\begin{equation}
p(\theta|D = D^*) \propto p(\theta)p(D|\theta)|_{D = D^*}
\end{equation}
where $D^*$ is the actual data set that was observed. Throughout this paper,
the data $D$ are a set of measured magnitudes of multiple images of a QSO; see
Figure~\ref{simdata} for an example data set.

The intrinsic quasar variability, in magnitudes, and in whatever band the
observations are carried out in, is described by a function $f(t)$. For a
four-image system, the resulting light curves, if observed without gaps and
noise, would be described by the functions
\begin{eqnarray}
y_1(t) &=& m_1 + f(t) + \mu_1(t) \\
y_2(t) &=& m_2 + f(t - \tau_2) + \mu_2(t) \\
y_3(t) &=& m_3 + f(t - \tau_3) + \mu_3(t) \\
y_4(t) &=& m_4 + f(t - \tau_4) + \mu_4(t) 
\end{eqnarray}
where $\{m_1, m_2, m_3, m_4\}$ are the mean magnitudes of the images,
$\{\tau_2, \tau_3, \tau_4\}$ are the time delays relative to image 1, and the
functions $\{\mu_1(t), \mu_2(t), \mu_3(t), \mu_4(t)\}$ describe the perturbing
effects of microlensing on each of the light curves. The data then consist of
noisy, irregularly sampled measurements of the light curves $\{y_1(t), y_2(t),
y_3(t), y_4(t)\}$.

{\bf GAUSSIAN ERRORS ON MAGNITUDES?}

{\bf IMPORTANCE OF MEAN MAGS FOR MILLILENSING, CORRELATION WITH MICROLENSING AND INTRINSIC AGN TIMESCALE, NEED FOR INFORMATIVE PRIORS}

We assume that the covariance function for the quasar variability is given by
\begin{eqnarray}
\textnormal{Cov}\left[f(t_1), f(t_2)\right] = C(|t_2 - t_1|)
\end{eqnarray}
where
\begin{eqnarray}
C(\Delta t) = \sigma_{\rm QSO}^2\exp\left(-\frac{\Delta t}{L_{\rm QSO}}\right)
\end{eqnarray}

{\bf COMMENT ON KELLY}

Similarly, the covariance function for each of the microlensing components is
\begin{eqnarray}
\textnormal{Cov}\left[\mu(t_1), \mu(t_2)\right] = C_{\rm micro}(|t_2 - t_1|)
\end{eqnarray}
where
\begin{eqnarray}
C_{\rm micro}(\Delta t) = \sigma_{\rm micro}^2\exp\left(-\left(\frac{\Delta t}{L_{\rm micro}}\right)^{\alpha_{\rm micro}}\right)
\end{eqnarray}

There is a $\sigma_{\rm micro}$  parameter and an $L_{\rm micro}$ parameter
for each image; this is because the stellar density and lens convergence is
different at each image position. In contrast, we use a single $\alpha_{\rm
micro}$   parameter for all images; this is because it is suspected that the
smoothness of the microlensing variations, which $\alpha$ describes, depends
primarily on the size of the microlensed source, and all images are of the
same source.

The covariance function of the noise-free light curves is then:
\begin{eqnarray}
& & \textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right] \nonumber\\
&=& \left<y_i(t_1)y_j(t_2)\right> \\
&=& \left<\left[f(t_1 - \tau_i) + \mu_i(t_1)\right]\left[f(t_2 - \tau_j) + \mu_j(t_2)\right]\right> \\
&=& \left<f(t_1 - \tau_i)f(t_2 - \tau_j)\right> + \left<\mu_i(t_1)\mu_j(t_2)\right> \\
&=& C_{\rm QSO}(t_1 - \tau_i - t_2 + \tau_j) + \delta_{ij}C_{\rm micro}(t_1 - t_2)
\end{eqnarray}
where the intrinsic variations have been assumed to be uncorrelated with the
microlensing contributions.

The probability distribution for the observed data $\mathbf{y}$ given the
model parameters $\boldsymbol{\theta}$ is a multivariate Gaussian:
\begin{equation}
p(\mathbf{y} | \theta) = \frac{1}{\sqrt{(2\pi)^n \det \mathbf{C}}} \exp\left[-\frac{1}{2}(\mathbf{y} - \mathbf{m})^T\mathbf{C}^{-1}(\mathbf{y} - \mathbf{m})\right]
\end{equation}
where $\mathbf{m}$ is the vector of mean values, and $\mathbf{C}$ is the
covariance matrix of the data, formed by evaluating the covariance function of
the GP at the times of the observations, and then adding diagonal components
to describe the observational errors. That is to say,

\begin{equation}
\mathbf{C} = \textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right] + \mathbf{C}_{\rm noise}.
\end{equation}
{\bf TODO: WRITE THIS PROPERLY}

Note that in this formalism we are using a  GP to describe the {\it
uncertainty} about the microlensing contributions to the light curves so that
we can model the data. This is not quite the same as asserting that the
microlensing magnifications were generated by a Gaussian Process; it is a way
of accoutning for microlensing fluctuations when inferring the time delays. It
is also the case that {\it other} sources of light curve fluctuation, such as
outliers caused by single epoch calibration errors, will be absorbed by the
``microlensing'' process -- the robustness of the time delay inference in that
situation is something we test in Section~\ref{simdata}. It would be more
realistic, yet much more computationally demanding, to use actual simulated
microlensing  light curves \citep{1999JCoAM.109..353W, 2008ApJ...676...80M,
2010NewA...15..181G}. However, for large source sizes or far from caustic
crossing events, we might expect that the GP approximation is adequate. In
Section~\ref{simdata} we test the model on simulated data containing realistic
microlensing effects, and demonstrate that the time delays are still reliably
inferred despite the GP assumption.

A common situation in astronomy is that final data products, such as light
curves, are returned, with error bars, from a pipeline. However, such error
bars are not always trustworthy, as they may have been underestimated due to
modelling assumptions and approximations in the pipeline. In the context of
light curve modelling, underestimated error bars can be detected if the
scatter in the light curve is better modelled by independent Gaussian noise
than it is by a time-correlated microlensing term. Thus, to improve the
robustness of our method, we include a ``noise boost'' parameter $\kappa$
describing the factor by which the reported error bars should be multiplied.
This is another testable mechanism for accounting for outliers.

In summary, the parameter vector is
\begin{equation}
\left\{ \{\tau\}, \{m\}, \sigma_{\rm QSO}, L_{\rm QSO}, \{\sigma_{\rm micro}\}, \{L_{\rm micro}\}, \alpha_{\rm micro}, \kappa \right\}
\end{equation}
where parameters in braces denote a set (one for each QSO image).  For a
double image lens, the total number of parameters in the model is~12. For a
quad, this number is~20. {\bf CHECK.}  A set of uninformative prior
distributions for these parameters are listed in Table~\ref{priors}; we assign
these priors in the analysis described in the first part of this paper. Later,
we will show how a more informative set of priors can be derived, and used to
improve the precision of the time delay inferences {\bf (WE HOPE)}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\begin{center}
\begin{tabular}{ccc}
Parameter & \vline & Prior \\
\hline
$\tau$ & \vline & $U(-\Delta t/2, \Delta t/2)$ \\ 
$m$ & \vline & $U(y_{\rm min}, y_{\rm max})$ \\
$\sigma_{\rm QSO}$ & \vline & $LU(10^{-3}, 10)$ \\
$L_{\rm QSO}$ & \vline & $LU(10^{-3}\Delta t, 10\Delta t)$ \\
$\sigma_{\rm micro}$ & \vline & $LU(10^{-3}, 10)$ \\
$L_{\rm micro}$ & \vline & $LU(10^{-3}\Delta t, 10\Delta t)$ \\
$\alpha_{\rm micro}$ & \vline & $U(1, 2)$ \\
$\kappa$ & \vline & $LU(1, 10)$
\end{tabular}
\end{center}
\caption{Prior probability distributions for the parameters. $U(a, b)$ denotes
a uniform prior between $a$ and $b$, and $LU(a, b)$ denotes a log-uniform
(that is, uniform in the logarithm of the parameter) prior $\propto 1/x$
between $a$ and $b$. $\Delta t$ is the duration of the data set. All priors
are independent.\label{priors}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Computational Details}

We use Diffusive Nested Sampling \citep{dnest} to explore the posterior
distribution for the parameters. Metropolis-Hastings updates are used for all
parameters. For those that do not affect the covariance matrix (the $m$'s),
the Cholesky decomposition is not recomputed, making these steps fast. {\bf
DESCRIBE CONSTRUCTION OF CHOLESKY DECOMP AND MOTIVATION - CITE KOSLOWSKI,
R\&W}

% -----------------------------------------------------------------------------

\section{Tests on Simulated Data}\label{simdata}

In this section, we demonstrate the method on simulated data. We first show
that, in the absence of microlensing, we can recover the input time delays and
GP parameters of the intrinsic AGN lightcurve. We then go on to test the
method on the same mock lens systems, but now with plausible microlensing
fluctuations added. These data violate the modelling assumptions, particularly
that of the probability distribution for the microlensing variations being a
GP.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{A Sample of Mock Lenses}

Motivated by the need to accurately measure time delays for lens cosmography
in the multi-filter optical imaging survey era, we  we selected three
quadruply-imaged quasar lens systems from the mock LSST  catalog provided by
\citet[][hereafter OM10]{OM10}.\footnote{The catalog is available from
\texttt{http://kipac.stanford.edu/collab/research/lensing}}

The OM10 mock lenses are designed to represent massive elliptical galaxies;
they have singular isothermal mass distributions, with velocity dispersions
$\vdisp$ drawn from the SDSS velocity function \citep[extrapolated to higher
redshifts assuming no evolution, as in][]{Ogu++08}. The sources were drawn
from a model quasar luminosity function that matches the observed bright and
faint ends. External shears, drawn from a plausible distribution, were
applied. The mock catalog is the product of performing the lens abundance
calculation for the LSST survey by Monte Carlo integration; OM10 found some
3000 well-measured lens systems to be detectable at LSST single-epoch depth
and image quality. Here we consider a typical slice through this sample: from
the 22 systems lying between $0.4 < \zd < 0.6$, $1.6 < \zs < 2.4$ and $150 <
\vdisp < 300$\kms, three were chosen as being representative of the possible
image configurations while also being particularly prone to microlensing
effects: 6 out of the 12 images lie within the lens galaxy effective radius.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.32\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3042348.eps}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3424637.eps}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=5058554.eps}
\end{minipage}\hfill
\caption{The three simulated lenses, showing image positions (blue) relative
to the lens light effective area (orange). The fraction of surface mass
density in stars ($f^{*}$) at each image position is shown under each image's\label{lenslayouts}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Intrinsic AGN Variability}

In this section we describe our simulated AGN lightcurves in more detail, and
then present the results of the GP inference, all in the absence of any
microlensing. Can we recover the time delays, in all observing strategies? Can
we simultaneously recover the input variability parameters?

\subsubsection{Methodology}

We use the damped random walk, or CAR(1) model of \citet{Kel++09} to simulate
the optical variability of lensed quasars. This is a stochastic model: the
statistics of lightcurves generated by the CAR(1) process are fuly described
by three parameters: the mean magnitude, the decay timescale $\tau_q$ and the
amplitude of  variability $\sigma_q$. While \citeauthor{Kel++09} describe an
efficient processs for generating simulated lightcurves,  we note that the
CAR(1) model is a Gaussian Process with exponential covariance function
\citep{Zu++10}: the CAR(1) parameters are exactly equivalent to those of our
model (Section~\ref{sec:model}).

We choose three representative values of $\tau_q$ and $\sigma_q$ to span the
space of possibilities for lensed quasars (Table~\ref{tab:simpars}).  These
were estimated from the  distributions for the Stripe 82 SDSS spectroscopic
quasar sample of \citet{Kel++09}. We note that the possible timescales range
over two orders of magnitude, making the distinction between observed frame
and rest-frame negligible. When simulating lightcurves for each image, we make
some assumptions to simplify the interpretation of our results. We only
consider measurements in the $i$-band, and we choose the mean magnitude of the
unlensed source quasar to be 23.5 in each case, such that the faintest image
in any given quad lens system lies at or above the 5-sigma $i$-band detection
limit \citep{Ive++10}. Indeed, we assume constant image quality and depth at
every observation epoch, resulting in a constant assumed 5-sigma AB limiting
magnitude per image of 24. 
The different magnifications of the images then lead
naturally to a range of different monitoring signal to noise ratios.
Fluxes corresponding to the lensed image magnitudes
at each epoch were computed, and Gaussian background sky noise added such that a 24th
magnitude point source lay at 5-sigma above zero flux. We ignore readnoise and
shot noise
from the images themselves, an approximation confirmed to be good to a few
percent using the LSST exposure time calculator\footnote{LSST ETC url here.}
with fiducial values of the
sky brightnessss. 
The resulting noisy
fluxes were then converted back to magnitudes. This procedure, while
approximate, captures the fact that it is the image fluxes that have Gaussian
distributed uncertainties: in our inference we assume Gaussian likelihoods for
the magnitudes, an assumption that we need to test. We provide an uncertainty
estimate for each magnitude consistent with these assumptions, based on
propagating the assumed errors on the flux to an uncertainty in the
corresponding magnitude.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\begin{center}
\begin{tabular}{ccc}
\hline
Parameter      & Possible values \\
\hline\hline
$tau_q$/days   & 31.62, {\bf 316.2}, 3162  \\ 
$sigma_q$      & 0.004, {\bf 0.008}, 0.016 \\ 
\hline
\end{tabular}
\end{center}
\caption{Simulated source parameters. The damped random walk variability model
is controlled by two parameters, $tau_q$ and $sigma_q$; the values used here
are representative of those observed in the SDSS stripe 82 spectroscopic 
quasar sample. Our fiducial source lies at the centre of the resulting 3 by 3
grid, and its variability parameters are highlighted in bold font.
\label{tab:simsrc}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We then apply the appropriate magnifications and time delays to generate
4-year lightcurves for each image of our three mock lenses. We sample the
lightcurves in several different ways, to explore the robustness of our time
delay inference. Keeping the monitoring campaign length fixed at 4 years, we
down-sample to a plausible cadence (separation between successive
observations) and season length. Again, we choose representative values:
COSMOGRAIL \citep{cosmograil} typically monitor lenses for 8 months per year
at a cadence of 4 days, LSST will likely have seasons closer to 4 months, with
a single-filter cadence closer to 8 days (although this is somewhat
pessimistic). 4-month seasons with 4 day cadence mimics LSST monitoring should
multi-filter joint analysis be possible, while 8-month seasons with 8-day
cadence represents a cheaper version of a dedicated monitoring campaign. These
choices resulted in 4 possible observing strategies, with total numbers of
measurements of 60, 120, 120 and 240. A more thorough investigation of time
delay measurement feasibility with LSST is left to future work. Four observing
strategies, times 3 times 3 possible variable sources, lensed by 3 deflectors 
makes a total of 108 simulated datasets, and 324 independent time delays to be
inferred.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\begin{tabular}{ccccc}
\hline
Observing strategy & Cadence (days) & Season length (days) 
& Epochs/season & Total epochs\\
\hline\hline
A  & 8  & 120 & 15  &  60 \\ 
\textbf{B}  & \textbf{4}  & \textbf{120} & \textbf{30}  & \textbf{120} \\
C  & 8  & 240 & 30  & 120 \\
D  & 4  & 240 & 60  & 240 \\
\hline
\end{tabular}
\end{center}
\caption{Simulating observing strategies explored. 4 day cadence and 8 month
seasons are an approximation to those used in the COSMOGRAIL survey, while
LSST will provide closer to 8 day cadence and 4 month seasons. Strategy B is
our fiducial one, corresponding approximately to (somehow) using multiple
LSST filters to achieve a 4-day cadence over four consecutive 4-month seasons.
\label{observing}}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\TODO{PJM}{Check these  numbers, esp COSMOGRAIL...}

Two example simulated datasets are shown in Figure~\ref{fig:simdata}. In  the
next sub-section we present results from our Gaussian Process modelling of all 108
datasets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.48\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3042348_tau=31_sigma=4_cadence=4_season=8_lightcurve.eps}
\end{minipage}
\begin{minipage}{0.48\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=5058554_tau=3162_sigma=16_cadence=8_season=4_lightcurve.eps}
\end{minipage}\hfill
\caption{Two example simulated lensed AGN datasets (with no microlensing).
Left: a minimally-variable source ($\tau_q = 31.62$ days, $\sigma_q = 0.004$),
lensed into a cross configuration (lens ID 3042348), and monitored with high
cadence and long season length (4 days, 8 months, strategy D). Right: a
maaximally-variable source ($\tau_q = 3162$ days, $\sigma_q = 0.016$), lensed
into a major axis cusp configuration (lens ID 5058554), and monitored with low
cadence and short season length (8 days, 4 months, strategy A).
\label{fig:simdata}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Results}

Strategy B, dt recovery across sample. 3x3x3 = 27 inferences, 27x18 dt estimates. Are
uncertainties on dt accurate? Trends with dt.

Effect of changing observing strategy.

Mean mag and var par recovery across sample. Trends with tau and sigma.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Including Microlensing}

In this section we describe our simulations of the microlensing occurring at
each image position, and then repeat the inferences of the previous section in
the presence of this contaminant. How is the time delay inference degraded?
Can the microlensing curves be modelled by the GP? How are the results
affected by observing strategy?

\subsubsection{Methodology}

\TODO{GD}{Write this section please!}
\TODO{PJM/BJB}{Suitable illustrative figure?}

Stellar mass distributions were ``painted on'' to the model lens galaxies, as
described in Dobler et al.\, (in preparation). Established galaxy scaling
relations were used to draw luminosities and effective radii given the model
velocity dispersions; a surface brightness distribution with the same
ellipticity as that of the lens mass, and with a de Vaucoleurs cite profile 
were assumed. Finally, a mass to light ratio was assumed, and the fraction of
surface mass density in the form of stars ($\fstar$) 
computed, at each image position. This parameter, the local lens
convergence, and its shear, were then used to generate a starfield at each
image position, and compute the corresponding source plane magnification map,
using the code of ...

Wambsganss maps. Then, parameters chosen to span the possibilities. Constant
speed traversals, 2 speeds (fast and slow).  2 source sizes, from Kochanek etc
work, separated by factor of 3 (large and small).  fstar values chosen in
catalog and in lens selection, check they are high (c. 0.5).  Check timescales
compared to length of dataset. Illustrative plot.

Apply to 3 lenses with "typical" intrinsic variabilities, ie the middle tau
and sigma of the 3x3 grid in the previous section.

\subsubsection{Results}

dt recovery across sample. 3x(2x2) = 12 inferences, 12x18=216 dt estimates.
Are uncertainties on dt accurate? Trends with dt.

Mean mag and ml par recovery across sample. 48 microlensing lightcurves (4
images times 3 lenses x 4 setups). Trends with fstar, source size and v.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Discussion}

What did we learn from simulated data?
When do we succeed and when do we fail?
WHat future work is required?

% -----------------------------------------------------------------------------

\section{Application to B1608$+$656}

B1608$+$656 is a gravitational lens system a radio-loud source galaxy; it has 
very precisely measured time delays, and has been successfully used as a
cosmographic probe \citep[e.g.][]{Suy++10}.  Three seasons of VLA monitoring
were taken by  \citet{2002ApJ...581..823F, 1999ApJ...527..498F},  and are
shown in Figure~\ref{b1608_data}. The dominant effect visible in the data is
the gradual change in the flux of the background AGN over time, visible in all
four images. This gradual change is not so useful for measuring the time
delays. Instead, the time delay information comes mostly from small short-term
fluctuations on top of the global trend. The size of the radio source is
assumed to be much larger than the Einstein radius of the stars in the lens
galaxy, leading to negligible microlensing. In this section we model the
B1608$+$656 radio lightcurve data with our combined intrinsic plus
microlensing Gaussian process, to test this assumption and re-measure the
system time delays.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\includegraphics[scale=0.75]{figs1/b1608_data.eps}
\caption{Monitoring data for B1608+656.\label{b1608_data}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\includegraphics{figs1/b1608_delays.eps}
\caption{Inferred time delays for B1608+656.\label{b1608_delays}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[scale=0.42]{figs1/correlation.eps}
\caption{Inferred intrinsic QSO variability parameters for B1608+656. This is the only significant correlation in the joint posterior distribution. The data are consistent with a small or large value for the variability amplitude, but if the amplitude is large, then the timescale must also be large so that we have only observed a small portion of the timescale.\label{correlation}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% -----------------------------------------------------------------------------

\section{Application to J1131}

\citep{2006astro.ph..5321M}

% -----------------------------------------------------------------------------

\section{Discussion and Further Work}

Robustness of time delays.

Effect of outliers. Do we deal with them well enough in th ecurrent implementation?

Mean mags. implications for millilensing.

Microlensing pars, good enough? Future: inference of physical microlensing parameters. 
Use of informative priors.

Multi-wavelength extension - colour variability.

GP approximations for speed up in LSST era.

% -----------------------------------------------------------------------------

\section{Conclusions}

\begin{itemize}
\item Microlensing can be approximated as GP
\item Time delays and covariance between them can be robustly estimated
\item B1608
\item J1131
\end{itemize}

% -----------------------------------------------------------------------------

\section{Acknowledgements}
David Hogg, Jo Bovy, Malte Tewes, Matt Auger, Sherry Suyu, Chris Fassnacht, Tommaso Treu, Alicia Berciano Alba
(some may be upgraded to authors)

% -----------------------------------------------------------------------------

\begin{thebibliography}{99} 
\bibitem[\protect\citeauthoryear{Brewer 
\& Stello}{2009}]{2009MNRAS.395.2226B} Brewer B.~J., Stello D., 2009, MNRAS, 395, 2226

\bibitem[{Brewer {et~al.}(2010)Brewer, P\'{a}rtay, \& Cs\'{a}nyi}]{dnest}
Brewer, B., P\'{a}rtay, L., \& Cs\'{a}nyi, G. 2010, Statistics and Computing, doi: 10.1007/s11222-010-9198-8, arXiv: 0912.2380

\bibitem[\protect\citeauthoryear{Eigenbrod et 
al.}{2005}]{2005A&A...436...25E} Eigenbrod A., Courbin F., Vuissoz C., Meylan G., Saha P., Dye S., 2005, A\&A, 436, 25

\bibitem[\protect\citeauthoryear{Fassnacht et 
al.}{2006}]{2006ApJ...642...30F} Fassnacht C.~D., Gal R.~R., Lubin L.~M., 
McKean J.~P., Squires G.~K., Readhead A.~C.~S., 2006, ApJ, 642, 30

\bibitem[\protect\citeauthoryear{Fassnacht et 
al.}{2002}]{2002ApJ...581..823F} Fassnacht C.~D., Xanthopoulos E., Koopmans 
L.~V.~E., Rusin D., 2002, ApJ, 581, 823

\bibitem[\protect\citeauthoryear{Fassnacht et 
al.}{1999}]{1999ApJ...527..498F} Fassnacht C.~D., Pearson T.~J., Readhead 
A.~C.~S., Browne I.~W.~A., Koopmans L.~V.~E., Myers S.~T., Wilkinson P.~N., 
1999, ApJ, 527, 498

\bibitem[\protect\citeauthoryear{Garsden 
\& Lewis}{2010}]{2010NewA...15..181G} Garsden H., Lewis G.~F., 2010, NewA, 15, 181

\bibitem[\protect\citeauthoryear{Gibbs \& MacKay}{1997}]{gibbsmackay}
Gibbs, M., Mackay, D., 1997, Efficient implementation of Gaussian processes, Cavendish Lab., Cambridge, U.K., Tech. Rep.

\bibitem[\protect\citeauthoryear{Kelly, Bechtold, 
\& Siemiginowska}{2009}]{2009ApJ...698..895K} Kelly B.~C., Bechtold J., Siemiginowska A., 2009, ApJ, 698, 895

\bibitem[\protect\citeauthoryear{MacLeod et 
al.}{2010}]{2010ApJ...721.1014M} MacLeod C.~L., et al., 2010, ApJ, 721, 
1014

\bibitem[\protect\citeauthoryear{Morgan et al.}{2008}]{2008ApJ...676...80M} 
Morgan C.~W., Eyler M.~E., Kochanek C.~S., Morgan N.~D., Falco E.~E., 
Vuissoz C., Courbin F., Meylan G., 2008, ApJ, 676, 80

\bibitem[\protect\citeauthoryear{Morgan et al.}{2006}]{2006astro.ph..5321M} 
Morgan N.~D., Kochanek C.~S., Falco E.~E., Dai X., 2006, astro, 
arXiv:astro-ph/0605321

\bibitem[\protect\citeauthoryear{Oguri 
\& Marshall}{2010}]{2010MNRAS.405.2579O} Oguri M., Marshall P.~J., 2010, MNRAS, 405, 2579

\bibitem[\protect\citeauthoryear{O'Hagan 
\& Forster}{2004}]{2004kats.book.....O} O'Hagan A., Forster J., 2004, Kendall's Advanced Theory of Statistics, Volume 2B: Bayesian Inference, London: Hodder Arnold, 2004

\bibitem[\protect\citeauthoryear{Paraficz 
\& Hjorth}{2010}]{2010ApJ...712.1378P} Paraficz D., Hjorth J., 2010, ApJ, 712, 1378

\bibitem[\protect\citeauthoryear{Pelt et 
al.}{1996}]{1996A&A...305...97P} Pelt J., Kayser R., Refsdal S., Schramm T., 1996, A\&A, 305, 97

\bibitem[\protect\citeauthoryear{Rasmussen \& Williams}{2006}]{rasmussen} Rasmussen C., Williams C., Gaussian Processes for Machine Learning, The MIT Press, 2006

\bibitem[\protect\citeauthoryear{Schechter}{2005}]{schechter} 
Schechter P.~L., 2005, IAUS, 225, 281

\bibitem[\protect\citeauthoryear{Suyu et al.}{2010}]{2010ApJ...711..201S} 
Suyu S.~H., Marshall P.~J., Auger M.~W., Hilbert S., Blandford R.~D., 
Koopmans L.~V.~E., Fassnacht C.~D., Treu T., 2010, ApJ, 711, 201

\bibitem[\protect\citeauthoryear{Suyu et al.}{2009}]{2009ApJ...691..277S} 
Suyu S.~H., Marshall P.~J., Blandford R.~D., Fassnacht C.~D., Koopmans 
L.~V.~E., McKean J.~P., Treu T., 2009, ApJ, 691, 277

\bibitem[\protect\citeauthoryear{Vuissoz et 
al.}{2008}]{2008A&A...488..481V} Vuissoz C., et al., 2008, A\&A, 488, 481

\bibitem[\protect\citeauthoryear{Wambsganss}{1999}]{1999JCoAM.109..353W} 
Wambsganss J., 1999, JCoAM, 109, 353

\bibitem[\protect\citeauthoryear{Zu, Kochanek, 
\& Peterson}{2011}]{2011ApJ...735...80Z} Zu Y., Kochanek C.~S., Peterson B.~M., 2011, ApJ, 735, 80 
\end{thebibliography}

\end{document}

### 2012-05-30 src/Python/PJM/make_simulated_data.py
Phil's python

@@ -1,0 +1,370 @@
+#!/usr/bin/env python
+# ======================================================================
+
+# Globally useful modules, imported here and then accessible by all
+# functions in this file:
+
+import matplotlib
+# Force matplotlib to not use any Xwindows backend:
+matplotlib.use('Agg')
+
+# Fonts, latex:
+matplotlib.rc('font',**{'family':'serif', 'serif':['TimesNewRoman']})
+matplotlib.rc('text', usetex=True)
+
+import numpy,atpy,pylab,sys,getopt,pyfits
+
+# ======================================================================
+
+def make_simulated_data(argv):
+
+  USAGE = """
+  NAME
+    make_simulated_data.py
+
+  PURPOSE
+    Given a FITS table of 10 year AGN lightcurves (from Greg) and some lens 
+    IDs, make simple simulated LSST data. 
+
+  COMMENTS
+    Use atpy to keep control of tables. Output is in BJB-compliant format.
+    Sampling patterns are generated internally, later on may want to import 
+    one from LSST opsim output.
+
+  USAGE
+    make_simulated_data.py [flags] [options]
+
+  FLAGS
+          -h --help    Print this message [0]
+          -v --verbose Be verbose [0]
+
+  INPUTS
+          lctable      FITS table of lightcurves, in GD format
+          ID1 ID2 etc  List of lens IDs to use.
+          
+  OPTIONAL INPUTS
+          --catalog    Lens catalog [def=oguri_qso_LSST_fstar.fits]
+          -m ms        Make all sources have the same magnitude
+          
+  OUTPUTS
+          stdout       Useful information
+
+  EXAMPLES
+
+  BUGS
+
+  HISTORY
+    2011-09-09 started Marshall (UCSB)    
+  """
+
+  # --------------------------------------------------------------------
+
+  try:
+      opts, args = getopt.getopt(argv[1:], "hvm:c:", ["help","verbose","source-magnitude","catalog="])
+  except getopt.GetoptError, err:
+      # print help information and exit:
+      print str(err) # will print something like "option -a not recognized"
+      print USAGE
+      sys.exit(2)
+
+  catalog = 'oguri_qso_LSST_fstar.fits'
+  ms = 0
+  verbose = False
+  for o,a in opts:
+      if o == "-v":
+          vb = True
+      elif o in ("-h", "--help"):
+          print USAGE
+          sys.exit()
[…]

### 2012-05-30 src/Python/PJM/plot_oguri_lens.py

@@ -1,0 +1,400 @@
+#!/usr/bin/env python
+# ======================================================================
+
+# Globally useful modules, imported here and then accessible by all
+# functions in this file:
+
+import matplotlib
+# Force matplotlib to not use any Xwindows backend:
+matplotlib.use('Agg')
+
+# Fonts, latex:
+matplotlib.rc('font',**{'family':'serif', 'serif':['TimesNewRoman']})
+matplotlib.rc('text', usetex=True)
+
+import numpy,pylab,sys,getopt,pyfits
+
+# ======================================================================
+
+def plot_oguri_lens(argv):
+
+  USAGE = """
+  NAME
+    plot_oguri_lens.py
+
+  PURPOSE
+    Given a lensID, pull out the information on that lens from the 
+    Oguri catalog and compute some basic quantities, 
+    then plot them on the sky.
+
+  COMMENTS
+    FITS table from Dobler has lens galaxy sizes in it. Best to use pyfits 
+    utilities.
+
+  USAGE
+    plot_oguri_lens.py [flags] [options] lensID
+
+  FLAGS
+          -h --help    Print this message [0]
+          -v --verbose Be verbose [0]
+          --eps        Postscript output [0, png is default]
+          --showdisk   Plot a seeing disk for comparison [0]
+
+  INPUTS
+          lensID       Lens ID number
+
+  OPTIONAL INPUTS
+          --catalog    Lens catalog [def=oguri_qso_LSST_fstar.fits]
+          
+  OUTPUTS
+          stdout       Useful information
+
+  EXAMPLES
+
+    First double in LSST catalog:
+          plot_oguri_lens.py 18060 
+
+    First quad in LSST catalog:
+          plot_oguri_lens.py 4376
+
+  BUGS
+    Image opening angles greater than 180deg are not possible - need to 
+    replace the largest with 360-sum(phi1:phi3)
+
+  HISTORY
+    2010-06-13 started Marshall (KIPAC)
+    2011-03-05 adaptd for FITS catalogs Marshall (UCSB)    
+  """
+
+  # --------------------------------------------------------------------
+
+  try:
[…]

### 2012-05-30 src/Python/PJM/plot_one_lens_lightcurves.py

@@ -1,0 +1,254 @@
+#!/usr/bin/env python
+# ======================================================================
+
+# Globally useful modules, imported here and then accessible by all
+# functions in this file:
+
+import matplotlib
+# Force matplotlib to not use any Xwindows backend:
+matplotlib.use('Agg')
+
+# Fonts, latex:
+matplotlib.rc('font',**{'family':'serif', 'serif':['TimesNewRoman']})
+matplotlib.rc('text', usetex=True)
+
+import numpy,pylab,sys,getopt,string
+
+# ======================================================================
+
+def plot_one_lens_lightcurves(argv):
+
+  USAGE = """
+  NAME
+    plot_one_lens_lightcurves.py
+
+  PURPOSE
+    Given a BJB lighcurve file, plot all 4 image lightcurves with sensible
+    offsets.
+
+  COMMENTS
+    Optionally include predicted curves? True curves? Can plot many files...
+
+  USAGE
+    plot_one_lens_lightcurves.py [flags] [options] lcfile
+
+  FLAGS
+          -h --help    Print this message [0]
+          -v --verbose Be verbose [0]
+          --eps        Postscript output [0, png is default]
+
+  INPUTS
+          lcfile       BJB format lightcurve file (or list of files)
+
+  OPTIONAL INPUTS
+          -t --truth   Overlay true lightcurve in grey (on each plot)
+          
+  OUTPUTS
+          stdout       Useful information
+
+  EXAMPLES
+
+
+  BUGS
+
+  HISTORY
+    2011-09-12 started Marshall (KIPAC)
+  """
+
+  # --------------------------------------------------------------------
+
+  try:
+      opts, args = getopt.getopt(argv[1:], "hvt:", ["help","eps","truth="])
+  except getopt.GetoptError, err:
+      # print help information and exit:
+      print str(err) # will print something like "option -a not recognized"
+      print USAGE
+      sys.exit(2)
+
+  vb = False
+  eps = False
+  truefile = 0
+  lcfiles = []
+  for o,a in opts:
+      if o == "-v":
+          vb = True
+      elif o in ("-h", "--help"):
+          print USAGE
+          sys.exit()
+      elif o in ("--eps"):
+          eps = True
+      elif o in ("-t","--truth"):
+          truefile = a
+          lcfiles.append(truefile)
+      else:
[…]

### 2012-05-30 src/Python/PJM/select_quads.py

@@ -1,0 +1,134 @@
+#!/usr/bin/env python
+# ======================================================================
+
+# Globally useful modules, imported here and then accessible by all
+# functions in this file:
+
+import matplotlib
+# Force matplotlib to not use any Xwindows backend:
+matplotlib.use('Agg')
+
+# Fonts, latex:
+matplotlib.rc('font',**{'family':'serif', 'serif':['TimesNewRoman']})
+matplotlib.rc('text', usetex=True)
+
+import numpy,pylab,sys,getopt,pyfits
+
+# ======================================================================
+
+def select_quads(argv):
+
+  USAGE = """
+  NAME
+    select_quads.py
+
+  PURPOSE
+    Select 3 quads from LSST catalog, and plot.
+
+  COMMENTS
+    FITS table from Dobler has lens galaxy sizes in it. Best to use pyfits 
+    utilities.
+
+  USAGE
+    select_quads.py [flags] [options]
+
+  FLAGS
+          -h --help    Print this message [0]
+          -v --verbose Be verbose [0]
+
+  INPUTS
+
+  OPTIONAL INPUTS
+          --catalog    Lens catalog [def=oguri_qso_LSST_fstar.fits]
+          
+  OUTPUTS
+          stdout       Useful information
+
+  EXAMPLES
+
+  BUGS
+
+  HISTORY
+    2011-03-05 started Marshall (UCSB)    
+  """
+
+  # --------------------------------------------------------------------
+
+  try:
+      opts, args = getopt.getopt(argv[1:], "hvl:c:", ["help","catalog=","logfile="])
+  except getopt.GetoptError, err:
+      # print help information and exit:
+      print str(err) # will print something like "option -a not recognized"
+      print USAGE
+      sys.exit(2)
+
+  catalog = 'oguri_qso_LSST_fstar.fits'
+  verbose = False
+  for o,a in opts:
+      if o == "-v":
+          vb = True
+      elif o in ("-h", "--help"):
+          print USAGE
+          sys.exit()
+      elif o in ("-c", "--catalog"):
+          catalog = a
+      else:
+          assert False, "unhandled option"
+
+  # --------------------------------------------------------------------
+
+  # Read in lens and image data in expected format:
+
+  hdulist = pyfits.open(catalog)
[…]

### 2012-05-30 src/DNest/Data.cpp
BJB's DNest code for the inference

@@ -1,0 +1,70 @@
+#include "Data.h"
+#include <fstream>
+#include <iostream>
+#include <cstdlib>
+#include <algorithm>
+
+using namespace std;
+
+Data::Data()
+:loaded(false)
+{
+
+}
+
+
+void Data::load(const char* filename)
+{
+	if(loaded)
+		cerr<<"# Already loaded file."<<endl;
+
+	fstream fin(filename, ios::in);
+	double temp1, temp2, temp3, temp4;
+
+	if(!fin)
+	{	
+		cerr<<"Error opening "<<filename<<"."<<endl;
+		exit(0);
+	}
+
+	// Skip past comments
+	while(fin.peek() == '#')
+		fin.ignore(1024*1024, '\n');
+
+	while(fin>>temp1 && fin>>temp2 && fin>>temp3 && fin>>temp4)
+	{
+		t.push_back(temp1);
+		y.push_back(temp2);
+		sig.push_back(temp3);
+		qsoID.push_back((int)temp4);
+	}
+	cout<<"# Loaded "<<t.size()<<" points from file "<<filename<<"."<<endl;
+
+	fin.close();
+
+	N = t.size();
+	numQSOs = 1 + *max_element(qsoID.begin(), qsoID.end());
+	tMin = *min_element(t.begin(), t.end());
+	tRange = *max_element(t.begin(), t.end()) - tMin;
+	yMin = *min_element(y.begin(), y.end());
+	yRange = *max_element(y.begin(), y.end()) - yMin;
+
+	loaded = true;
+}
+
+double Data::getTMin() const
+{
+	return tMin;
+}
+
+double Data::getTRange() const
+{
+	return tRange;
+}
+
+int Data::getN() const
+{
+	return N;
+}
+
+

### 2012-05-30 src/DNest/Data.h

@@ -1,0 +1,31 @@
+#ifndef _Data_
+#define _Data_
+
+#include <vector>
+
+class Data
+{
+	friend class TDModel;
+
+	private:
+		std::vector<double> t, y, sig;
+		std::vector<int> qsoID;
+
+		bool loaded;
+
+		int N;
+		int numQSOs;
+		double tMin, tRange;
+		double yMin, yRange;
+
+	public:
+		Data();
+		void load(const char* filename);
+		double getTMin() const;
+		double getTRange() const;
+		int getN() const;
+
+};
+
+#endif
+

### 2012-05-30 src/DNest/TDModel.cpp

@@ -1,0 +1,361 @@
+#include "TDModel.h"
+#include "Utils.h"
+#include "RandomNumberGenerator.h"
+#include <cmath>
+#include <iostream>
+#include <Eigen/Dense>
+
+using namespace Eigen;
+using namespace std;
+using namespace DNest;
+
+Data TDModel::data;
+
+const double TDModel::minSig = 1E-3;
+const double TDModel::maxSig = 10.;
+
+TDModel::TDModel()
+:timeDelays(data.numQSOs), sigMicrolensing(data.numQSOs), tauMicrolensing(data.numQSOs), meanMagnitudes(data.numQSOs)
+,covMat(data.N, data.N)
+{
+	if(!data.loaded)
+	{
+		cerr<<"Data has not been loaded! Cannot construct TDModel"<<endl;
+		exit(0);
+	}
+}
+
+TDModel::~TDModel()
+{
+
+}
+
+DNest::Model* TDModel::factory() const
+{
+	return new TDModel;
+}
+
+DNest::Model* TDModel::clone() const
+{
+	return new TDModel(*this);
+}
+
+void TDModel::copyFrom(const DNest::Model* other)
+{
+	*this = *((TDModel*)other);
+}
+
+void TDModel::fromPrior()
+{
+	for(size_t i=0; i<timeDelays.size(); i++)
+		timeDelays[i] = (i==0)?(0):(data.tRange*(-0.5 + randomU()));
+	for(size_t i=0; i<sigMicrolensing.size(); i++)
+		sigMicrolensing[i] = exp(log(minSig) + log(maxSig/minSig)*randomU());
+	for(size_t i=0; i<tauMicrolensing.size(); i++)
+		tauMicrolensing[i] = exp(log(1E-2*data.tRange) + log(1E4)*randomU());
+
+	for(size_t i=0; i<meanMagnitudes.size(); i++)
+		meanMagnitudes[i] = data.yMin + data.yRange*randomU();
+
+	sigIntrinsic = exp(log(minSig) + log(maxSig/minSig)*randomU());
+	tauIntrinsic = exp(log(1E-2*data.tRange) + log(1E4)*randomU());
+	alphaIntrinsic = 1.0 + randomU();
+	alphaMicrolensing = 1.0 + randomU();
+	sigmaBoost = exp(log(10.0)*randomU());
+
+	DNest::Model::fromPrior();
+	formCovarianceMatrix();
[…]

### 2012-05-30 src/DNest/TDModel.h

@@ -1,0 +1,62 @@
+#ifndef _TDModel_
+#define _TDModel_
+
+#include "Model.h"
+#include "Data.h"
+#include <ostream>
+#include <istream>
+#include <vector>
+#include <Eigen/Dense>
+
+class TDModel:public DNest::Model
+{
+	private:
+		std::vector<double> timeDelays;
+		std::vector<double> sigMicrolensing;
+		std::vector<double> tauMicrolensing;
+		std::vector<double> meanMagnitudes;
+
+		double sigIntrinsic, tauIntrinsic, alphaIntrinsic, alphaMicrolensing, sigmaBoost;
+		Eigen::MatrixXd covMat;
+		Eigen::LLT< Eigen::MatrixXd > cholesky;
+
+		static const double minSig;
+		static const double maxSig;
+
+		static Data data;
+
+		void formCovarianceMatrix();
+
+		double perturbHelper1();
+		double perturbHelper2();
+		double perturbHelper3();
+		double perturbHelper4();
+		double perturbHelper5();
+		double perturbHelper6();
+		double perturbHelper7();
+		double perturbHelper8();
+		double perturbHelper9();
+
+		// Noise-free covariance function
+		double covariance(double t1, double t2, int qsoID1, int qsoID2);
+
+	public:
+		TDModel();
+		~TDModel();
+		DNest::Model* factory() const;
+		DNest::Model* clone() const;
+		void copyFrom(const DNest::Model* other);
+		void fromPrior();
+		void calculateLogLikelihood();
+		double perturb();
+		double getValue();
+		double getLogLikelihood() const;
+		void print(std::ostream& out) const;
+		void read(std::istream& in);
+		std::vector<double> evaluate(const std::vector<double>& t, const std::vector<int>& qsoID);
+		static void loadData(const char* filename);
+		static Data& getData();
+};
+
+#endif
+

### 2012-05-30 src/DNest/main.cpp

@@ -1,0 +1,120 @@
+/*
+    Copyright (C) 2011 Brendon J. Brewer
+    This file is part of DNest, the Diffusive Nested Sampler.
+
+    DNest is free software: you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation, either version 3 of the License, or
+    (at your option) any later version.
+
+    DNest is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with DNest.  If not, see <http://www.gnu.org/licenses/>.
+*/
+
+
+#include <iostream>
+#include "TDModel.h"
+#include "MultiSampler.h"
+#include "RandomNumberGenerator.h"
+#include "DNestSampler.h"
+#include <boost/thread.hpp>
+#include <boost/program_options.hpp>
+
+using namespace std;
+using namespace DNest;
+namespace po = boost::program_options;
+
+void runMultiThread(const Model* exampleModel, const string& levelsFile);				// Run multithreaded sampler
+void runLegacy(const Model* exampleModel, const string& levelsFile);	// Run legacy sampler
+
+int main(int argc, char** argv)
+{
+	// What kind of model will we be running?
+	TDModel::loadData("Data/j1131.txt");
+	TDModel exampleModel;
+
+
+	// Which kind of sampler, and should we pre-load level structure?
+	bool legacy = false;
+	string levelsFile = "";
+
+	// Handle command line options
+	{
+		// Declare the supported options.
+		po::options_description desc("Welcome to DNest. Allowed options");
+		desc.add_options()
+		    ("help", "produce help message")
+		    ("levels", po::value<string>(), "Input levels file")
+		    ("legacy", "use legacy sampler (serial)")
+		;
+
+		po::variables_map vm;
+		po::store(po::parse_command_line(argc, argv, desc), vm);
+		po::notify(vm);    
+
+		if (vm.count("help"))
+		{
+		    cout<<desc<<endl;
+		    return 1;
+		}
+
+		if(vm.count("legacy")) 
+			legacy = true;
+
[…]

### 2012-05-30 doc/ms.tex
Removed a bunch of old stuff (it is still accessible from the history, don't worry!)

@@
 
-\input{macros.tex}
+%\input{macros.tex}
 
@@
 
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\subsection{A Sample of Mock Lenses}
-
-Motivated by the need to accurately measure time delays for lens cosmography
-in the multi-filter optical imaging survey era, we  we selected three
-quadruply-imaged quasar lens systems from the mock LSST  catalog provided by
-\citet[][hereafter OM10]{OM10}.\footnote{The catalog is available from
-\texttt{http://kipac.stanford.edu/collab/research/lensing}}
-
-The OM10 mock lenses are designed to represent massive elliptical galaxies;
-they have singular isothermal mass distributions, with velocity dispersions
-$\vdisp$ drawn from the SDSS velocity function \citep[extrapolated to higher
-redshifts assuming no evolution, as in][]{Ogu++08}. The sources were drawn
-from a model quasar luminosity function that matches the observed bright and
-faint ends. External shears, drawn from a plausible distribution, were
-applied. The mock catalog is the product of performing the lens abundance
-calculation for the LSST survey by Monte Carlo integration; OM10 found some
-3000 well-measured lens systems to be detectable at LSST single-epoch depth
-and image quality. Here we consider a typical slice through this sample: from
-the 22 systems lying between $0.4 < \zd < 0.6$, $1.6 < \zs < 2.4$ and $150 <
-\vdisp < 300$\kms, three were chosen as being representative of the possible
-image configurations while also being particularly prone to microlensing
-effects: 6 out of the 12 images lie within the lens galaxy effective radius.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{minipage}{0.32\linewidth}
-\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3042348.eps}
-\end{minipage}
-\begin{minipage}{0.32\linewidth}
-\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3424637.eps}
-\end{minipage}
-\begin{minipage}{0.32\linewidth}
-\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=5058554.eps}
-\end{minipage}\hfill
-\caption{The three simulated lenses, showing image positions (blue) relative
-to the lens light effective area (orange). The fraction of surface mass
-density in stars ($f^{*}$) at each image position is shown under each image's\label{lenslayouts}}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
 \subsection{Intrinsic AGN Variability}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-\begin{center}
-\begin{tabular}{ccc}
-\hline
-Parameter      & Possible values \\
-\hline\hline
-$tau_q$/days   & 31.62, {\bf 316.2}, 3162  \\ 
-$sigma_q$      & 0.004, {\bf 0.008}, 0.016 \\ 
-\hline
-\end{tabular}
-\end{center}
-\caption{Simulated source parameters. The damped random walk variability model
-is controlled by two parameters, $tau_q$ and $sigma_q$; the values used here
-are representative of those observed in the SDSS stripe 82 spectroscopic 
-quasar sample. Our fiducial source lies at the centre of the resulting 3 by 3
-grid, and its variability parameters are highlighted in bold font.
-\label{tab:simsrc}}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We then apply the appropriate magnifications and time delays to generate
-4-year lightcurves for each image of our three mock lenses. We sample the
-lightcurves in several different ways, to explore the robustness of our time
-delay inference. Keeping the monitoring campaign length fixed at 4 years, we
-down-sample to a plausible cadence (separation between successive
-observations) and season length. Again, we choose representative values:
-COSMOGRAIL \citep{cosmograil} typically monitor lenses for 8 months per year
-at a cadence of 4 days, LSST will likely have seasons closer to 4 months, with
-a single-filter cadence closer to 8 days (although this is somewhat
-pessimistic). 4-month seasons with 4 day cadence mimics LSST monitoring should
-multi-filter joint analysis be possible, while 8-month seasons with 8-day
-cadence represents a cheaper version of a dedicated monitoring campaign. These
-choices resulted in 4 possible observing strategies, with total numbers of
-measurements of 60, 120, 120 and 240. A more thorough investigation of time
-delay measurement feasibility with LSST is left to future work. Four observing
-strategies, times 3 times 3 possible variable sources, lensed by 3 deflectors 
-makes a total of 108 simulated datasets, and 324 independent time delays to be
-inferred.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table*}
-\begin{center}
-\begin{tabular}{ccccc}
-\hline
-Observing strategy & Cadence (days) & Season length (days) 
-& Epochs/season & Total epochs\\
-\hline\hline
-A  & 8  & 120 & 15  &  60 \\ 
-\textbf{B}  & \textbf{4}  & \textbf{120} & \textbf{30}  & \textbf{120} \\
-C  & 8  & 240 & 30  & 120 \\
-D  & 4  & 240 & 60  & 240 \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Simulating observing strategies explored. 4 day cadence and 8 month
-seasons are an approximation to those used in the COSMOGRAIL survey, while
-LSST will provide closer to 8 day cadence and 4 month seasons. Strategy B is
-our fiducial one, corresponding approximately to (somehow) using multiple
-LSST filters to achieve a 4-day cadence over four consecutive 4-month seasons.
-\label{observing}}
-\end{table*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\TODO{PJM}{Check these  numbers, esp COSMOGRAIL...}
-
-Two example simulated datasets are shown in Figure~\ref{fig:simdata}. In  the
-next sub-section we present results from our Gaussian Process modelling of all 108
-datasets.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{minipage}{0.48\linewidth}
-\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3042348_tau=31_sigma=4_cadence=4_season=8_lightcurve.eps}
-\end{minipage}
-\begin{minipage}{0.48\linewidth}
-\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=5058554_tau=3162_sigma=16_cadence=8_season=4_lightcurve.eps}
-\end{minipage}\hfill
-\caption{Two example simulated lensed AGN datasets (with no microlensing).
-Left: a minimally-variable source ($\tau_q = 31.62$ days, $\sigma_q = 0.004$),
-lensed into a cross configuration (lens ID 3042348), and monitored with high
-cadence and long season length (4 days, 8 months, strategy D). Right: a
-maaximally-variable source ($\tau_q = 3162$ days, $\sigma_q = 0.016$), lensed
-into a major axis cusp configuration (lens ID 5058554), and monitored with low
-cadence and short season length (8 days, 4 months, strategy A).
-\label{fig:simdata}}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Results}
-
-Strategy B, dt recovery across sample. 3x3x3 = 27 inferences, 27x18 dt estimates. Are
-uncertainties on dt accurate? Trends with dt.
-
-Effect of changing observing strategy.
-
-Mean mag and var par recovery across sample. Trends with tau and sigma.
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
 \subsection{Including Microlensing}
@@
 \subsubsection{Methodology}
-
-\TODO{GD}{Write this section please!}
-\TODO{PJM/BJB}{Suitable illustrative figure?}
-
-Stellar mass distributions were ``painted on'' to the model lens galaxies, as
-described in Dobler et al.\, (in preparation). Established galaxy scaling
-relations were used to draw luminosities and effective radii given the model
-velocity dispersions; a surface brightness distribution with the same
-ellipticity as that of the lens mass, and with a de Vaucoleurs cite profile 
-were assumed. Finally, a mass to light ratio was assumed, and the fraction of
-surface mass density in the form of stars ($\fstar$) 
-computed, at each image position. This parameter, the local lens
-convergence, and its shear, were then used to generate a starfield at each
-image position, and compute the corresponding source plane magnification map,
-using the code of ...
-
-Wambsganss maps. Then, parameters chosen to span the possibilities. Constant
-speed traversals, 2 speeds (fast and slow).  2 source sizes, from Kochanek etc
-work, separated by factor of 3 (large and small).  fstar values chosen in
-catalog and in lens selection, check they are high (c. 0.5).  Check timescales
-compared to length of dataset. Illustrative plot.
-
-Apply to 3 lenses with "typical" intrinsic variabilities, ie the middle tau
-and sigma of the 3x3 grid in the previous section.
-
-\subsubsection{Results}
-
-dt recovery across sample. 3x(2x2) = 12 inferences, 12x18=216 dt estimates.
-Are uncertainties on dt accurate? Trends with dt.
-
-Mean mag and ml par recovery across sample. 48 microlensing lightcurves (4
-images times 3 lenses x 4 setups). Trends with fstar, source size and v.
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
 \subsection{Discussion}

### 2012-05-30 doc/ms.tex
No figures - paper now compiles

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\includegraphics[scale=0.75]{figs1/b1608_data.eps}
-\caption{Monitoring data for B1608+656.\label{b1608_data}}
-\end{figure*}
+%\begin{figure*}
+%\includegraphics[scale=0.75]{figs1/b1608_data.eps}
+%\caption{Monitoring data for B1608+656.\label{b1608_data}}
+%\end{figure*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\includegraphics{figs1/b1608_delays.eps}
-\caption{Inferred time delays for B1608+656.\label{b1608_delays}}
-\end{figure*}
+%\begin{figure*}
+%\includegraphics{figs1/b1608_delays.eps}
+%\caption{Inferred time delays for B1608+656.\label{b1608_delays}}
+%\end{figure*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
-\includegraphics[scale=0.42]{figs1/correlation.eps}
-\caption{Inferred intrinsic QSO variability parameters for B1608+656.
+%\begin{figure}
+%\includegraphics[scale=0.42]{figs1/correlation.eps}
+%\caption{Inferred intrinsic QSO variability parameters for B1608+656.
 This is the only significant correlation in the joint posterior distribution.
 The data are consistent with a small or large value for the variability amplitude, but if the amplitude is large, then the timescale must also be large so that we have only observed a small portion of the timescale.\label{correlation}}
-\end{figure}
+%\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-05-30 doc/ms.tex
Random edits

@@
 
-\section{The Model}
+\section{The Model for the Inference}
 
@@
 \begin{equation}
-p(\mathbf{y} | \theta) = \frac{1}{\sqrt{(2\pi)^n \det \mathbf{C}}} \exp\left[-\frac{1}{2}(\mathbf{y} - \mathbf{m})^T\mathbf{C}^{-1}(\mathbf{y} - \mathbf{m})\right]
+p(\mathbf{y} | \theta) = \frac{1}{\sqrt{(2\pi)^n \det \mathbf{C}}}
+\exp\left[-\frac{1}{2}
+(\mathbf{y} - \mathbf{m})^T\mathbf{C}^{-1}(\mathbf{y} - \mathbf{m})\right]
 \end{equation}
@@
 the GP at the times of the observations, and then adding diagonal components
-to describe the observational errors. That is to say,
-
+to describe the observational errors. That is,
 \begin{equation}
@@
 describing the factor by which the reported error bars should be multiplied.
-This is another testable mechanism for accounting for outliers.
+This is another mechanism for accounting for outliers.
 
@@
 \begin{equation}
-\left\{ \{\tau\}, \{m\}, \sigma_{\rm QSO}, L_{\rm QSO}, \{\sigma_{\rm micro}\}, \{L_{\rm micro}\}, \alpha_{\rm micro}, \kappa \right\}
+\left\{ \{\tau\}, \{m\}, \sigma_{\rm QSO}, L_{\rm QSO}, \{\sigma_{\rm micro}\}
+\{L_{\rm micro}\}, \alpha_{\rm micro}, \kappa \right\}
 \end{equation}
@@
 double image lens, the total number of parameters in the model is~12. For a
-quad, this number is~20. {\bf CHECK.}  A set of uninformative prior
+quad, this number is~20. A set of uninformative prior
 distributions for these parameters are listed in Table~\ref{priors}; we assign

### 2012-06-01 src/Python/emcee/Data.py
Started working on emcee version

@@ -1,0 +1,55 @@
+import numpy as np
+
+class Data
+	"""
+	An object of this class is a dataset in Flotsam time delay inference
+	format. Columns: (time, magnitude, sig_magnitude, image ID)
+	"""
+
+	def __init__(self):
+		"""
+		Sets loaded to False, does not construct data arrays
+		"""
+		self.loaded = False
+		pass
+
+	def load(self, filename):
+		"""
+		Loads the data from the given file.
+		"""
+		data = np.loadtxt(filename)
+		self.t = data[:,0]
+		self.y = data[:,1]
+		self.sig = data[:,2]
+		self.id = data[:,3].astype('int')
+		self.loaded = True
+
+	# Useful summary statistics
+	@property
+	def tRange(self):
+		return self.t.max() - self.t.min()
+
+	@property
+	def weights(self):
+		w = self.sig**(-2)
+		w = w/w.sum()
+		return w
+
+	@property
+	def yMean(self):
+		return (self.weights*self.y).sum()
+
+	@property
+	def ySqMean(self):
+		return (self.weights*self.y**2).sum()
+
+	@property
+	def yStDev(self):
+		return np.sqrt(self.ySqMean - self.yMean**2)
+
+if __name__ == '__main__':
+	"""
+	This is just for simple testing purposes.
+	"""
+	
+	

### 2012-06-01 src/Python/emcee/TDModel.py

@@ -1,0 +1,11 @@
+import numpy as np
+import numpy.random as rng
+import matplotlib.pyplot as plt
+
+class TDModel:
+	"""
+	An object of this class is a point in the Flotsam parameter space
+	for inferring time delays in the presence of microlensing.
+	"""
+	def __init__(self):
+		

### 2012-06-01 src/Python/emcee/Data.py
Nice plotting

@@ -1,6 +1,7 @@
 import numpy as np
+import matplotlib.pyplot as plt
 
-class Data
+class Data:
 	"""
 	An object of this class is a dataset in Flotsam time delay inference
 	format. Columns: (time, magnitude, sig_magnitude, image ID)
@@ -23,6 +24,13 @@ class Data
 		self.sig = data[:,2]
 		self.id = data[:,3].astype('int')
 		self.loaded = True
+
+	def plot(self):
+		for i in xrange(self.id.min(), self.id.max() + 1):
+			which = np.nonzero(self.id == i)[0]
+			plt.errorbar(self.t[which], self.y[which],\
+					yerr=self.sig[which], fmt='.')
+		plt.show()
 
 	# Useful summary statistics
 	@property
@@ -52,4 +60,7 @@ if __name__ == '__main__':
 	This is just for simple testing purposes.
 	"""
 	
-	
+	data = Data()
+	data.load('j1131.txt')
+	data.plot()
+

### 2012-06-01 src/Python/emcee/Data.py
Summaries aren't @properties

@@ -23,6 +23,8 @@ class Data:
 		self.y = data[:,1]
 		self.sig = data[:,2]
 		self.id = data[:,3].astype('int')
+		assert self.id.min() == 0
+		self.computeSummaries()
 		self.loaded = True
 
 	def plot(self):
@@ -32,28 +34,25 @@ class Data:
 					yerr=self.sig[which], fmt='.')
 		plt.show()
 
-	# Useful summary statistics
-	@property
-	def tRange(self):
-		return self.t.max() - self.t.min()
+	def computeSummaries(self):
+		"""
+		Compute useful summary statistics
+		"""
+		self.numImages = self.id.max() + 1
+		self.tRange = self.t.max() - self.t.min()
 
-	@property
-	def weights(self):
-		w = self.sig**(-2)
-		w = w/w.sum()
-		return w
-
-	@property
-	def yMean(self):
-		return (self.weights*self.y).sum()
-
-	@property
-	def ySqMean(self):
-		return (self.weights*self.y**2).sum()
-
-	@property
-	def yStDev(self):
-		return np.sqrt(self.ySqMean - self.yMean**2)
+		# Summary statistics for each image
+		self.yMean = np.empty(self.numImages)
+		self.ySqMean = np.empty(self.numImages)
+		self.yStDev = np.empty(self.numImages)
+		for i in xrange(0, self.numImages):
+			which = np.nonzero(self.id == i)[0]
+			w = self.sig[which]**(-2)
+			w = w/w.sum()
+			self.yMean[i] = (w*self.y[which]).sum()
+			self.ySqMean[i] = (w*self.y[which]**2).sum()
+			self.yStDev[i] = np.sqrt(self.ySqMean[i]\
+							-self.yMean[i]**2)
 
 if __name__ == '__main__':
 	"""
@@ -62,5 +61,6 @@ if __name__ == '__main__':
 	
 	data = Data()
 	data.load('j1131.txt')
+	print(data.tRange, data.yMean, data.yStDev)
 	data.plot()
 

### 2012-06-01 src/Python/emcee/TDModel.py
Limits class, other stuff

@@ -1,6 +1,22 @@
 import numpy as np
 import numpy.random as rng
 import matplotlib.pyplot as plt
+from Data import Data
+
+data = Data()
+data.load('j1131.txt')
+
+class Limits:
+	"""
+	A singleton class that just holds prior bounds
+	"""
+	mMin = data.yMean - 10.*data.yStDev
+	mMax = data.yMean + 10.*data.yStDev
+	mRange = mMax - mMin
+
+	tauMin = -data.tRange
+	tauMax =  data.tRange
+	tauRange = tauMax - tauMin
 
 class TDModel:
 	"""
@@ -8,4 +24,42 @@ class TDModel:
 	for inferring time delays in the presence of microlensing.
 	"""
 	def __init__(self):
-		
+		pass
+
+	def fromPrior(self):
+		# Mean magnitudes
+		self.m = Limits.mMin + Limits.mRange*rng.rand(data.numImages)
+
+		# Time delays
+		self.tau = Limits.tauMin\
+				+ Limits.tauRange*rng.rand(data.numImages)
+		self.tau[0] = 0.
+
+	@property
+	def logPrior(self):
+		logP = 0.
+		if np.any(np.logical_or(self.m < Limits.mMin,\
+				self.m > limits.mMax)):
+			logP = -np.inf
+		if np.any(np.logical_or(self.tau < Limits.tauMin,\
+				self.tau > limits.tauMax)):
+			logP = -np.inf
+		return logP
+
+	@property
+	def logLikelihood(self):
+		logL = 0.
+		return logL
+
+	@property
+	def vector(self):
+		"""
+		Stack all parameters into a vector
+		"""
+		return np.hstack([self.m, self.tau])
+
+if __name__ == '__main__':
+	model = TDModel()
+	model.fromPrior()
+	print(model.m)
+

### 2012-06-01 src/Python/emcee/TDModel.py
Data no longer global

@@ -3,20 +3,19 @@ import numpy.random as rng
 import matplotlib.pyplot as plt
 from Data import Data
 
-data = Data()
-data.load('j1131.txt')
-
 class Limits:
 	"""
 	A singleton class that just holds prior bounds
 	"""
-	mMin = data.yMean - 10.*data.yStDev
-	mMax = data.yMean + 10.*data.yStDev
-	mRange = mMax - mMin
+	@staticmethod
+	def initialise(data):
+		Limits.mMin = data.yMean - 10.*data.yStDev
+		Limits.mMax = data.yMean + 10.*data.yStDev
+		Limits.mRange = Limits.mMax - Limits.mMin
 
-	tauMin = -data.tRange
-	tauMax =  data.tRange
-	tauRange = tauMax - tauMin
+		Limits.tauMin = -data.tRange
+		Limits.tauMax =  data.tRange
+		Limits.tauRange = Limits.tauMax - Limits.tauMin
 
 class TDModel:
 	"""
@@ -59,6 +58,10 @@ class TDModel:
 		return np.hstack([self.m, self.tau])
 
 if __name__ == '__main__':
+	data = Data()
+	data.load('j1131.txt')
+	Limits.initialise(data)
+
 	model = TDModel()
 	model.fromPrior()
 	print(model.m)

### 2012-06-01 src/Python/emcee/TDModel.py
Can now explore the prior for a few parameters with emcee

@@ -22,26 +22,26 @@ class TDModel:
 	An object of this class is a point in the Flotsam parameter space
 	for inferring time delays in the presence of microlensing.
 	"""
-	def __init__(self):
-		pass
+	def __init__(self, numImages):
+		self.numImages = numImages
 
 	def fromPrior(self):
 		# Mean magnitudes
-		self.m = Limits.mMin + Limits.mRange*rng.rand(data.numImages)
+		self.m = Limits.mMin + Limits.mRange*rng.rand(self.numImages)
 
 		# Time delays
 		self.tau = Limits.tauMin\
-				+ Limits.tauRange*rng.rand(data.numImages)
+				+ Limits.tauRange*rng.rand(self.numImages)
 		self.tau[0] = 0.
 
 	@property
 	def logPrior(self):
 		logP = 0.
 		if np.any(np.logical_or(self.m < Limits.mMin,\
-				self.m > limits.mMax)):
+				self.m > Limits.mMax)):
 			logP = -np.inf
 		if np.any(np.logical_or(self.tau < Limits.tauMin,\
-				self.tau > limits.tauMax)):
+				self.tau > Limits.tauMax)):
 			logP = -np.inf
 		return logP
 
@@ -57,12 +57,20 @@ class TDModel:
 		"""
 		return np.hstack([self.m, self.tau])
 
+	def fromVector(self, vec):
+		"""
+		Set all parameters from the vector
+		"""
+		self.m = vec[0:self.numImages]
+		self.tau = vec[self.numImages:2*self.numImages]
+		
+
 if __name__ == '__main__':
 	data = Data()
 	data.load('j1131.txt')
 	Limits.initialise(data)
 
-	model = TDModel()
+	model = TDModel(data.numImages)
 	model.fromPrior()
 	print(model.m)
 

### 2012-06-01 src/Python/emcee/main.py

@@ -1,0 +1,44 @@
+import numpy as np
+import emcee
+from Data import Data
+from TDModel import Limits, TDModel
+
+# Define the Density
+class Flotsam(object):
+	def __init__(self, data):
+		self.data = data
+		self.model = TDModel(data.numImages)
+
+	def __call__(self, params):
+		self.model.fromVector(params)
+		return model.logPrior + model.logLikelihood
+
+data = Data()
+data.load('j1131.txt')
+Limits.initialise(data)
+
+nwalkers = 100
+
+# Make an initial guess for the positions.
+model = TDModel(data.numImages)
+model.fromPrior()
+params = np.empty((nwalkers, model.vector.size))
+for i in xrange(0, nwalkers):
+	model.fromPrior()
+	params[i, :] = model.vector
+
+# Instantiate the class
+flotsam = Flotsam(data)
+
+# The sampler object
+sampler = emcee.EnsembleSampler(nwalkers, model.vector.size, flotsam, threads=10)
+
+# Sample, outputting to a file
+f = open("flotsam.out", "w")
+for pos, prob, rstate in sampler.sample(params, iterations=2000):
+	# Write the current position to a file, one line per walker
+	f.write("\n".join(["\t".join([str(q) for q in p]) for p in pos]))
+	f.write("\n")
+	f.flush()
+f.close()
+

### 2012-06-01 src/Python/emcee/TDModel.py
Runs but acceptance prob is low (?)

@@ -9,13 +9,18 @@ class Limits:
 	"""
 	@staticmethod
 	def initialise(data):
-		Limits.mMin = data.yMean - 10.*data.yStDev
-		Limits.mMax = data.yMean + 10.*data.yStDev
-		Limits.mRange = Limits.mMax - Limits.mMin
+		Limits.magMin = data.yMean - 10.*data.yStDev
+		Limits.magMax = data.yMean + 10.*data.yStDev
+		Limits.magRange = Limits.magMax - Limits.magMin
 
 		Limits.tauMin = -data.tRange
 		Limits.tauMax =  data.tRange
 		Limits.tauRange = Limits.tauMax - Limits.tauMin
+
+		Limits.logSig_mlMin = np.log(1E-2*data.yStDev)
+		Limits.logSig_mlMax = np.log(1E2*data.yStDev)
+		Limits.logSig_mlRange = Limits.logSig_mlMax\
+						- Limits.logSig_mlMin
 
 class TDModel:
 	"""
@@ -27,21 +32,28 @@ class TDModel:
 
 	def fromPrior(self):
 		# Mean magnitudes
-		self.m = Limits.mMin + Limits.mRange*rng.rand(self.numImages)
+		self.mag = Limits.magMin + Limits.magRange*rng.rand(self.numImages)
 
 		# Time delays
 		self.tau = Limits.tauMin\
 				+ Limits.tauRange*rng.rand(self.numImages)
 		self.tau[0] = 0.
 
+		# Microlensing amplitudes
+		self.logSig_ml = Limits.logSig_mlMin + Limits.logSig_mlRange\
+					*rng.rand(self.numImages)
+
 	@property
 	def logPrior(self):
 		logP = 0.
-		if np.any(np.logical_or(self.m < Limits.mMin,\
-				self.m > Limits.mMax)):
+		if np.any(np.logical_or(self.mag < Limits.magMin,\
+				self.mag > Limits.magMax)):
 			logP = -np.inf
 		if np.any(np.logical_or(self.tau < Limits.tauMin,\
 				self.tau > Limits.tauMax)):
+			logP = -np.inf
+		if np.any(np.logical_or(self.logSig_ml < Limits.logSig_mlMin,\
+				self.logSig_ml > Limits.logSig_mlMax)):
 			logP = -np.inf
 		return logP
 
@@ -55,15 +67,15 @@ class TDModel:
 		"""
 		Stack all parameters into a vector
 		"""
[…]

### 2012-06-01 src/Python/emcee/main.py

@@ -11,7 +11,7 @@ class Flotsam(object):
 
 	def __call__(self, params):
 		self.model.fromVector(params)
-		return model.logPrior + model.logLikelihood
+		return self.model.logPrior + model.logLikelihood
 
 data = Data()
 data.load('j1131.txt')
@@ -22,6 +22,7 @@ nwalkers = 100
 # Make an initial guess for the positions.
 model = TDModel(data.numImages)
 model.fromPrior()
+
 params = np.empty((nwalkers, model.vector.size))
 for i in xrange(0, nwalkers):
 	model.fromPrior()
@@ -31,14 +32,16 @@ for i in xrange(0, nwalkers):
 flotsam = Flotsam(data)
 
 # The sampler object
-sampler = emcee.EnsembleSampler(nwalkers, model.vector.size, flotsam, threads=10)
+sampler = emcee.EnsembleSampler(nwalkers, model.vector.size, flotsam, threads=1)
 
 # Sample, outputting to a file
 f = open("flotsam.out", "w")
-for pos, prob, rstate in sampler.sample(params, iterations=2000):
+for pos, prob, rstate in sampler.sample(params, iterations=1000):
 	# Write the current position to a file, one line per walker
 	f.write("\n".join(["\t".join([str(q) for q in p]) for p in pos]))
 	f.write("\n")
 	f.flush()
+	print('Acceptance Fraction = ', sampler.acceptance_fraction)
+
 f.close()
 

### 2012-06-01 src/Python/emcee/main.py
Print acceptance frac

@@ -41,7 +41,7 @@ for pos, prob, rstate in sampler.sample(params, iterations=1000):
 	f.write("\n".join(["\t".join([str(q) for q in p]) for p in pos]))
 	f.write("\n")
 	f.flush()
-	print('Acceptance Fraction = ', sampler.acceptance_fraction)
+	print('Acceptance Fraction = %.3f'%sampler.acceptance_fraction.mean())
 
 f.close()
 

### 2012-06-01 src/Python/emcee/TDModel.py
Alpha for microlensing

@@ -17,10 +17,17 @@ class Limits:
 		Limits.tauMax =  data.tRange
 		Limits.tauRange = Limits.tauMax - Limits.tauMin
 
+		# Microlensing
 		Limits.logSig_mlMin = np.log(1E-2*data.yStDev)
 		Limits.logSig_mlMax = np.log(1E2*data.yStDev)
 		Limits.logSig_mlRange = Limits.logSig_mlMax\
 						- Limits.logSig_mlMin
+		Limits.alpha_mlMin = 1.
+		Limits.alpha_mlMax = 2.
+		Limits.alpha_mlRange = Limits.alpha_mlMax\
+						- Limits.alpha_mlMin
+
+
 
 class TDModel:
 	"""
@@ -43,6 +50,10 @@ class TDModel:
 		self.logSig_ml = Limits.logSig_mlMin + Limits.logSig_mlRange\
 					*rng.rand(self.numImages)
 
+		# Microlensing smoothness
+		self.alpha_ml = Limits.alpha_mlMin + Limits.alpha_mlRange\
+					*rng.rand()
+
 	@property
 	def logPrior(self):
 		logP = 0.
@@ -54,6 +65,9 @@ class TDModel:
 			logP = -np.inf
 		if np.any(np.logical_or(self.logSig_ml < Limits.logSig_mlMin,\
 				self.logSig_ml > Limits.logSig_mlMax)):
+			logP = -np.inf
+		if self.alpha_ml < Limits.alpha_mlMin or self.alpha_ml >\
+					Limits.alpha_mlMax:
 			logP = -np.inf
 		return logP
 
@@ -67,7 +81,8 @@ class TDModel:
 		"""
 		Stack all parameters into a vector
 		"""
-		return np.hstack([self.mag, self.tau, self.logSig_ml])
+		return np.hstack([self.mag, self.tau, self.logSig_ml,\
+					self.alpha_ml])
 
 	def fromVector(self, vec):
 		"""
@@ -76,6 +91,7 @@ class TDModel:
 		self.mag = vec[0:self.numImages]
 		self.tau = vec[self.numImages:2*self.numImages]
 		self.logSig_ml = vec[2*self.numImages:3*self.numImages]
+		self.alpha_ml = vec[3*self.numImages]
 
 if __name__ == '__main__':
 	data = Data()

### 2012-06-01 src/Python/emcee/TDModel.py
QSO variability parameters are in there

@@ -9,10 +9,12 @@ class Limits:
 	"""
 	@staticmethod
 	def initialise(data):
+		# Mean magnitudes
 		Limits.magMin = data.yMean - 10.*data.yStDev
 		Limits.magMax = data.yMean + 10.*data.yStDev
 		Limits.magRange = Limits.magMax - Limits.magMin
 
+		# Time delays
 		Limits.tauMin = -data.tRange
 		Limits.tauMax =  data.tRange
 		Limits.tauRange = Limits.tauMax - Limits.tauMin
@@ -27,7 +29,15 @@ class Limits:
 		Limits.alpha_mlRange = Limits.alpha_mlMax\
 						- Limits.alpha_mlMin
 
-
+		# QSO Variability
+		Limits.logSig_qsoMin = np.log(1E-2*data.yStDev).mean()
+		Limits.logSig_qsoMax = np.log(1E2*data.yStDev).mean()
+		Limits.logSig_qsoRange = Limits.logSig_qsoMax\
+						- Limits.logSig_qsoMin
+		Limits.logTau_qsoMin = np.log(1E-2*data.tRange)
+		Limits.logTau_qsoMax = np.log(1E3*data.tRange)
+		Limits.logTau_qsoRange = Limits.logTau_qsoMax\
+						- Limits.logTau_qsoMin
 
 class TDModel:
 	"""
@@ -54,6 +64,12 @@ class TDModel:
 		self.alpha_ml = Limits.alpha_mlMin + Limits.alpha_mlRange\
 					*rng.rand()
 
+		# QSO Variability
+		self.logSig_qso = Limits.logSig_qsoMin +\
+					Limits.logSig_qsoRange*rng.rand()
+		self.logTau_qso = Limits.logTau_qsoMin +\
+					Limits.logTau_qsoRange*rng.rand()
+
 	@property
 	def logPrior(self):
 		logP = 0.
@@ -69,6 +85,12 @@ class TDModel:
 		if self.alpha_ml < Limits.alpha_mlMin or self.alpha_ml >\
 					Limits.alpha_mlMax:
 			logP = -np.inf
+		if self.logSig_qso < Limits.logSig_qsoMin or\
+			self.logSig_qso > Limits.logSig_qsoMax:
+			logP = -np.inf
+		if self.logTau_qso < Limits.logTau_qsoMin or\
+			self.logTau_qso > Limits.logTau_qsoMax:
+			logP = -np.inf
 		return logP
 
 	@property
[…]

### 2012-06-01 src/Python/emcee/display.py
Check prior exploration

@@ -1,0 +1,10 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import time
+
+sample = np.loadtxt('flotsam.out')
+for i in xrange(0, sample.shape[1]):
+	plt.hist(sample[:,i], 20)
+	plt.title(i)
+	plt.show()
+

### 2012-06-01 src/Python/emcee/main.py
Fixes

@@ -11,7 +11,7 @@ class Flotsam(object):
 
 	def __call__(self, params):
 		self.model.fromVector(params)
-		return self.model.logPrior + model.logLikelihood
+		return self.model.logPrior + self.model.logLikelihood(data)
 
 data = Data()
 data.load('j1131.txt')
@@ -20,19 +20,19 @@ Limits.initialise(data)
 nwalkers = 100
 
 # Make an initial guess for the positions.
-model = TDModel(data.numImages)
-model.fromPrior()
+m = TDModel(data.numImages)
+m.fromPrior()
 
-params = np.empty((nwalkers, model.vector.size))
+params = np.empty((nwalkers, m.vector.size))
 for i in xrange(0, nwalkers):
-	model.fromPrior()
-	params[i, :] = model.vector
+	m.fromPrior()
+	params[i, :] = m.vector
 
 # Instantiate the class
 flotsam = Flotsam(data)
 
 # The sampler object
-sampler = emcee.EnsembleSampler(nwalkers, model.vector.size, flotsam, threads=1)
+sampler = emcee.EnsembleSampler(nwalkers, m.vector.size, flotsam, threads=1)
 
 # Sample, outputting to a file
 f = open("flotsam.out", "w")

### 2012-06-01 src/Python/emcee/TDModel.py
Working towards likelihood

@@ -93,8 +93,23 @@ class TDModel:
 			logP = -np.inf
 		return logP
 
-	@property
-	def logLikelihood(self):
+	def logLikelihood(self, data):
+		assert self.numImages == data.numImages
+
+		which = [np.nonzero(data.id == i)[0]\
+				for i in xrange(0, self.numImages)]
+
+		# Mean vector
+		m = np.empty(data.t.size)
+		for i in xrange(0, self.numImages):
+			m[which[i]] = self.mag[i]
+
+		# Covariance matrix
+		C = np.zeros((data.t.size, data.t.size))
+		ids = np.meshgrid(data.id, data.id)
+
+		equal = ids[0] == ids[1]
+
 		logL = 0.
 		return logL
 

### 2012-06-02 src/Python/emcee/TDModel.py
Dumped in Linear Algebra stuff

@@ -1,5 +1,6 @@
 import numpy as np
 import numpy.random as rng
+import scipy.linalg as la
 import matplotlib.pyplot as plt
 from Data import Data
 
@@ -110,8 +111,16 @@ class TDModel:
 
 		equal = ids[0] == ids[1]
 
-		logL = 0.
-		return logL
+##		try:
+##			L = la.cholesky(C)
+##		except:
+##			return -np.inf
+#		y = data.y - m
+#		logDeterminant = 2.0*np.sum(np.log(np.diag(L)))
+#		solution = la.cho_solve((L, True), y)
+#		exponent = np.dot(y, solution)
+#		logL = -0.5*data.t.size*np.log(2.0*np.pi) - 0.5*logDeterminant - 0.5*exponent
+		return 0.
 
 	@property
 	def vector(self):

### 2012-06-03 src/Python/emcee/TDModel.py
Lags

@@ -106,8 +106,12 @@ class TDModel:
 			m[which[i]] = self.mag[i]
 
 		# Covariance matrix
-		C = np.zeros((data.t.size, data.t.size))
+		[t1, t2] = np.meshgrid(data.t, data.t)
+		lags = np.abs(t2 - t1)
+		plt.imshow(lags)
+		plt.show()
 		ids = np.meshgrid(data.id, data.id)
+
 
 		equal = ids[0] == ids[1]
 

### 2012-06-03 src/DNest/Data.cpp
Tidied up Data class

@@ -3,6 +3,7 @@
 #include <iostream>
 #include <cstdlib>
 #include <algorithm>
+#include <cmath>
 
 using namespace std;
 
@@ -23,48 +24,70 @@ void Data::load(const char* filename)
 
 	if(!fin)
 	{	
-		cerr<<"Error opening "<<filename<<"."<<endl;
+		cerr<<"# Error opening "<<filename<<"."<<endl;
 		exit(0);
 	}
 
 	// Skip past comments
 	while(fin.peek() == '#')
-		fin.ignore(1024*1024, '\n');
+		fin.ignore(1048576, '\n');
 
 	while(fin>>temp1 && fin>>temp2 && fin>>temp3 && fin>>temp4)
 	{
 		t.push_back(temp1);
 		y.push_back(temp2);
 		sig.push_back(temp3);
-		qsoID.push_back((int)temp4);
+		ID.push_back(static_cast<int>(temp4));
 	}
 	cout<<"# Loaded "<<t.size()<<" points from file "<<filename<<"."<<endl;
-
 	fin.close();
 
-	N = t.size();
-	numQSOs = 1 + *max_element(qsoID.begin(), qsoID.end());
-	tMin = *min_element(t.begin(), t.end());
-	tRange = *max_element(t.begin(), t.end()) - tMin;
-	yMin = *min_element(y.begin(), y.end());
-	yRange = *max_element(y.begin(), y.end()) - yMin;
-
+	computeSummaries();
 	loaded = true;
 }
 
-double Data::getTMin() const
+void Data::computeSummaries()
 {
-	return tMin;
-}
+	cout<<"# Summary statistics :"<<endl;
 
-double Data::getTRange() const
-{
-	return tRange;
-}
+	numPoints = t.size();
+	numImages = 1 + *max_element(ID.begin(), ID.end());
+	tMin = *min_element(t.begin(), t.end());
+	tMax = *max_element(t.begin(), t.end());
+	tRange = tMax - tMin;
+	cout<<"# Number of images = "<<numImages<<endl;
+	cout<<"# Time range = "<<tRange<<endl;
 
-int Data::getN() const
-{
-	return N;
-}
+	yMean.assign(numImages, 0.);
+	vector<double> ySqMean(numImages, 0.);
+	yStDev.assign(numImages, 0.);
 
[…]

### 2012-06-03 src/DNest/Data.h

@@ -5,25 +5,29 @@
 
 class Data
 {
-	friend class TDModel;
-
 	private:
+		// The data
 		std::vector<double> t, y, sig;
-		std::vector<int> qsoID;
+		std::vector<int> ID;
 
 		bool loaded;
 
-		int N;
-		int numQSOs;
-		double tMin, tRange;
-		double yMin, yRange;
+		// Some summary statistics
+		int numPoints, numImages;
+		double tMin, tMax, tRange;
+		std::vector<double> yMean, yStDev; // One for each image
+
+		void computeSummaries();
 
 	public:
 		Data();
 		void load(const char* filename);
-		double getTMin() const;
-		double getTRange() const;
-		int getN() const;
+
+		// A bunch of getters
+		double get_tMin() const { return tMin; }
+		double get_tRange() const { return tRange; }
+		int get_numPoints() const { return numPoints; }
+		int get_numImages() const { return numImages; }
 
 };
 

### 2012-06-03 src/DNest/Data.cpp
Static instance

@@ -7,6 +7,8 @@
 
 using namespace std;
 
+Data Data::instance;
+
 Data::Data()
 :loaded(false)
 {

### 2012-06-03 src/DNest/Data.h

@@ -19,6 +19,9 @@ class Data
 
 		void computeSummaries();
 
+		// Singleton
+		static Data instance;
+
 	public:
 		Data();
 		void load(const char* filename);
@@ -28,6 +31,8 @@ class Data
 		double get_tRange() const { return tRange; }
 		int get_numPoints() const { return numPoints; }
 		int get_numImages() const { return numImages; }
+
+		static Data get_instance() { return instance; }
 
 };
 

### 2012-06-03 src/DNest/main.cpp
DNest3 style main

@@ -1,120 +1,24 @@
-/*
-    Copyright (C) 2011 Brendon J. Brewer
-    This file is part of DNest, the Diffusive Nested Sampler.
-
-    DNest is free software: you can redistribute it and/or modify
-    it under the terms of the GNU General Public License as published by
-    the Free Software Foundation, either version 3 of the License, or
-    (at your option) any later version.
-
-    DNest is distributed in the hope that it will be useful,
-    but WITHOUT ANY WARRANTY; without even the implied warranty of
-    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-    GNU General Public License for more details.
-
-    You should have received a copy of the GNU General Public License
-    along with DNest.  If not, see <http://www.gnu.org/licenses/>.
-*/
-
-
 #include <iostream>
+#include "Start.h"
+#include "Data.h"
 #include "TDModel.h"
-#include "MultiSampler.h"
-#include "RandomNumberGenerator.h"
-#include "DNestSampler.h"
-#include <boost/thread.hpp>
-#include <boost/program_options.hpp>
 
 using namespace std;
-using namespace DNest;
-namespace po = boost::program_options;
-
-void runMultiThread(const Model* exampleModel, const string& levelsFile);				// Run multithreaded sampler
-void runLegacy(const Model* exampleModel, const string& levelsFile);	// Run legacy sampler
+using namespace DNest3;
 
 int main(int argc, char** argv)
 {
-	// What kind of model will we be running?
-	TDModel::loadData("Data/j1131.txt");
-	TDModel exampleModel;
+	// Process command line options and load data
+	CommandLineOptions options(argc, argv);
+	string dataFile = options.get_dataFile();
+	Data::get_instance().load(dataFile.c_str());
 
+	// Initialise the sampler
+	Sampler<TDModel> sampler = setup<TDModel>(options);
 
-	// Which kind of sampler, and should we pre-load level structure?
-	bool legacy = false;
-	string levelsFile = "";
-
-	// Handle command line options
-	{
-		// Declare the supported options.
-		po::options_description desc("Welcome to DNest. Allowed options");
-		desc.add_options()
-		    ("help", "produce help message")
[…]

### 2012-06-03 src/DNest/TDModel.h
DNest3 format

@@ -4,58 +4,62 @@
 #include "Model.h"
 #include "Data.h"
 #include <ostream>
-#include <istream>
 #include <vector>
 #include <Eigen/Dense>
 
-class TDModel:public DNest::Model
+// Shorter names for Eigen3 types
+typedef Eigen::VectorXd Vector;
+typedef Eigen::MatrixXd Matrix;
+typedef Eigen::LLT< Eigen::MatrixXd > Cholesky;
+
+class TDModel:public DNest3::Model
 {
 	private:
-		std::vector<double> timeDelays;
-		std::vector<double> sigMicrolensing;
-		std::vector<double> tauMicrolensing;
-		std::vector<double> meanMagnitudes;
+		// Mean magnitudes
+		std::vector<double> mag;
 
-		double sigIntrinsic, tauIntrinsic, alphaIntrinsic, alphaMicrolensing, sigmaBoost;
-		Eigen::MatrixXd covMat;
-		Eigen::LLT< Eigen::MatrixXd > cholesky;
+		// Time delays (first is zero by definition)
+		std::vector<double> tau;		
 
-		static const double minSig;
-		static const double maxSig;
+		// Microlensing amplitudes
+		std::vector<double> logSig_ml;
 
-		static Data data;
+		// Microlensing smoothness
+		double alpha;
 
-		void formCovarianceMatrix();
+		// QSO Variability amplitude
+		double logSig_qso;
 
-		double perturbHelper1();
-		double perturbHelper2();
-		double perturbHelper3();
-		double perturbHelper4();
-		double perturbHelper5();
-		double perturbHelper6();
-		double perturbHelper7();
-		double perturbHelper8();
-		double perturbHelper9();
+		// QSO variability timescale
+		double logTau_qso;
+
+		// Mean vector
+		Vector meanVector;
+
+		// Covariance matrix and its Cholesky decomposition
+		Matrix covarianceMatrix;
+		Cholesky cholesky;
 
 		// Noise-free covariance function
-		double covariance(double t1, double t2, int qsoID1, int qsoID2);
+		double covariance(double t1, double t2, int ID1, int ID2);
 
 	public:
 		TDModel();
-		~TDModel();
-		DNest::Model* factory() const;
-		DNest::Model* clone() const;
-		void copyFrom(const DNest::Model* other);
+
+		// Generate the point from the prior
 		void fromPrior();
-		void calculateLogLikelihood();
+
+		// Metropolis-Hastings proposals
 		double perturb();
-		double getValue();
[…]

### 2012-06-03 src/DNest/Limits.h
Limits class

@@ -1,0 +1,45 @@
+#ifndef _Limits_
+#define _Limits_
+
+#include "Data.h"
+
+/*
+* Define prior bounds on the parameters.
+*/
+
+class Limits
+{
+	private:
+		// Naming convention: varName_min, varName_max, varName_range
+
+		// Mean magnitudes
+		std::vector<double> mag_min, mag_max, mag_range;
+
+		// Time delays (first is zero by definition)
+		std::vector<double> tau_min, tau_max, tau_range;
+
+		// Microlensing amplitudes
+		std::vector<double> logSig_ml_min, logSig_ml_max,
+						logSig_ml_range;
+
+		// Microlensing smoothness
+		double alpha_min, alpha_max, alpha_range;
+
+		// QSO Variability amplitude
+		double logSig_qso_min, logSig_qso_max, logSig_qso_range;
+
+		// QSO variability timescale
+		double logTau_qso_min, logTau_qso_max, logTau_qso_range;
+		
+
+	public:
+		// Default constructor: Do nothing
+		Limits();
+
+		// Use data to construct limits
+		void set(const Data& data);
+
+};
+
+#endif
+

### 2012-06-03 src/DNest/Limits.h
TDModel is a friend

@@ -9,6 +9,8 @@
 
 class Limits
 {
+	friend class TDModel;
+
 	private:
 		// Naming convention: varName_min, varName_max, varName_range
 

### 2012-06-04 src/DNest/Data.h
Some more getters

@@ -31,6 +31,8 @@ class Data
 		double get_tRange() const { return tRange; }
 		int get_numPoints() const { return numPoints; }
 		int get_numImages() const { return numImages; }
+		const std::vector<double>& get_yMean() const { return yMean; }
+		const std::vector<double>& get_yStDev() const { return yStDev; }
 
 		static Data get_instance() { return instance; }
 

### 2012-06-04 src/DNest/TDModel.h
static Limits instance

@@ -15,6 +15,9 @@ typedef Eigen::LLT< Eigen::MatrixXd > Cholesky;
 class TDModel:public DNest3::Model
 {
 	private:
+		// Prior bounds on everything
+		static Limits limits;
+
 		// Mean magnitudes
 		std::vector<double> mag;
 

### 2012-06-04 src/DNest/Limits.cpp
Started work on set()

@@ -1,0 +1,23 @@
+#include "Limits.h"
+
+Limits::Limits()
+:isSet(false)
+{
+
+}
+
+void Limits::set(const Data& data)
+{
+	mag_min.resize(data.get_numImages());
+	mag_max.resize(data.get_numImages());
+	mag_range.resize(data.get_numImages());
+	for(int i=0; i<data.get_numImages(); i++)
+	{
+		mag_min[i] = data.get_yMean()[i] - 10.*data.get_yStDev()[i]; 
+		mag_max[i] = data.get_yMean()[i] + 10.*data.get_yStDev()[i];
+		mag_range[i] = mag_max[i] - mag_min[i];
+	}
+
+	isSet = true;
+}
+

### 2012-06-04 src/DNest/Limits.h

@@ -32,7 +32,9 @@ class Limits
 
 		// QSO variability timescale
 		double logTau_qso_min, logTau_qso_max, logTau_qso_range;
-		
+
+		// Whether set() has been called
+		bool isSet;
 
 	public:
 		// Default constructor: Do nothing

### 2012-06-04 src/DNest/Limits.cpp
tau limits

@@ -18,6 +18,16 @@ void Limits::set(const Data& data)
 		mag_range[i] = mag_max[i] - mag_min[i];
 	}
 
+	tau_min.resize(data.get_numImages());
+	tau_max.resize(data.get_numImages());
+	tau_range.resize(data.get_numImages());
+	for(int i=0; i<data.get_numImages(); i++)
+	{
+		tau_min[i] = -data.get_tRange();
+		tau_max[i] =  data.get_tRange();
+		tau_range[i] = tau_max[i] - tau_min[i];
+	}
+
 	isSet = true;
 }
 

### 2012-06-04 src/DNest3/Limits.cpp
More limits

@@ -1,4 +1,5 @@
 #include "Limits.h"
+#include <cmath>
 
 Limits::Limits()
 :isSet(false)
@@ -26,6 +27,16 @@ void Limits::set(const Data& data)
 		tau_min[i] = -data.get_tRange();
 		tau_max[i] =  data.get_tRange();
 		tau_range[i] = tau_max[i] - tau_min[i];
+	}
+
+	logSig_ml_min.resize(data.get_numImages());
+	logSig_ml_max.resize(data.get_numImages());
+	logSig_ml_range.resize(data.get_numImages());
+	for(int i=0; i<data.get_numImages(); i++)
+	{
+		logSig_ml_min[i] = log(1E-2*data.get_yStDev()[i]);
+		logSig_ml_max[i] = log(1E+2*data.get_yStDev()[i]);
+		logSig_ml_range[i] = logSig_ml_max[i] - logSig_ml_min[i];
 	}
 
 	isSet = true;

### 2012-06-04 src/DNest3/Limits.cpp
More limits

@@ -39,6 +39,23 @@ void Limits::set(const Data& data)
 		logSig_ml_range[i] = logSig_ml_max[i] - logSig_ml_min[i];
 	}
 
+	alpha_min = 1.;
+	alpha_max = 2.;
+	alpha_range = alpha_max - alpha_min;
+
+	logSig_qso_min = 0.;
+	logSig_qso_max = 0.;
+	for(int i=0; i<data.get_numImages(); i++)
+	{
+		logSig_qso_min += 1E-2*data.get_yStDev()[i]/data.get_numImages();
+		logSig_qso_max += 1E+2*data.get_yStDev()[i]/data.get_numImages();
+	}
+	logSig_qso_range = logSig_qso_max - logSig_qso_min;
+
+	logTau_qso_min = log(1E-2*data.get_tRange());
+	logTau_qso_max = log(1E+3*data.get_tRange());
+	logTau_qso_range = logTau_qso_max - logTau_qso_min;
+
 	isSet = true;
 }
 

### 2012-06-04 src/DNest3/Data.h
One more getter

@@ -33,6 +33,7 @@ class Data
 		int get_numImages() const { return numImages; }
 		const std::vector<double>& get_yMean() const { return yMean; }
 		const std::vector<double>& get_yStDev() const { return yStDev; }
+		bool get_loaded() const { return loaded; }
 
 		static Data get_instance() { return instance; }
 

### 2012-06-04 src/DNest3/TDModel.h
include Limits.h

@@ -3,6 +3,7 @@
 
 #include "Model.h"
 #include "Data.h"
+#include "Limits.h"
 #include <ostream>
 #include <vector>
 #include <Eigen/Dense>

### 2012-06-04 src/DNest3/TDModel.cpp
fromPrior

@@ -1,75 +1,37 @@
 #include "TDModel.h"
-#include "Utils.h"
+#include "Data.h"
 #include "RandomNumberGenerator.h"
-#include <cmath>
 #include <iostream>
-#include <Eigen/Dense>
 
-using namespace Eigen;
 using namespace std;
-using namespace DNest;
-
-Data TDModel::data;
-
-const double TDModel::minSig = 1E-3;
-const double TDModel::maxSig = 10.;
+using namespace DNest3;
 
 TDModel::TDModel()
-:timeDelays(data.numQSOs), sigMicrolensing(data.numQSOs), tauMicrolensing(data.numQSOs), meanMagnitudes(data.numQSOs)
-,covMat(data.N, data.N)
 {
-	if(!data.loaded)
+	if(!Data::get_instance().get_loaded())
 	{
-		cerr<<"Data has not been loaded! Cannot construct TDModel"<<endl;
+		cerr<<"# Data has not been loaded! Cannot construct TDModel"<<endl;
 		exit(0);
 	}
 }
 
-TDModel::~TDModel()
-{
-
-}
-
-DNest::Model* TDModel::factory() const
-{
-	return new TDModel;
-}
-
-DNest::Model* TDModel::clone() const
-{
-	return new TDModel(*this);
-}
-
-void TDModel::copyFrom(const DNest::Model* other)
-{
-	*this = *((TDModel*)other);
-}
-
 void TDModel::fromPrior()
 {
-	for(size_t i=0; i<timeDelays.size(); i++)
-		timeDelays[i] = (i==0)?(0):(data.tRange*(-0.5 + randomU()));
-	for(size_t i=0; i<sigMicrolensing.size(); i++)
-		sigMicrolensing[i] = exp(log(minSig) + log(maxSig/minSig)*randomU());
-	for(size_t i=0; i<tauMicrolensing.size(); i++)
-		tauMicrolensing[i] = exp(log(1E-2*data.tRange) + log(1E4)*randomU());
+	for(int i=0; i<Data::get_instance().get_numImages(); i++)
+	{
+		mag[i] = limits.mag_min[i] + limits.mag_range[i]*randomU();
+		tau[i] = limits.tau_min[i] + limits.tau_range[i]*randomU();
+		logSig_ml[i] = limits.logSig_ml_min[i]
+					+ limits.logSig_ml_range[i]*randomU();
+	}
 
[…]

### 2012-06-04 src/DNest3/Limits.cpp
Forgot tau_ml

@@ -39,6 +39,16 @@ void Limits::set(const Data& data)
 		logSig_ml_range[i] = logSig_ml_max[i] - logSig_ml_min[i];
 	}
 
+	logTau_ml_min.resize(data.get_numImages());
+	logTau_ml_max.resize(data.get_numImages());
+	logTau_ml_range.resize(data.get_numImages());
+	for(int i=0; i<data.get_numImages(); i++)
+	{
+		logTau_ml_min[i] = log(1E-2*data.get_tRange());
+		logTau_ml_max[i] = log(1E+3*data.get_tRange());
+		logTau_ml_range[i] = logSig_ml_max[i] - logSig_ml_min[i];
+	}
+
 	alpha_min = 1.;
 	alpha_max = 2.;
 	alpha_range = alpha_max - alpha_min;

### 2012-06-04 src/DNest3/Limits.h

@@ -24,6 +24,10 @@ class Limits
 		std::vector<double> logSig_ml_min, logSig_ml_max,
 						logSig_ml_range;
 
+		// Microlensing timescales
+		std::vector<double> logTau_ml_min, logTau_ml_max,
+						logTau_ml_range;
+
 		// Microlensing smoothness
 		double alpha_min, alpha_max, alpha_range;
 

### 2012-06-04 src/DNest3/TDModel.h
logTau_ml

@@ -28,6 +28,9 @@ class TDModel:public DNest3::Model
 		// Microlensing amplitudes
 		std::vector<double> logSig_ml;
 
+		// Microlensing timescales
+		std::vector<double> logTau_ml;
+
 		// Microlensing smoothness
 		double alpha;
 

### 2012-06-04 src/DNest3/TDModel.cpp
covariance method

@@ -23,6 +23,8 @@ void TDModel::fromPrior()
 		tau[i] = limits.tau_min[i] + limits.tau_range[i]*randomU();
 		logSig_ml[i] = limits.logSig_ml_min[i]
 					+ limits.logSig_ml_range[i]*randomU();
+		logTau_ml[i] = limits.logTau_ml_min[i]
+					+ limits.logTau_ml_range[i]*randomU();
 	}
 
 	alpha = limits.alpha_min + limits.alpha_range*randomU();
@@ -31,7 +33,7 @@ void TDModel::fromPrior()
 }
 
 
-/*double TDModel::perturbHelper1()
+/*double TDModel::perturb1()
 {
 	int which = 1+randInt(timeDelays.size()-1);
 	timeDelays[which] += data.tRange*pow(10.0, 1.5-6*randomU())*randn();
@@ -39,7 +41,7 @@ void TDModel::fromPrior()
 	return 0;
 }
 
-double TDModel::perturbHelper2()
+double TDModel::perturb2()
 {
 	int which = randInt(sigMicrolensing.size());
 	double temp = log(sigMicrolensing[which]);
@@ -49,7 +51,7 @@ double TDModel::perturbHelper2()
 	return 0.0;
 }
 
-double TDModel::perturbHelper3()
+double TDModel::perturb3()
 {
 	int which = randInt(tauMicrolensing.size());
 	double temp = log(tauMicrolensing[which]);
@@ -59,7 +61,7 @@ double TDModel::perturbHelper3()
 	return 0.0;
 }
 
-double TDModel::perturbHelper4()
+double TDModel::perturb4()
 {
 	int which = randInt(meanMagnitudes.size());
 	meanMagnitudes[which] += data.yRange*pow(10.0, 1.5-6*randomU())*randn();
@@ -67,7 +69,7 @@ double TDModel::perturbHelper4()
 	return 0.0;
 }
 
-double TDModel::perturbHelper5()
+double TDModel::perturb5()
 {
 	double temp = log(sigIntrinsic);
 	temp += log(maxSig/minSig)*pow(10.0, 1.5-6*randomU())*randn();
@@ -76,7 +78,7 @@ double TDModel::perturbHelper5()
 	return 0.0;	
 }
 
-double TDModel::perturbHelper6()
+double TDModel::perturb6()
 {
 	double temp = log(tauIntrinsic);
[…]

### 2013-04-16 src/DNest3/Limits.cpp
Found bug in limits, also extended some

@@ -1,5 +1,8 @@
 #include "Limits.h"
 #include <cmath>
+#include <iostream>
+
+using namespace std;
 
 Limits::Limits()
 :isSet(false)
@@ -14,8 +17,8 @@ void Limits::set(const Data& data)
 	mag_range.resize(data.get_numImages());
 	for(int i=0; i<data.get_numImages(); i++)
 	{
-		mag_min[i] = data.get_yMean()[i] - 10.*data.get_yStDev()[i]; 
-		mag_max[i] = data.get_yMean()[i] + 10.*data.get_yStDev()[i];
+		mag_min[i] = data.get_yMean()[i] - 30.*data.get_yStDev()[i]; 
+		mag_max[i] = data.get_yMean()[i] + 30.*data.get_yStDev()[i];
 		mag_range[i] = mag_max[i] - mag_min[i];
 	}
 
@@ -28,8 +31,8 @@ void Limits::set(const Data& data)
 	logSig_ml_range.resize(data.get_numImages());
 	for(int i=0; i<data.get_numImages(); i++)
 	{
-		logSig_ml_min[i] = log(1E-2*data.get_yStDev()[i]);
-		logSig_ml_max[i] = log(1E+2*data.get_yStDev()[i]);
+		logSig_ml_min[i] = log(1E-3*data.get_yStDev()[i]);
+		logSig_ml_max[i] = log(1E+3*data.get_yStDev()[i]);
 		logSig_ml_range[i] = logSig_ml_max[i] - logSig_ml_min[i];
 	}
 
@@ -38,7 +41,7 @@ void Limits::set(const Data& data)
 	logTau_ml_range.resize(data.get_numImages());
 	for(int i=0; i<data.get_numImages(); i++)
 	{
-		logTau_ml_min[i] = log(1E-2*data.get_tRange());
+		logTau_ml_min[i] = log(1E-3*data.get_tRange());
 		logTau_ml_max[i] = log(1E+3*data.get_tRange());
 		logTau_ml_range[i] = logSig_ml_max[i] - logSig_ml_min[i];
 	}
@@ -51,12 +54,16 @@ void Limits::set(const Data& data)
 	logSig_qso_max = 0.;
 	for(int i=0; i<data.get_numImages(); i++)
 	{
-		logSig_qso_min += 1E-2*data.get_yStDev()[i]/data.get_numImages();
[…]

### 2013-04-16 doc/Leon/ms.tex
Writing some words

% Every Latex document starts with a documentclass command
\documentclass[a4paper, 11pt]{article}

% Load some packages
\usepackage{graphicx} % This allows you to put figures in
\usepackage{natbib}   % This allows for relatively pain-free reference lists
\usepackage[left=3cm,top=3cm,right=3cm]{geometry} % The way I like the margins

% This helps with figure placement
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm
\newcommand{\yy}{\mathbf{y}}
\newcommand{\mm}{\mathbf{m}_\theta}
\newcommand{\CC}{\mathbf{C}_\theta}

% Set values so you can have a title
\title{}
\author{}
\date{} % Otherwise, it'll show today's date

% Document starts here
\begin{document}

Suppose that the underlying QSO variability as a function of time is
described by:
\begin{eqnarray}
y_{\rm QSO}(t)
\end{eqnarray}
Then the light curves of the two images $A$ and $B$ will be given by:
\begin{eqnarray}
y_A(t) &=& y_{\rm QSO}(t) + \mu_A(t)\\
y_B(t) &=& y_{\rm QSO}(t - \tau) + \mu_B(t)
\end{eqnarray}
where $\mu_A(t)$ and $\mu_B(t)$ are microlensing perturbations.

The data for all images are in magnitudes:
\begin{eqnarray}
\yy = \{y_1, y_2, ..., y_N\}
\end{eqnarray}

The probability distribution for the data $\yy$ given the parameters
$\theta$
is a multivariate normal distribution, where the mean vector $\mm$
and the covariance matrix $\CC$ depend on the parameters:
\begin{eqnarray}
p(\yy | \theta) &=& \frac{1}{\sqrt{(2\pi)^N\det{\CC}}}
\exp\left[
-\frac{1}{2}
\left(
\yy - \mm
\right)^T
\CC^{-1}
\left(
\yy - \mm
\right)
\right]
\end{eqnarray}

Numerically, the logarithm of the determinant and the
$\CC^{-1}\left(\yy - \mm\right)$ term are implemented using
Cholesky decompositions.

\end{document}

### 2013-04-21 src/DNest3/TDModel.cpp
Better proposal

@@ -193,39 +193,43 @@ double TDModel::perturb()
 {
 	double logH = 0.;
 
-	// Propose things that are many-parameter more often
-	int which = randInt(9);
-
-	switch(which)
+	// Flag -- whether to do each kind of proposal
+	vector<bool> change(9, false);
+	int num = 0;
+	double chance = pow(10., 0.5 - 3.*randomU());
+	for(size_t i=0; i<change.size(); i++)
 	{
-		case 0:
-			logH += perturb1();
-			break;
-		case 1:
-			logH += perturb2();
-			break;
-		case 2:
-			logH += perturb3();
-			break;
-		case 3:
-			logH += perturb4();
-			break;
-		case 4:
-			logH += perturb5();
-			break;
-		case 5:
-			logH += perturb6();
-			break;
-		case 6:
-			logH += perturb7();
-			break;
-		case 7:
-			logH += perturb8();
-			break;
-		case 8:
-			logH += perturb9();
-			break;
+		if(randomU() <= chance)
+		{
+			change[i] = true;
+			num++;
+		}
 	}
+	if(num == 0)
+	{
+		change[randInt(change.size())] = true;
+		num++; 
+	}
+
+	if(change[0])
+		logH += perturb1();
+	if(change[1])
+		logH += perturb2();
+	if(change[2])
+		logH += perturb3();
+	if(change[3])
+		logH += perturb4();
+	if(change[4])
+		logH += perturb5();
+	if(change[5])
+		logH += perturb6();
+	if(change[6])
+		logH += perturb7();
+	if(change[7])
+		logH += perturb8();
+	if(change[8])
+		logH += perturb9();
+
 
 	formMeanVector();
 	formCovarianceMatrix();

### 2013-04-21 src/DNest3/TDModel.cpp
Fixed description string

@@ -349,6 +349,6 @@ string TDModel::description() const
 {
 	return string("mag, tau, logSig_ml, logTau_ml, alpha")
 			+ string(", logSig_qso, logTau_qso")
-			+ string(", meanLogSigmaBoost, logStdevLogSigmaBoost");
+			+ string(", f_bad, boost");
 }
 

### 2013-04-21 src/DNest3/TDModel.cpp
Read method

@@ -345,6 +345,24 @@ void TDModel::print(ostream& out) const
 	out<<f_bad<<' '<<boost<<' ';
 }
 
+void TDModel::read(istream& in)
+{
+	for(int i=0; i<numImages; i++)
+		in>>mag[i];
+	for(int i=0; i<numImages; i++)
+		in>>tau[i];
+	for(int i=0; i<numImages; i++)
+		in>>logSig_ml[i];
+	for(int i=0; i<numImages; i++)
+		in>>logTau_ml[i];
+
+	in>>alpha;
+	in>>logSig_qso;
+	in>>logTau_qso;
+	in>>f_bad;
+	in>>boost;
+}
+
 string TDModel::description() const
 {
 	return string("mag, tau, logSig_ml, logTau_ml, alpha")

### 2013-04-21 src/DNest3/TDModel.h

@@ -4,7 +4,7 @@
 #include "Model.h"
 #include "Data.h"
 #include "Limits.h"
-#include <ostream>
+#include <iostream>
 #include <vector>
 #include "Matrix.h"
 #include "Vector.h"
@@ -81,6 +81,9 @@ class TDModel:public DNest3::Model
 
 		// Print to stream
 		void print(std::ostream& out) const;
+
+		// Read from stream
+		void read(std::istream& in);
 
 		// Return string with column information
 		std::string description() const;

### 2013-04-21 src/DNest3/TDModel.cpp
Print2 method

@@ -3,6 +3,7 @@
 #include "Utils.h"
 #include "RandomNumberGenerator.h"
 #include <iostream>
+#include <iomanip>
 #include <cmath>
 #include <gsl/gsl_linalg.h>
 
@@ -343,6 +344,17 @@ void TDModel::print(ostream& out) const
 	out<<logSig_qso<<' ';
 	out<<logTau_qso<<' ';
 	out<<f_bad<<' '<<boost<<' ';
+}
+
+void TDModel::print2(ostream& out) const
+{
+	out<<setprecision(12);
+	// Write out mean vector and covariance matrix
+	for(int i=0; i<numPoints; i++)
+		out<<meanVector(i)<<' ';
+	for(int i=0; i<numPoints; i++)
+		for(int j=0; j<numPoints; j++)
+			out<<covarianceMatrix(i, j)<<' ';
 }
 
 void TDModel::read(istream& in)

### 2013-04-21 src/DNest3/TDModel.h

@@ -82,6 +82,9 @@ class TDModel:public DNest3::Model
 		// Print to stream
 		void print(std::ostream& out) const;
 
+		// Print mean vector and covariance matrix
+		void print2(std::ostream& out) const;
+
 		// Read from stream
 		void read(std::istream& in);
 

### 2013-04-21 src/DNest3/TDModel.cpp
Need to reconstruct object

@@ -373,6 +373,9 @@ void TDModel::read(istream& in)
 	in>>logTau_qso;
 	in>>f_bad;
 	in>>boost;
+
+	formMeanVector();
+	formCovarianceMatrix();
 }
 
 string TDModel::description() const

### 2013-04-21 src/DNest3/TDModel.cpp
Need to write out bad_uniforms

@@ -344,6 +344,9 @@ void TDModel::print(ostream& out) const
 	out<<logSig_qso<<' ';
 	out<<logTau_qso<<' ';
 	out<<f_bad<<' '<<boost<<' ';
+
+	for(int i=0; i<numPoints; i++)
+		out<<bad_uniforms[i]<<' ';
 }
 
 void TDModel::print2(ostream& out) const
@@ -373,6 +376,9 @@ void TDModel::read(istream& in)
 	in>>logTau_qso;
 	in>>f_bad;
 	in>>boost;
+
+	for(int i=0; i<numPoints; i++)
+		in>>bad_uniforms[i];
 
 	formMeanVector();
 	formCovarianceMatrix();

### 2013-04-21 src/DNest3/TDModel.cpp
Return the istream

@@ -360,7 +360,7 @@ void TDModel::print2(ostream& out) const
 			out<<covarianceMatrix(i, j)<<' ';
 }
 
-void TDModel::read(istream& in)
+istream& TDModel::read(istream& in)
 {
 	for(int i=0; i<numImages; i++)
 		in>>mag[i];
@@ -382,6 +382,8 @@ void TDModel::read(istream& in)
 
 	formMeanVector();
 	formCovarianceMatrix();
+
+	return in;
 }
 
 string TDModel::description() const

### 2013-04-21 src/DNest3/TDModel.h

@@ -86,7 +86,7 @@ class TDModel:public DNest3::Model
 		void print2(std::ostream& out) const;
 
 		// Read from stream
-		void read(std::istream& in);
+		std::istream& read(std::istream& in);
 
 		// Return string with column information
 		std::string description() const;

### 2013-04-21 src/DNest3/mean_covariance.cpp
Code for posterior predictive checks

@@ -1,0 +1,38 @@
+#include <iostream>
+#include <fstream>
+#include "CommandLineOptions.h"
+#include "Data.h"
+#include "TDModel.h"
+
+using namespace std;
+using namespace DNest3;
+
+int main(int argc, char** argv)
+{
+	// Process command line options and load data
+	CommandLineOptions options(argc, argv);
+	string dataFile = options.get_dataFile();
+	Data::get_instance().load(dataFile.c_str());
+
+	TDModel t;
+
+	// Load posterior samples
+	// and save the mean and covariance to a file
+	fstream fin("posterior_sample.txt", ios::in);
+	fstream fout("mean_covariance.txt", ios::out);
+	if(!fin)
+		cerr<<"# ERROR. Couldn't load posterior_sample.txt"<<endl;
+	int k = 0;
+	while(t.read(fin))
+	{
+		t.print2(fout); fout<<endl;
+		cout<<++k<<endl;
+		if(k > 10)
+			break;
+	}
+	fout.close();
+	fin.close();
+
+	return 0;
+}
+

### 2013-04-21 src/DNest3/posterior_predictive_check.py

@@ -1,0 +1,27 @@
+import numpy as np
+import numpy.random as rng
+import matplotlib.pyplot as plt
+import scipy.linalg as la
+
+data = np.loadtxt('leon.txt')
+models = np.atleast_2d(np.loadtxt('mean_covariance.txt'))
+
+num_points = int((-1 + np.sqrt(1. + 4.*models.shape[1]))/2.)
+
+plt.ion()
+for i in xrange(0, models.shape[0]):
+	model = models[i, :]
+	mu = model[0:num_points]
+	C  = model[num_points:].reshape((num_points, num_points))
+	L = np.matrix((la.cholesky(C)).T)
+	y = mu + L*np.matrix(rng.randn(num_points, 1))
+
+	plt.hold(False)
+	plt.plot(data[:,0], data[:,1], 'b.', markersize=1)
+	plt.hold(True)
+	plt.plot(data[:,0], y, 'r.', markersize=1)
+	plt.draw()
+
+plt.ioff()
+plt.show()
+

### 2013-04-21 src/DNest3/posterior_predictive_check.py
Fixed matrix algebra

@@ -11,7 +11,7 @@ num_points = int((-1 + np.sqrt(1. + 4.*models.shape[1]))/2.)
 plt.ion()
 for i in xrange(0, models.shape[0]):
 	model = models[i, :]
-	mu = model[0:num_points]
+	mu = np.matrix(model[0:num_points]).T
 	C  = model[num_points:].reshape((num_points, num_points))
 	L = np.matrix((la.cholesky(C)).T)
 	y = mu + L*np.matrix(rng.randn(num_points, 1))

### 2013-04-21 src/DNest3/Limits.cpp
Removed printing which was only meant to be temporary

@@ -60,7 +60,6 @@ void Limits::set(const Data& data)
 	logSig_qso_min = log(logSig_qso_min);
 	logSig_qso_max = log(logSig_qso_max);
 
-	cout<<logSig_qso_min<<' '<<logSig_qso_max<<endl;
 	logSig_qso_range = logSig_qso_max - logSig_qso_min;
 
 	logTau_qso_min = log(1E-3*data.get_tRange());

### 2013-04-21 src/DNest3/posterior_predictive_check.py
Improved plotting

@@ -2,6 +2,10 @@ import numpy as np
 import numpy.random as rng
 import matplotlib.pyplot as plt
 import scipy.linalg as la
+import time
+
+plt.rc("font", size=18, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
 
 data = np.loadtxt('leon.txt')
 models = np.atleast_2d(np.loadtxt('mean_covariance.txt'))
@@ -14,13 +18,20 @@ for i in xrange(0, models.shape[0]):
 	mu = np.matrix(model[0:num_points]).T
 	C  = model[num_points:].reshape((num_points, num_points))
 	L = np.matrix((la.cholesky(C)).T)
-	y = mu + L*np.matrix(rng.randn(num_points, 1))
 
-	plt.hold(False)
-	plt.plot(data[:,0], data[:,1], 'b.', markersize=1)
-	plt.hold(True)
-	plt.plot(data[:,0], y, 'r.', markersize=1)
-	plt.draw()
+	for k in xrange(0, 3):
+		y = mu + L*np.matrix(rng.randn(num_points, 1))
+
+		plt.hold(False)
+		plt.plot(data[:,0], data[:,1], 'b.', markersize=1, label='Actual Data')
+		plt.hold(True)
+		plt.plot(data[:,0], y, 'r.', markersize=1, label='Simulated Data')
+		plt.title('{count}/{total}'.format(count=(3*i + k + 1), total=3*models.shape[0]))
+		plt.legend(numpoints=1, loc='lower left')
+		plt.xlabel('Time')
+		plt.ylabel('Magnitude')
+		plt.draw()
+		time.sleep(1)
 
 plt.ioff()
 plt.show()

### 2013-04-21 doc/Leon/ms.tex
A table

@@
 \end{eqnarray}
-where $\mu_A(t)$ and $\mu_B(t)$ are microlensing perturbations.
+where $\mu_A(t)$ and $\mu_B(t)$ are perturbations due to microlensing.
 
@@
 \end{eqnarray}
-
 Numerically, the logarithm of the determinant and the
@@
 
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Parameter & Description & Prior\\
+\hline
+$\{m_k\}$ & Mean magnitudes for each image & Uniform$(m_{\rm min}, m_{\rm max})$\\
+$\{\tau_k\}$ & Time delays for each image & Cauchy$(0, \tau_{\rm range}/10)T(\tau_{\rm min}, \tau_{\rm max})$\\
+\end{tabular}
+\end{center}
+\end{table}
+
 \end{document}

### 2013-04-22 doc/Leon/ms.tex
More work on the table

@@
 % Every Latex document starts with a documentclass command
-\documentclass[a4paper, 11pt]{article}
+\documentclass[a4paper, 10pt]{article}
 
@@
 \hline
-$\{m_k\}$ & Mean magnitudes for each image & Uniform$(m_{\rm min}, m_{\rm max})$\\
-$\{\tau_k\}$ & Time delays for each image & Cauchy$(0, \tau_{\rm range}/10)T(\tau_{\rm min}, \tau_{\rm max})$\\
+$\{m_k\}$ & Mean magnitudes & Uniform$(m_{\rm min}, m_{\rm max})$\\
+$\{\tau_k\}$ & Time delays & Cauchy$(0, \tau_{\rm range}/10)T(\tau_{\rm min}, \tau_{\rm max})$\\
+$\{\sigma_{\rm ML}\}$ & Standard deviations of microlensing signals & LogUniform$()$\\
+$\{L_{\rm ML}\}$ & Timescales of microlensing signals & LogUniform$()$\\
+$\alpha$ & Smoothness parameter for microlensing signals & Uniform$(1,2)$\\
+$\sigma_{\rm QSO}$ & Standard deviation of quasar variability & LogUniform$()$\\
+$L_{\rm QSO}$ & Timescale of quasar variability & LogUniform$()$\\
+$f_{\rm bad}$ & Fraction of data points with incorrect errorbars & Uniform$(0,1)$\\
+$K$ & Boost factor for bad errorbars & LogUniform$(1, 100)$\\
+$\{b_i\}$ & Flag for whether each errorbar is bad & Bernoulli$(f_{\rm bad})$\\
+\hline
 \end{tabular}

### 2013-04-26 src/DNest3/display.py
Nicer plots

@@ -1,15 +1,52 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
+cut = 0.5
 sample = np.atleast_2d(np.loadtxt('sample.txt'))
 sample_info = np.atleast_2d(np.loadtxt('sample_info.txt'))
+cut = int(cut*sample.shape[0])
+sample = sample[cut:]
+sample_info = sample_info[cut:]
 
-for i in xrange(0, sample.shape[1]):
-	plt.subplot(2,1,1)
-	plt.plot(sample[:,i])
-	plt.title(i)
-	plt.subplot(2,1,2)
-	plt.plot(sample_info[:,0], sample[:,i], 'k.', markersize=2)
-	plt.title('Convergence Plot')
-	plt.show()
+#for i in xrange(0, sample.shape[1]):
+#	plt.subplot(2,1,1)
+#	plt.plot(sample[:,i])
+#	plt.title(i)
+#	plt.subplot(2,1,2)
+#	plt.plot(sample_info[:,0], sample[:,i], 'k.', markersize=2)
+#	plt.title('Convergence Plot')
+#	plt.show()
+
+plt.rc("font", size=18, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
+posterior_sample = np.atleast_2d(np.loadtxt('posterior_sample.txt'))
+
+plt.hist(posterior_sample[:,3], 30, normed=True, label='Posterior Samples')
+plt.xlabel('Time Delay $\\tau$ (days)')
+plt.ylabel('Posterior Probability')
+plt.title('Time Delay = {mean:.2f} $\\pm$ {sd:.2f} days'\
+		.format(mean=posterior_sample[:,3].mean(),
+		sd=posterior_sample[:,3].std()))
+plt.savefig('hist.pdf', bbox_inches='tight')
+plt.show()
+
+import triangle
+index = np.array([11, 12, 3])
+labels = [r'$f_{\rm bad}$',
+		r'$K$', r'$\tau$']
+
+ndim, nsamples = index.size, posterior_sample.shape[0]
+triangle.corner(posterior_sample[:, index], labels=labels, quantiles=[0.16, 0.5, 0.84], bins=30, smooth=False)
+plt.show()
+
+plt.plot(posterior_sample[:,11], posterior_sample[:,12],
+		'bo', markersize=5, alpha=0.25, label='Posterior Samples')
[…]

### 2013-04-26 src/DNest3/mean_covariance.cpp

@@ -27,7 +27,7 @@ int main(int argc, char** argv)
 	{
 		t.print2(fout); fout<<endl;
 		cout<<++k<<endl;
-		if(k > 10)
+		if(k >= 16)
 			break;
 	}
 	fout.close();

### 2013-04-26 src/DNest3/posterior_predictive_check.py

@@ -12,27 +12,35 @@ models = np.atleast_2d(np.loadtxt('mean_covariance.txt'))
 
 num_points = int((-1 + np.sqrt(1. + 4.*models.shape[1]))/2.)
 
-plt.ion()
-for i in xrange(0, models.shape[0]):
+which = rng.randint(9)
+plt.figure()
+for i in xrange(0, 9):
 	model = models[i, :]
 	mu = np.matrix(model[0:num_points]).T
 	C  = model[num_points:].reshape((num_points, num_points))
 	L = np.matrix((la.cholesky(C)).T)
 
-	for k in xrange(0, 3):
-		y = mu + L*np.matrix(rng.randn(num_points, 1))
+	y = mu + L*np.matrix(rng.randn(num_points, 1))
 
-		plt.hold(False)
-		plt.plot(data[:,0], data[:,1], 'b.', markersize=1, label='Actual Data')
-		plt.hold(True)
-		plt.plot(data[:,0], y, 'r.', markersize=1, label='Simulated Data')
-		plt.title('{count}/{total}'.format(count=(3*i + k + 1), total=3*models.shape[0]))
-		plt.legend(numpoints=1, loc='lower left')
-		plt.xlabel('Time')
-		plt.ylabel('Magnitude')
-		plt.draw()
-		time.sleep(1)
+	plt.subplot(3,3,i+1)
+	if i == which:
+		plt.plot(data[:,0], data[:,1], 'b.', markersize=1)
+	else:
+		plt.plot(data[:,0], y, 'b.', markersize=1)
+	plt.xlim([data[:,0].min(), data[:,0].max()])
+	plt.gca().set_xticklabels([])
+	plt.gca().set_yticklabels([])
+#	plt.title('{count}/{total}'.format(count=(3*i + k + 1), total=3*models.shape[0]))
+#	plt.legend(numpoints=1, loc='lower left')
+#	plt.xlabel('Time')
+#	plt.ylabel('Magnitude')
 
-plt.ioff()
+plt.subplot(3,3,8)
+plt.xlabel(r'$t$ (days)')
+plt.subplot(3,3,4)
+plt.ylabel(r'$y$ (magnitudes)')
+
+plt.tight_layout()
+plt.savefig('posterior_predictive_check.pdf', bbox_inches='tight')
 plt.show()
 

### 2013-04-29 doc/Leon/ms.tex
Include beautiful figures
Finished table

--- 2013-04-21 doc/Leon/ms.tex
@@
 \hline
-$\{m_k\}$ & Mean magnitudes for each image & Uniform$(m_{\rm min}, m_{\rm max})$\\
-$\{\tau_k\}$ & Time delays for each image & Cauchy$(0, \tau_{\rm range}/10)T(\tau_{\rm min}, \tau_{\rm max})$\\
+$\{m_k\}$ & Mean magnitudes & Uniform$(m_{\rm min}, m_{\rm max})$\\
+$\{\tau_k\}$ & Time delays & Cauchy$(0, \tau_{\rm range}/10)T(\tau_{\rm min}, \tau_{\rm max})$\\
+$\{\sigma_{\rm ML}\}$ & Standard deviations of microlensing signals & LogUniform$(\sigma_{\rm min}, \sigma_{\rm max})$\\
+$\{L_{\rm ML}\}$ & Timescales of microlensing signals & LogUniform$(L_{\rm min}, L_{\rm max})$\\
+$\alpha$ & Smoothness parameter for microlensing signals & Uniform$(1,2)$\\
+$\sigma_{\rm QSO}$ & Standard deviation of quasar variability & LogUniform$(\sigma_{\rm min}, \sigma_{\rm max})$\\
+$L_{\rm QSO}$ & Timescale of quasar variability & LogUniform$(L_{\rm min}, L_{\rm max})$\\
+$f_{\rm bad}$ & Fraction of data points with incorrect errorbars & Uniform$(0,1)$\\
+$K$ & Boost factor for bad errorbars & LogUniform$(1, 100)$\\
+$\{b_i\}$ & Flag for whether each errorbar is bad & Bernoulli$(f_{\rm bad})$\\
+\hline
 \end{tabular}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{hist.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{error_bars.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{posterior_predictive_check.pdf}
+\end{center}
+\end{figure}
+
 \end{document}

### 2013-04-29 doc/Leon/ms.tex
Adding details

@@
 
+\usepackage{dsfont}
 % Load some packages
@@
 
-Suppose that the underlying QSO variability as a function of time is
-described by:
+Throughout this section we consider the measured signal in magnitudes,
+rather than fluxes. Suppose that the underlying QSO variability as a function
+of time is described by the following function:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Then the light curves of the two images $A$ and $B$ will be given by:
+Then the noise-free light curves of the two images
+$A$ and $B$ will be given by:
 \begin{eqnarray}
 y_A(t) &=& y_{\rm QSO}(t) + \mu_A(t)\\
-y_B(t) &=& y_{\rm QSO}(t - \tau) + \mu_B(t)
+y_B(t) &=& y_{\rm QSO}(t - \tau) + \mu_B(t).
 \end{eqnarray}
-where $\mu_A(t)$ and $\mu_B(t)$ are perturbations due to microlensing.
+The variability of image $A$ is the underlying QSO variability
+plus a signal $\mu_A(t)$ that is due to microlensing or perhaps
+other effects. The variability of image $B$ is given by
+a lagged version of the QSO variability, plus another
+perturbation $\mu_B(t)$ caused my microlensing.
 
-The data for all images are in magnitudes:
+The data are assumed to be noisy measurements of $y_A(t)$ and
+$y_B(t)$ at various times. We denote the entire dataset (both images together)
+as a vector of magnitude measurements:
 \begin{eqnarray}
-\yy = \{y_1, y_2, ..., y_N\}
+\yy = \{y_1, y_2, ..., y_N\}.
 \end{eqnarray}
 
-The probability distribution for the data $\yy$ given the parameters
+If we make the simplifying assumption that our prior knowledge about
+$\mu_A(t)$, $\mu_B(t)$, and
+$y_{\rm QSO}(t)$ can be described by Gaussian Processes \citep{rasmussen},
+then the probability distribution for the data $\yy$ given the parameters
 $\theta$
@@
 \end{eqnarray}
+The elements of the mean vector are just the mean magnitudes of the
+appropriate images:
+\begin{eqnarray}
+m_i = 
+\end{eqnarray}
+
+The covariance matrix elements are given by:
+\begin{eqnarray}
+C_{ij} &=& \textnormal{Cov}(y_i, y_j) \\
+&=& \mathds{E}\left[
+\left(y_i - m_i\right)
+\left(y_j - m_j\right)
+\right]
+\end{eqnarray}
+
 Numerically, the logarithm of the determinant and the
@@
 
+\begin{thebibliography}{99} 
+\bibitem[\protect\citeauthoryear{Rasmussen \& Williams}{2006}]{rasmussen} Rasmussen C., Williams C., Gaussian Processes for Machine Learning, The MIT Press, 2006
+\end{thebibliography}
+
 \end{document}

### 2013-05-03 doc/Leon/ms.tex
Covariance functions
Working on the document

@@
 \begin{document}
+In this section we describe the Bayesian model used to infer the
+time delay from the data. This model will be described in greater detail in
+a forthcoming contribution (Brewer and Marshall, in prep).
+Throughout this section we consider the measured signal in magnitudes,
+rather than fluxes, as the model assumptions (Section~\ref{sec:assumptions})
+make more sense in terms of magnitudes.
 
-Throughout this section we consider the measured signal in magnitudes,
-rather than fluxes. Suppose that the underlying QSO variability as a function
-of time is described by the following function:
+\section{The Data}
+We denote the entire dataset (both images together)
+as a vector of magnitude measurements:
+\begin{eqnarray}
+\yy = \{y_1, y_2, ..., y_N\}.
+\end{eqnarray}
+Along with these measurements we have the timestamps $\{t_i\}_{i=1}^N$,
+error bars $\{\sigma_i\}_{i=1}^N$ and flags
+$\{c_i\}_{i=1}^N$ where a flag $c_i \in \{A, B\}$ tells us whether the
+measurement is of image $A$ or image $B$. The likelihood will be
+written as a probability distribution for $\yy$ given some parameters, and the
+timestamps, errorbars and flags will be considered as part of the prior
+information.
+
+\section{Model Assumptions}\label{sec:assumptions}
+Suppose that the underlying QSO variability as a function
+of time, as seen through image $A$, is described by the following function:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Then the noise-free light curves of the two images
+Then the light curves of the two images
 $A$ and $B$ will be given by:
 \begin{eqnarray}
-y_A(t) &=& y_{\rm QSO}(t) + \mu_A(t)\\
-y_B(t) &=& y_{\rm QSO}(t - \tau) + \mu_B(t).
+y_A(t) &=& M_A + y_{\rm QSO}(t)\\
+y_B(t) &=& M_B + y_{\rm QSO}(t - \tau).
 \end{eqnarray}
-The variability of image $A$ is the underlying QSO variability
+The light curve of image $A$ is just the underlying QSO variability,
+and the light curve of
+
 plus a signal $\mu_A(t)$ that is due to microlensing or perhaps
@@
 The data are assumed to be noisy measurements of $y_A(t)$ and
-$y_B(t)$ at various times. We denote the entire dataset (both images together)
-as a vector of magnitude measurements:
-\begin{eqnarray}
-\yy = \{y_1, y_2, ..., y_N\}.
-\end{eqnarray}
+$y_B(t)$ at various times.
 
@@
 \begin{eqnarray}
-m_i = 
+m_i = M_{c_i}
 \end{eqnarray}
@@
 \right]
+\end{eqnarray}
+
+We assume the following:
+\begin{eqnarray}
+\textnormal{Cov}\left(y_{\rm QSO}(t_1), y_{\rm QSO}(t_2)\right)
+&=& \sigma_{\rm QSO}^2\exp\left[-\frac{\left|t_2 - t_1\right|}{L_{\rm QSO}}\right]
+\end{eqnarray}
+
+\begin{eqnarray}
+\textnormal{Cov}\left(\mu_A(t_1), \mu_A(t_2)\right)
+&=& \sigma_{{\rm ML}, A}^2\exp\left[-\left(\frac{\left|t_2 - t_1\right|}{L_{{\rm ML}, A}}\right)^\alpha\right]
+\end{eqnarray}
+
+\begin{eqnarray}
+\textnormal{Cov}\left(\mu_B(t_1), \mu_B(t_2)\right)
+&=& \sigma_{{\rm ML}, B}^2\exp\left[-\left(\frac{\left|t_2 - t_1\right|}{L_{{\rm ML}, B}}\right)^\alpha\right]
 \end{eqnarray}

### 2013-05-03 doc/Leon/ms.tex
Much more writing

@@
 \begin{eqnarray}
-y_A(t) &=& M_A + y_{\rm QSO}(t)\\
-y_B(t) &=& M_B + y_{\rm QSO}(t - \tau).
+y_A(t) &=& M_A + y_{\rm QSO}(t) + \mu_A(t)\\
+y_B(t) &=& M_B + y_{\rm QSO}(t - \tau) + \mu_B(t).
 \end{eqnarray}
 The light curve of image $A$ is just the underlying QSO variability,
-and the light curve of
+fluctuating around the mean magnitude $M_A$ of image $A$, plus
+a microlensing signal $\mu_A(t)$. Similarly,
+the light curve of image $B$ is its mean magnitude $M_B$ plus the QSO
+variability delayed by a lag $\tau$, plus a different microlensing
+signal $\mu_B(t)$. The data are assumed to be noisy measurements of $y_A(t)$ and
+$y_B(t)$ at various times.
 
-plus a signal $\mu_A(t)$ that is due to microlensing or perhaps
-other effects. The variability of image $B$ is given by
-a lagged version of the QSO variability, plus another
-perturbation $\mu_B(t)$ caused my microlensing.
+\subsection{Error Bar Wrongness}
+We do not assume that the given error bars
+$\{\sigma_i\}_{i=1}^N$ are entirely correct. Instead, we assume that each
+measurement is either ``good'', in which case $\sigma_i$ is used, otherwise it
+is ``bad'', in which case $K\sigma_i$ is used, where $K > 1$. Letting $b_i = 1$
+denote a bad measurement and $b_i=0$ denote a good measurement, the PDF for the
+data can be written as:
+\begin{eqnarray}
+y_i \sim \mathcal{N}
+\left(
+y_{c_i}(t_i), \sigma_i'^2
+\right)\label{eq:noise}
+\end{eqnarray}
+where
+\begin{eqnarray}
+\sigma_i' &=& \left\{
+\begin{array}{lrl}
+\sigma_i, & b_i = 1\\
+K\sigma_i, & b_i = 0 & .
+\end{array}
+\right.
+\end{eqnarray}
 
-The data are assumed to be noisy measurements of $y_A(t)$ and
-$y_B(t)$ at various times.
+\section{Unknown Parameters}
+In Table~\ref{tab:params} we list the unknown parameters that are to be inferred
+and the prior distributions that were used.
+
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Parameter & Description & Prior\\
+\hline
+$\{M_A, M_B\}$ & Mean magnitudes of the two images & Uniform$(M_{\rm min}, M_{\rm max})$\\
+$\tau$ & Time delay & Cauchy$(0, \tau_{\rm range}/10)T(\tau_{\rm min}, \tau_{\rm max})$\\
+$\{\sigma_{{\rm ML}, A}, \sigma_{{\rm ML}, B}\}$ & Standard deviations of microlensing signals & LogUniform$(\sigma_{\rm min}, \sigma_{\rm max})$\\
+$\{L_{{\rm ML, A}}, L_{{\rm ML, B}}\}$ & Timescales of microlensing signals & LogUniform$(L_{\rm min}, L_{\rm max})$\\
+$\alpha$ & Smoothness parameter for microlensing signals & Uniform$(1,2)$\\
+$\sigma_{\rm QSO}$ & Standard deviation of quasar variability & LogUniform$(\sigma_{\rm min}, \sigma_{\rm max})$\\
+$L_{\rm QSO}$ & Timescale of quasar variability & LogUniform$(L_{\rm min}, L_{\rm max})$\\
+$f_{\rm bad}$ & Fraction of data points with incorrect errorbars & Uniform$(0,1)$\\
+$K$ & Boost factor for bad errorbars & LogUniform$(1, 100)$\\
+$\{b_i\}_{i=1}^N$ & Flag for whether each errorbar is bad & Bernoulli$(f_{\rm bad})$\\
+\hline
+\end{tabular}
+\end{center}
+\caption{A list of the model parameters and their prior probability
+distributions. All of the priors are vague.\label{tab:params}}
+\end{table}
+
+\section{Gaussian Processes}
+Inspection of Equation~\ref{eq:noise} suggests that this is a simple curve
+fitting problem. However, this is not the case, because the noise-free
+model predictions $y_{c_i}(t_i)$ cannot be computed as a function of the
+parameters listed in Table~\ref{tab:params}.
 
@@
 \right)
-\right]
+\right]\label{eq:likelihood}
 \end{eqnarray}
+The Gaussian Process assumption is questionable in the case of the microlensing
+signals $\mu_A(t)$ and $\mu_B(t)$ (especially if the source is not large),
+although it is still a flexible model for any temporally-correlated signal.
+The use of a Gaussian Process model for the QSO variability is more
+well established \citep{kelly, xu}.
+
 The elements of the mean vector are just the mean magnitudes of the
@@
 \end{eqnarray}
-
 The covariance matrix elements are given by:
@@
 \end{eqnarray}
+and are obtained by evaluating the following covariance {\it functions}
+at the relevant time differences.
+\begin{eqnarray}
+\textnormal{Cov}\left(y_{\rm QSO}(t_i), y_{\rm QSO}(t_j)\right)
+&=& \sigma_{\rm QSO}^2\exp\left[-\frac{\left|t_i - t_j\right|}{L_{\rm QSO}}\right]\\
+\textnormal{Cov}\left(\mu_A(t_i), \mu_A(t_j)\right)
+&=& \sigma_{{\rm ML}, A}^2\exp\left[-\left(\frac{\left|t_i - t_j\right|}{L_{{\rm ML}, A}}\right)^\alpha\right]\\
+\textnormal{Cov}\left(\mu_B(t_i), \mu_B(t_j)\right)
+&=& \sigma_{{\rm ML}, B}^2\exp\left[-\left(\frac{\left|t_i - t_j\right|}{L_{{\rm ML}, B}}\right)^\alpha\right]
+\end{eqnarray}
+Numerically, the log of the likelihood function (Equation~\ref{eq:likelihood})
+requires the log of $\det(\mathbf{C})$ and the
+$\CC^{-1}\left(\yy - \mm\right)$ term. These can be implemented using
+Cholesky decompositions. Unfortunately the evaluation of this likelihood
+function takes $O(N^3)$ time, but is tractable for the B1600 data where $N=604$.
 
-We assume the following:
-\begin{eqnarray}
-\textnormal{Cov}\left(y_{\rm QSO}(t_1), y_{\rm QSO}(t_2)\right)
-&=& \sigma_{\rm QSO}^2\exp\left[-\frac{\left|t_2 - t_1\right|}{L_{\rm QSO}}\right]
-\end{eqnarray}
-
-\begin{eqnarray}
-\textnormal{Cov}\left(\mu_A(t_1), \mu_A(t_2)\right)
-&=& \sigma_{{\rm ML}, A}^2\exp\left[-\left(\frac{\left|t_2 - t_1\right|}{L_{{\rm ML}, A}}\right)^\alpha\right]
-\end{eqnarray}
-
-\begin{eqnarray}
-\textnormal{Cov}\left(\mu_B(t_1), \mu_B(t_2)\right)
-&=& \sigma_{{\rm ML}, B}^2\exp\left[-\left(\frac{\left|t_2 - t_1\right|}{L_{{\rm ML}, B}}\right)^\alpha\right]
-\end{eqnarray}
-
-Numerically, the logarithm of the determinant and the
-$\CC^{-1}\left(\yy - \mm\right)$ term are implemented using
-Cholesky decompositions.
-
-\begin{table}[h!]
-\begin{center}
-\begin{tabular}{|c|c|c|}
-\hline
-Parameter & Description & Prior\\
-\hline
-$\{m_k\}$ & Mean magnitudes & Uniform$(m_{\rm min}, m_{\rm max})$\\
-$\{\tau_k\}$ & Time delays & Cauchy$(0, \tau_{\rm range}/10)T(\tau_{\rm min}, \tau_{\rm max})$\\
-$\{\sigma_{\rm ML}\}$ & Standard deviations of microlensing signals & LogUniform$(\sigma_{\rm min}, \sigma_{\rm max})$\\
-$\{L_{\rm ML}\}$ & Timescales of microlensing signals & LogUniform$(L_{\rm min}, L_{\rm max})$\\
-$\alpha$ & Smoothness parameter for microlensing signals & Uniform$(1,2)$\\
-$\sigma_{\rm QSO}$ & Standard deviation of quasar variability & LogUniform$(\sigma_{\rm min}, \sigma_{\rm max})$\\
-$L_{\rm QSO}$ & Timescale of quasar variability & LogUniform$(L_{\rm min}, L_{\rm max})$\\
-$f_{\rm bad}$ & Fraction of data points with incorrect errorbars & Uniform$(0,1)$\\
-$K$ & Boost factor for bad errorbars & LogUniform$(1, 100)$\\
-$\{b_i\}$ & Flag for whether each errorbar is bad & Bernoulli$(f_{\rm bad})$\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+\section{Results}
+The MCMC was run using the Diffusive Nested Sampling algorithm
+\citep[][software available at {\tt http://github.com/eggplantbren/DNest3}]{dnest}.
+The posterior PDF for the time delay is shown in Figure~\ref{fig:time_delay}.
 
@@
 \includegraphics[scale=0.5]{hist.pdf}
+\caption{The posterior PDF for the time delay.\label{fig:time_delay}}
 \end{center}
@@
 \includegraphics[scale=0.5]{error_bars.pdf}
+\caption{The joint posterior PDF for the fraction of ``bad'' error bars
+and the error bar scaling parameter $K$. The dotted lines correspond to the
+assumption that the reported error bars are correct.\label{fig:time_delay}}
 \end{center}
 \end{figure}
+
+\section{Posterior Predictive Check}
+A useful sanity check for a Bayesian model is to verify that the observed data
+is typical of the kind of data that the model predicts. This is formalised
+using ``posterior predictive checks'' \citep{gelman}. Figure~\ref{fig:ppc}
+shows the actual data along with eight simulated datasets.
 
@@
 \includegraphics[scale=0.7]{posterior_predictive_check.pdf}
+\caption{Eight data sets simulated from the model
+using parameters drawn from the posterior distribution. The actual data set
+is also plotted and appears in the lower left panel. The actual data appears
+typical of the distribution of simulated data. The most atypical feature of
+the real data is the rapid variability in the final season, although similar
+features can be seen in the first two simulated datasets.
+\label{fig:ppc}}
 \end{center}
@@
 \begin{thebibliography}{99}
+
+\bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,
+\& Cs{\'a}nyi}{2011}]{dnest} Brewer B.~J., P{\'a}rtay L.~B.,
+Cs{\'a}nyi G., 2011, Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380
+
+\bibitem[\protect\citeauthoryear{Gelman, Meng, \& Stern}{1996}]{gelman} Gelman, A.,
+Meng X., Stern, H., Posterior predictive assessment of model fitness via realized discrepancies,
+Statistica Sinica 6 (1996): 733-759.
+
+\bibitem[Kelly et al.(2009)]{kelly} Kelly, B.~C., Bechtold, 
+J., Siemiginowska, A.\ 2009.\ Are the Variations in Quasar Optical Flux 
+Driven by Thermal Fluctuations?.\ The Astrophysical Journal 698, 895-910.
+
 \bibitem[\protect\citeauthoryear{Rasmussen \& Williams}{2006}]{rasmussen} Rasmussen C., Williams C., Gaussian Processes for Machine Learning, The MIT Press, 2006
+
+\bibitem[Zu et al.(2013)]{zu} Zu, Y., Kochanek, C.~S., 
+Koz{\l}owski, S., Udalski, A.\ 2013.\ Is Quasar Optical Variability a 
+Damped Random Walk?.\ The Astrophysical Journal 765, 106. 
 \end{thebibliography}

### 2013-08-14 doc/Paper1/ms.tex
New title
Working through the intro

--- 2012-05-30 doc/ms.tex
@@
 
-\title[Time Delays with Gaussian Processes]
-{Inference of Gravitational Lens Time Delays Using Gaussian Processes} 
-\author[B. J. Brewer]{Brendon J. Brewer$^1$, Philip J. Marshall$^2$, Gregory Dobler$^3$ \\
-$^1$Department of Physics, University of California, Santa Barbara, CA, USA \\
-$^2$Department of Physics, University of Oxford, Keble Road, Oxford, OX1 3RH, UK \\
-$^3$KITP}
+\title[Measuring Time Delays]
+{A Model for Extracting Time Delays from Noisy Time Series Data} 
+\author[B. J. Brewer]{Brendon J. Brewer$^1$, Philip J. Marshall$^2$\\
+$^1$Department of Statistics, The University of Auckland, Private Bag 92019,
+Auckland 1142, New Zealand\\
+$^2$Department of Physics, University of Oxford, Keble Road, Oxford, OX1 3RH, UK}
 
@@
 \begin{abstract}
-
 Observations of gravitationally lensed quasars provide a unique method for
@@
 A key part of
-this process is the extraction of the time delays from the photometric
-monitoring data of the multiple images of the quasar. Quantifying the
-uncertainty in this process is challenging, particularly when the time delays
-are obscured by the presence of microlensing in the light curves. We present a
+this process is the measurement of the time delays from the photometric
+monitoring of the multiple images of the quasar. The uncertainties in the time
+delays propagate to the uncertainties about the cosmological parameters, so it
+is important to have an accurate understanding of the size of these uncertainties.
+Quantifying the
+uncertainty in the time delay measurements is challenging, particularly when the time series
+data is affected by microlensing and other complex physical effects. We present a
 fully probabilistic model for the process of inferring the time delays from
 the light curve data. In this model, the probability distribution over light
-curves is described by a Gaussian Process (GP) whose covariance function
+curves is described by a Gaussian Process (GP) probability distribution whose covariance function
 depends on the time delays as well as the properties of the microlensing.
@@
 the time delays, while allowing for a wide range of intrinsic, microlensing, and 
-noise-related variability. We test this method extensively on simulated data, 
+noise-related variability. We test this method on a simple simulated data set
+and a challenging one, 
 and find the inferences (including parameter uncertainties) to be robust.
 We apply our method to radio and optical monitoring data of the well
-studied lenses B1608+656 and RXJ1131$-$123, deriving new time delay estimates complete 
-with covariances. The effect on the $H_0$ precision from B1608 is ...
-
+studied gravitational lens B1608+656, deriving new time delay estimates and uncertainties. The effect on the $H_0$ precision from B1608 is {\bf no idea yet}.
 \end{abstract}
@@
 
-However, any inference of $H_0$ from the gravitational lensing method inherits
+However, any inference of $H_0$ or other cosmological parameters
+from the gravitational lensing technique inherits
 the uncertainty from the measurements of the time delays themselves. A
@@
 \citep{2004kats.book.....O} so the results obtained are probability
-distributions for the time delays, and the uncertainties in the time delays
-are indicated by the widths of these probability distributions. To describe
+distributions over the space of possible solutions for the time delays, and the uncertainties in the time delays
+can be summarised by the widths of these probability distributions. To describe
 the underlying time variation of the quasar and the effects of microlensing on
@@
 is fully Bayesian and allows for exploration of the time delay parameter
-space, rather than providing a point estimate. Secondly, it is the likelihood
+space, rather than providing a point estimate of the time delays.
+Secondly, it is the likelihood
 function, rather than the scatter, which guides the inference, making it clear
-how uncertainties are to be obtained.
+how uncertainties are to be obtained. A recent contribution by \citep{2013PhRvD..87l3512H}
+described a similar approach to the one presented here. The main differences
+are that we implement the model in a Bayesian framework and obtain posterior
+samples using MCMC methods, and we also implement several robustness features
+(e.g. we allow the possibility of systematically underestimated error bars)
+that are not present in the approach of \citet{2013PhRvD..87l3512H}.
 
 An introduction to Gaussian Processes and their applications can be found in
-\citet{rasmussen}. however, in their applications, GPs are usually used as a
-prior distribution over the space of functions. Here, GPs describe the
+\citet{rasmussen}. In their applications, GPs are usually used as a choice of
+prior distribution over the space of functions. In this work, GPs describe the
 sampling distributions, or the probability of the data given the model
 parameters. This is a common situation in astrophysics in fields where the
-signals can be modelled by a stochastic process, for example quasar
-variability \citep{2009ApJ...698..895K, 2010ApJ...721.1014M, 2011ApJ...735...80Z} and solar-like
-oscillations of stars \citep{2009MNRAS.395.2226B}.
+signals are complex and the best models we can think of are 
+stochastic processes. Two prominent examples are quasar
+variability \citep{2009ApJ...698..895K, 2010ApJ...721.1014M, 2011ApJ...735...80Z} and solar-like oscillations of stars \citep{2009MNRAS.395.2226B}.
 
@@
 \section{The Model for the Inference}
-
 We begin by defining our model for the prior knowledge about how the time
 delays give rise rise to the observed data. This results in a probability
-distribution for the data given some unknown parameters, i.e. sampling
-distributions $p(D|\theta)$. Once specific data is in hand, the sampling
+distribution for the data given some unknown parameters, i.e. the ``sampling
+distribution'' (actually distributions, one for each value of the parameters)
+$p(D|\theta)$. Once specific data is in hand, the sampling
 distributions become a likelihood function for the parameters which can be
-combined with a prior distribution $p(\theta)$ to give inferences. By Bayes'
-rule, the posterior distribution for some parameters $\theta$ given data $D$
-is given by:
+combined with a prior distribution $p(\theta)$ to give the posterior distribution.
+The posterior distribution for some parameters $\theta$ given data $D$
+is given by Bayes' rule:
 \begin{equation}
@@
 The intrinsic quasar variability, in magnitudes, and in whatever band the
-observations are carried out in, is described by a function $f(t)$. For a
-four-image system, the resulting light curves, if observed without gaps and
+observations are carried out in, can be described by a function $q(t)$. For a
+two-image system, the resulting light curves, if observed without gaps and
 noise, would be described by the functions
@@
 y_2(t) &=& m_2 + f(t - \tau_2) + \mu_2(t) \\
-y_3(t) &=& m_3 + f(t - \tau_3) + \mu_3(t) \\
-y_4(t) &=& m_4 + f(t - \tau_4) + \mu_4(t) 
 \end{eqnarray}
-where $\{m_1, m_2, m_3, m_4\}$ are the mean magnitudes of the images,
-$\{\tau_2, \tau_3, \tau_4\}$ are the time delays relative to image 1, and the
-functions $\{\mu_1(t), \mu_2(t), \mu_3(t), \mu_4(t)\}$ describe the perturbing
-effects of microlensing on each of the light curves. The data then consist of
-noisy, irregularly sampled measurements of the light curves $\{y_1(t), y_2(t),
-y_3(t), y_4(t)\}$.
+where $\{m_1, m_2\}$ are the mean magnitudes of the images,
+$\tau_2$ is the time delay of image 2 relative to image 1, and the
+functions $\{\mu_1(t), \mu_2(t)\}$ describe any perturbations that are not
+shared by both images. A major source of perturbations is the effect of
+microlensing on each of the light curves. The data then consist of
+noisy, irregularly sampled measurements of the light curves $\{y_1(t), y_2(t)\}$.
+
+The actual data we observe is as follows. We will combine the measurements of all
+images into a single vector of $N$ magnitude measurements.
+At a set of times
+$\{t_1, t_2, ..., t_N\}$, we measure the magnitude of one of the images. These
+noisy measurements are denoted $\{Y_1, Y_2, ..., Y_N\}$. We also keep track of
+a set of flags $\{k_1, k_2, ..., k_N\}, \forall k_i \in \{1, 2\}$ which specifies
+which image was measured. For example, if $k_4 = 2$ then we know that $Y_4$ was
+a noisy measurement of image two at time $t_4$.
+
+When we use Bayes' rule, the flags $\{k_1, k_2, ..., k_N\}$ and measurement
+times $\{t_1, t_2, ..., t_N\}$ will be subsumed into the prior information. The
+data will be the measurements
+\begin{eqnarray}
+\boldsymbol{D} = \{Y_1, Y_2, ..., Y_N\}.
+\end{eqnarray}
+
+The sampling distributions (and hence the likelihood function) will then be
+written as a probability distribution for the measurements $\boldsymbol{D}$
+given the time delay $\tau$ along with some nuisance parameters.
 
@@
 
+\bibitem[\protect\citeauthoryear{Hojjati, Kim, 
+\& Linder}{2013}]{2013PhRvD..87l3512H} Hojjati A., Kim A.~G., Linder E.~V., 2013, PhRvD, 87, 123512
+
 \bibitem[\protect\citeauthoryear{Kelly, Bechtold, 

### 2013-08-14 doc/Paper1/ms.tex
Phil's affiliation

@@
 Auckland 1142, New Zealand\\
-$^2$Department of Physics, University of Oxford, Keble Road, Oxford, OX1 3RH, UK}
+$^2$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA}
 

### 2013-08-14 doc/Paper1/ms.tex
Slight fix

@@
 written as a probability distribution for the measurements $\boldsymbol{D}$
-given the time delay $\tau$ along with some nuisance parameters.
+given the time delay $\tau_2$ along with some nuisance parameters.
+
+\section{Sampling Distribution}
 

### 2013-08-14 doc/Paper1/ms.tex
SOme edits

--- 2013-08-14 doc/Paper1/ms.tex
@@
 \usepackage{amsmath}
+
+\newcommand{\params}{\boldsymbol{\theta}}
+\newcommand{\data}{\boldsymbol{D}}
 
@@
 
-\section{The Model for the Inference}
+\section{Bayesian Inference}
+For inferring unknown parameters $\params$ from data $\data$,
+Bayes' rule gives the posterior distribution
+\begin{eqnarray}
+p(\params | \data) \propto p(\params)p(\data|\params)
+\end{eqnarray}
+The prior distribution $p(\params)$ describes prior knowledge or uncertainty
+about the values of the parameter(s). The ``sampling distributions''
+$p(\data|\params)$
+describes prior knowledge about the data we expect to observe, as a function of
+$\params$ (Caticha). When $\data$ is fixed $p(\data|\params)$ is a function of
+$\params$ only and is called the likelihood function.
+
 We begin by defining our model for the prior knowledge about how the time
-delays give rise rise to the observed data. This results in a probability
-distribution for the data given some unknown parameters, i.e. the ``sampling
-distribution'' (actually distributions, one for each value of the parameters)
-$p(D|\theta)$. Once specific data is in hand, the sampling
-distributions become a likelihood function for the parameters which can be
-combined with a prior distribution $p(\theta)$ to give the posterior distribution.
-The posterior distribution for some parameters $\theta$ given data $D$
-is given by Bayes' rule:
-\begin{equation}
-p(\theta|D = D^*) \propto p(\theta)p(D|\theta)|_{D = D^*}
-\end{equation}
-where $D^*$ is the actual data set that was observed. Throughout this paper,
-the data $D$ are a set of measured magnitudes of multiple images of a QSO; see
-Figure~\ref{simdata} for an example data set.
+delays give rise rise to the observed data. This will determine the sampling
+distribution, and also the set of parameters (including nuisance parameters) to
+be inferred.
 
+\section{Sampling Distributions}
 The intrinsic quasar variability, in magnitudes, and in whatever band the
@@
 The sampling distributions (and hence the likelihood function) will then be
-written as a probability distribution for the measurements $\boldsymbol{D}$
-given the time delay $\tau$ along with some nuisance parameters.
+written as a probability distribution for the measurements $\data$
+given the time delay $\tau_2$ along with some nuisance parameters.
 
@@
 microlensing contributions.
+
+{RANT ABOUT HOW THE GP ASSUMPTION FOR "MICROLENSING" IS A DESCRIPTION OF OUR
+PRIOR UNCERTAINTY ABOUT THE SIGNAL **IN OUR PARTICULAR DATA SET**}
 

### 2013-08-16 doc/Paper1/ms.tex
Table of variables

@@
 \newcommand{\data}{\boldsymbol{D}}
+\newcommand{\info}{\boldsymbol{I}}
 
@@
 \section{Bayesian Inference}
-For inferring unknown parameters $\params$ from data $\data$,
+For inferring unknown parameters $\params$ from data $\data$ in the context
+of prior information $\info$,
 Bayes' rule gives the posterior distribution
 \begin{eqnarray}
-p(\params | \data) \propto p(\params)p(\data|\params)
+p(\params | \data, \info) \propto p(\params|\info)p(\data|\params, \info)
 \end{eqnarray}
-The prior distribution $p(\params)$ describes prior knowledge or uncertainty
+The prior distribution $p(\params|\info)$ describes prior knowledge or uncertainty
 about the values of the parameter(s). The ``sampling distributions''
-$p(\data|\params)$
+$p(\data|\params,\info)$
 describes prior knowledge about the data we expect to observe, as a function of
-$\params$ (Caticha). When $\data$ is fixed $p(\data|\params)$ is a function of
+$\params$ (Caticha). When $\data$ is fixed $p(\data|\params,\info)$ is a function of
 $\params$ only and is called the likelihood function.
@@
 given the time delay $\tau_2$ along with some nuisance parameters.
+
+\begin{table*}
+\begin{tabular}{lll}
+\hline
+{\bf Parameter} &	{\bf Meaning}			& {\bf Prior}\\
+\hline
+$\tau_2$	&	Time delay relative to image 1	& Cauchy()$T(,)$\\
+$\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform()\\
+\hline
+{\bf Data}	&	{\bf Meaning}			& {\bf Sampling Distribution}\\
+\hline
+$\{Y_1, ..., Y_N\}$ &	Noisy magnitude measurements	& $\mathcal{N}\left(\mathbf{m}(\params), \mathbf{C}(\params)\right)$\\
+\hline
+{\bf Prior Information} & {\bf Meaning}			&\\
+\hline
+$N$		&	Number of measurements (all images together)\\
+$\{t_1, ..., t_N\}$ &	Measurement times\\
+$\{k_1, ..., k_N\}$ &	Flags identifying which image was observed
+\end{tabular}
+\end{table*}
 

### 2013-08-20 doc/Paper1/ms.tex
ell is a good letter

@@
 $\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform()\\
+$L$		&	Length scale of QSO variations	& LogUniform()\\
+$\{\ell_1, \ell_2\}$ &	Length scale of image-specific variations & \\
 \hline

### 2013-08-20 src/DNest3/Limits.cpp
Use sig/sqrt(tau), not sig

@@ -50,15 +50,8 @@ void Limits::set(const Data& data)
 	alpha_max = 2.;
 	alpha_range = alpha_max - alpha_min;
 
-	logSig_qso_min = 0.;
-	logSig_qso_max = 0.;
-	for(int i=0; i<data.get_numImages(); i++)
-	{
-		logSig_qso_min += 1E-3*data.get_yStDev()[i]/data.get_numImages();
-		logSig_qso_max += 1E+3*data.get_yStDev()[i]/data.get_numImages();
-	}
-	logSig_qso_min = log(logSig_qso_min);
-	logSig_qso_max = log(logSig_qso_max);
+	logSig_qso_min = log(1E-4);
+	logSig_qso_max = log(1E2);
 
 	logSig_qso_range = logSig_qso_max - logSig_qso_min;
 

### 2013-08-20 src/DNest3/TDModel.cpp

@@ -236,7 +236,6 @@ double TDModel::perturb()
 	if(change[8])
 		logH += perturb9();
 
-
 	formMeanVector();
 	formCovarianceMatrix();
 	return logH;
@@ -288,7 +287,7 @@ double TDModel::covariance(double t1, double t2, int ID1, int ID2)
 	double tau_qso = exp(logTau_qso);
 
 	double exponent = abs((t1 - tau[ID1]) - (t2 - tau[ID2]))/tau_qso;
-	double C = pow(sig_qso, 2)
+	double C = pow(sig_qso*sqrt(tau_qso), 2)
 				*exp(-exponent);
 
 	vector<double> sig_ml(numImages);

### 2013-08-20 src/DNest3/TDModel.h

@@ -33,7 +33,7 @@ class TDModel:public DNest3::Model
 		// Microlensing smoothness
 		double alpha;
 
-		// QSO Variability amplitude
+		// QSO Variability amplitude IN UNITS OF SQRT(TAU)
 		double logSig_qso;
 
 		// QSO variability timescale

### 2013-08-23 doc/Paper1/ms.tex
Filling in parameter table

--- 2013-08-16 doc/Paper1/ms.tex
@@
 \usepackage{amsmath}
+\usepackage{dsfont}
 
@@
 
+Part of our model involves Gaussian Processes, which are a family of probability
+distributions over the space of possible functions.
 An introduction to Gaussian Processes and their applications can be found in
@@
 stochastic processes. Two prominent examples are quasar
-variability \citep{2009ApJ...698..895K, 2010ApJ...721.1014M, 2011ApJ...735...80Z} and solar-like oscillations of stars \citep{2009MNRAS.395.2226B}.
+variability \citep{2009ApJ...698..895K, 2010ApJ...721.1014M, 2011ApJ...735...80Z}
+and solar-like oscillations of stars \citep{2009MNRAS.395.2226B}.
+
+In Section~\ref{sec:bayes} we briefly discuss Bayesian Inference, and in
+Section~\ref{sec:sampling} we describe our specific
+modelling assumptions. The computational implementation is described in
+Section~\ref{sec:computation}. In Section~\ref{sec:tests} we test the inference
+with two simulated data sets, one ``easy'' data set and one ``hard'' data set.
+In Section~\ref{sec:b1608} we apply the model to B1608+656.
 
@@
 
-\section{Bayesian Inference}
+\section{Bayesian Inference}\label{sec:bayes}
 For inferring unknown parameters $\params$ from data $\data$ in the context
@@
 $p(\data|\params,\info)$
-describes prior knowledge about the data we expect to observe, as a function of
-$\params$ (Caticha). When $\data$ is fixed $p(\data|\params,\info)$ is a function of
+describe prior knowledge about the data we expect to observe, as a function of
+$\params$ \citep{caticha}. When $\data$ is fixed $p(\data|\params,\info)$ is a function of
 $\params$ only and is called the likelihood function.
@@
 delays give rise rise to the observed data. This will determine the sampling
-distribution, and also the set of parameters (including nuisance parameters) to
-be inferred.
+distributions, and also the set of parameters (including nuisance parameters) to
+be inferred. Our choice of prior distribution for the parameters will be
+discussed in Section
 
-\section{Sampling Distributions}
+\section{Sampling Distributions}\label{sec:sampling}
 The intrinsic quasar variability, in magnitudes, and in whatever band the
@@
 microlensing on each of the light curves. The data then consist of
-noisy, irregularly sampled measurements of the light curves $\{y_1(t), y_2(t)\}$.
+noisy, irregularly sampled measurements of the light curves
+$\{y_1(t), y_2(t)\}$.
 
@@
 \hline
-$\tau_2$	&	Time delay relative to image 1	& Cauchy()$T(,)$\\
-$\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform()\\
+$\tau_2$	&	Time delay relative to image 1	& Cauchy$\left(0, 0.1\tau_{\rm range}\right)T\left(\tau_{\rm min},\tau_{\rm max}\right)$\\
+$\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform$(m_{\rm min}, m_{\rm max})$\\
+$L$		&	Length scale of QSO variations	& LogUniform$(L_{\rm min}, L_{\rm max})$\\
+$\beta$		&	Size of QSO variations		& LogUniform$(\beta_{\rm min}, \beta_{\rm max})$\\
+$A$		&	Size of image-specific variations	& LogUniform$\left(A_{\rm min}, A_{\rm max}\right)$\\
+$\{\ell_1, \ell_2\}$ &	Length scale of image-specific variations & LogUniform$(\ell_{\rm min}, \ell_{\rm max})$\\
+$\alpha$	&	Smoothness of image-specific variations	& Uniform$(1, 2)$\\
+$f_{\rm bad}$	&	Fraction of ``bad'' data points	& Uniform$(0,1)$\\
+$\kappa_1$	&	Error-bar boost for ``good'' data & $\frac{1}{2}\delta(\kappa_1 - 1) + \frac{1}{2}$LogUniform$(1,100)$\\
+$\kappa_2$	&	Error-bar boost for ``bad'' data & LogUniform$(\kappa_1, 100\kappa_1)$\\
+$\{b_i\}_{i=1}^N$ &	Badness hidden variables for each data point & Uniform$(0,1)$\\
+\hline
+{\bf Derived Parameter} & {\bf Meaning}	&	{\bf Definition}\\
+\hline
+$\{B_i\}_{i=1}^N$	&	Badness flag for each data point & $\mathds{1}\left(b_i \leq f_{\rm bad}\right)$\\
 \hline
@@
 between $a$ and $b$. $\Delta t$ is the duration of the data set. All priors
-are independent.\label{priors}}
+are independent.\label{tab:prior}}
 \end{table}
@@
 
+\bibitem[\protect\citeauthoryear{Caticha}{2008}]{caticha} 
+Caticha A., 2008, arXiv, arXiv:0808.0012
+
 \bibitem[\protect\citeauthoryear{Eigenbrod et 

### 2013-08-23 doc/Paper1/ms.tex
Random work

@@
 \citep{2004kats.book.....O} so the results obtained are probability
-distributions over the space of possible solutions for the time delays, and the uncertainties in the time delays
+distributions over the space of possible solutions for the time delays, and the
+uncertainties in the time delays
 can be summarised by the widths of these probability distributions. To describe
-the underlying time variation of the quasar and the effects of microlensing on
+prior knowledge about the underlying time variation of the quasar and the
+effects of microlensing on
 the light curves, we use Gaussian Processes (GPs), a family of probability
@@
 with two simulated data sets, one ``easy'' data set and one ``hard'' data set.
-In Section~\ref{sec:b1608} we apply the model to B1608+656.
+In Section~\ref{sec:b1608} we apply the model to B1608+656. Software implementing
+the model described in this paper is freely available online at
+{\tt www.github.com/eggplantnbren/Flotsam}.
 
@@
 We begin by defining our model for the prior knowledge about how the time
-delays give rise rise to the observed data. This will determine the sampling
+delays give rise to the observed data. This will determine the sampling
 distributions, and also the set of parameters (including nuisance parameters) to
-be inferred. Our choice of prior distribution for the parameters will be
-discussed in Section
+be inferred.
 
@@
 \begin{eqnarray}
-y_1(t) &=& m_1 + f(t) + \mu_1(t) \\
-y_2(t) &=& m_2 + f(t - \tau_2) + \mu_2(t) \\
+y_1(t) &=& m_1 + q(t) + \mu_1(t) \\
+y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t) \\
 \end{eqnarray}
@@
 $\tau_2$ is the time delay of image 2 relative to image 1, and the
-functions $\{\mu_1(t), \mu_2(t)\}$ describe any perturbations that are not
-shared by both images. A major source of perturbations is the effect of
-microlensing on each of the light curves. The data then consist of
+functions $\{\mu_1(t), \mu_2(t)\}$ describe any additional fluctuations that
+are not shared by both images. A major source of additional fluctuations
+is the effect of
+microlensing on each of the light curves, which appears because the foreground
+lens galaxy is actually composed of stars rather than smoothly
+distributed matter \citep{falco}.
+The data then consist of
 noisy, irregularly sampled measurements of the light curves
@@
 times $\{t_1, t_2, ..., t_N\}$ will be subsumed into the prior information. The
-data will be the measurements
+data (to which we will assign a ``sampling distribution'') will be the
+measurements
 \begin{eqnarray}
@@
 written as a probability distribution for the measurements $\data$
-given the time delay $\tau_2$ along with some nuisance parameters.
+given the time delay $\tau_2$ along with some nuisance parameters. For a full
+list of parameters, see Table~\ref{tab:prior}. The total number of parameters
+for a two-image system is $13+N$, for a four image system the number of
+parameters is $21+N$.
 
@@
 \hline
-$\tau_2$	&	Time delay relative to image 1	& Cauchy$\left(0, 0.1\tau_{\rm range}\right)T\left(\tau_{\rm min},\tau_{\rm max}\right)$\\
+$\tau_2$	&	Time delay of image 2	&	Cauchy$\left(0, 0.1\tau_{\rm range}\right)T\left(\tau_{\rm min},\tau_{\rm max}\right)$\\
 $\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform$(m_{\rm min}, m_{\rm max})$\\
+$\beta$		&	Size of QSO variations		& LogUniform$(\beta_{\rm min}, \beta_{\rm max})$\\
 $L$		&	Length scale of QSO variations	& LogUniform$(L_{\rm min}, L_{\rm max})$\\
-$\beta$		&	Size of QSO variations		& LogUniform$(\beta_{\rm min}, \beta_{\rm max})$\\
-$A$		&	Size of image-specific variations	& LogUniform$\left(A_{\rm min}, A_{\rm max}\right)$\\
+$\{S_1, S_2\}$	&	Size of image-specific variations	& LogUniform$\left(A_{\rm min}, A_{\rm max}\right)$\\
 $\{\ell_1, \ell_2\}$ &	Length scale of image-specific variations & LogUniform$(\ell_{\rm min}, \ell_{\rm max})$\\
@@
 \hline
+$S_{\rm QSO}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
 $\{B_i\}_{i=1}^N$	&	Badness flag for each data point & $\mathds{1}\left(b_i \leq f_{\rm bad}\right)$\\
@@
 \end{tabular}
+\caption{A full list of parameters, prior information, and data in the
+FLOTSAM model. If there are more than two images, extra parameters are
+added accordingly.\label{tab:prior}}
 \end{table*}
@@
 
+\bibitem[\protect\citeauthoryear{Falco, Wambsganss, 
+\& Schneider}{1991}]{falco} Falco E.~E., Wambsganss J., Schneider P., 1991, MNRAS, 251, 698
+
 \bibitem[\protect\citeauthoryear{Fassnacht et 

### 2013-08-24 doc/Paper1/ms.tex
Cleaning up

@@
 $\beta$		&	Size of QSO variations		& LogUniform$(\beta_{\rm min}, \beta_{\rm max})$\\
-$L$		&	Length scale of QSO variations	& LogUniform$(L_{\rm min}, L_{\rm max})$\\
+$L_q$		&	Length scale of QSO variations	& LogUniform$(L_{\rm min}, L_{\rm max})$\\
 $\{S_1, S_2\}$	&	Size of image-specific variations	& LogUniform$\left(A_{\rm min}, A_{\rm max}\right)$\\
@@
 \hline
-$S_{\rm QSO}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
+$S_{q}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
 $\{B_i\}_{i=1}^N$	&	Badness flag for each data point & $\mathds{1}\left(b_i \leq f_{\rm bad}\right)$\\
@@
 
-We assume that the covariance function for the quasar variability is given by
+We model our uncertainties about the underlying QSO light curve $q(t)$ by
+a GP with mean function $\mathds{E}\left(q(t)\right) = 0$ and covariance function
 \begin{eqnarray}
-\textnormal{Cov}\left[f(t_1), f(t_2)\right] = C(|t_2 - t_1|)
-\end{eqnarray}
-where
-\begin{eqnarray}
-C(\Delta t) = \sigma_{\rm QSO}^2\exp\left(-\frac{\Delta t}{L_{\rm QSO}}\right)
+\textnormal{Cov}\left[q(t_1), q(t_2)\right] = S_q\exp
+\left[-\frac{|t_2 - t_1|}{L_q}\right]
 \end{eqnarray}
 
-{\bf COMMENT ON KELLY}
+This model for QSO variability has been studied extensively by \citep{}
+{\bf COMMENT ON KELLY}. We also note that this distribution is a MaxEnt
+distribution \citep{} with XX constraints \citep{sivia}.
 
-Similarly, the covariance function for each of the microlensing components is
+Similarly, our uncertainty about the image-specific variability is modelled
+as a GP with mean function $\mathds{E}\left(\mu_i(t)\right) = 0$ and
+covariance function
 \begin{eqnarray}
-\textnormal{Cov}\left[\mu(t_1), \mu(t_2)\right] = C_{\rm micro}(|t_2 - t_1|)
-\end{eqnarray}
-where
-\begin{eqnarray}
-C_{\rm micro}(\Delta t) = \sigma_{\rm micro}^2\exp\left(-\left(\frac{\Delta t}{L_{\rm micro}}\right)^{\alpha_{\rm micro}}\right)
+\textnormal{Cov}\left[\mu_i(t_1), \mu_i(t_2)\right] =
+S_i\exp\left[
+-\left(\frac{|t_2 - t_1|
+}{\ell_i}\right)^\alpha
+\right]
 \end{eqnarray}
 
-There is a $\sigma_{\rm micro}$  parameter and an $L_{\rm micro}$ parameter
+There is a separate $S_i$ parameter $\ell_i$ parameter
 for each image; this is because the stellar density and lens convergence is
-different at each image position. In contrast, we use a single $\alpha_{\rm
-micro}$ parameter for all images; this is because it is suspected that the
+different at each image position. If the image-specific variability is
+predominantly caused by microlensing, the typical timescales and amplitudes
+of the microlensing are unlikely to be the same for each image.
+In contrast, we use a single $\alpha$ parameter for all images;
+this is because it is suspected that the
 smoothness of the microlensing variations, which $\alpha$ describes, depends

### 2013-08-24 doc/Paper1/ms.tex
Starting to derive covariance matrix

@@
 primarily on the size of the microlensed source, and all images are of the
-same source.
+same source. We also assume
+prior independence of $q(t)$, $\mu_1(t)$, and $\mu_2(t)$.
 
-The covariance function of the noise-free light curves is then:
+Recall that the noise-free light curves of the two images are given by:
 \begin{eqnarray}
-& & \textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right] \nonumber\\
-&=& \left<y_i(t_1)y_j(t_2)\right> \\
-&=& \left<\left[f(t_1 - \tau_i) + \mu_i(t_1)\right]\left[f(t_2 - \tau_j) + \mu_j(t_2)\right]\right> \\
-&=& \left<f(t_1 - \tau_i)f(t_2 - \tau_j)\right> + \left<\mu_i(t_1)\mu_j(t_2)\right> \\
-&=& C_{\rm QSO}(t_1 - \tau_i - t_2 + \tau_j) + \delta_{ij}C_{\rm micro}(t_1 - t_2)
+y_1(t) &=& m_1 + q(t) + \mu_1(t) \\
+y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t) \\
+\end{eqnarray}
+Using this, and the model assumptions discussed above, we can obtain the
+covariance of $y_1$ with itself at a different time, as well as $y_2$ with itself
+at a different time and $y_1$ and $y_2$ at any time lag.
+
+\begin{eqnarray}
+\textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right]
+&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_j(t_2) - m_j\right)\right]
 \end{eqnarray}

### 2013-08-24 doc/Paper1/ms.tex
More covariance

@@
 \begin{eqnarray}
-y_1(t) &=& m_1 + q(t) + \mu_1(t) \\
-y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t) \\
+y_1(t) &=& m_1 + q(t - \tau_1) + \mu_1(t) \\
+y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t)
 \end{eqnarray}
+where $\tau_1 = 0$.
 Using this, and the model assumptions discussed above, we can obtain the
 covariance of $y_1$ with itself at a different time, as well as $y_2$ with itself
-at a different time and $y_1$ and $y_2$ at any time lag.
+at a different time, and $y_1$ and $y_2$ at any time lag.
 
 \begin{eqnarray}
-\textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right]
-&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_j(t_2) - m_j\right)\right]
+\textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right]\\
+= \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_j(t_2) - m_j\right)\right]\\
+= \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
+\left(q(t_2 - \tau_j) + \mu_j(t_2)\right)\right]
 \end{eqnarray}
-where the intrinsic variations have been assumed to be uncorrelated with the
-microlensing contributions.
 

### 2013-08-24 doc/Paper1/ms.tex
Covariance derivations in appendix

@@
 
-Recall that the noise-free light curves of the two images are given by:
-\begin{eqnarray}
-y_1(t) &=& m_1 + q(t - \tau_1) + \mu_1(t) \\
-y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t)
-\end{eqnarray}
-where $\tau_1 = 0$.
-Using this, and the model assumptions discussed above, we can obtain the
-covariance of $y_1$ with itself at a different time, as well as $y_2$ with itself
-at a different time, and $y_1$ and $y_2$ at any time lag.
-
-\begin{eqnarray}
-\textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right]\\
-= \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_j(t_2) - m_j\right)\right]\\
-= \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
-\left(q(t_2 - \tau_j) + \mu_j(t_2)\right)\right]
-\end{eqnarray}
+See the Appendix for derivations.
 
@@
 improve the precision of the time delay inferences {\bf (WE HOPE)}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-\begin{center}
-\begin{tabular}{ccc}
-Parameter & \vline & Prior \\
-\hline
-$\tau$ & \vline & $U(-\Delta t/2, \Delta t/2)$ \\ 
-$m$ & \vline & $U(y_{\rm min}, y_{\rm max})$ \\
-$\sigma_{\rm QSO}$ & \vline & $LU(10^{-3}, 10)$ \\
-$L_{\rm QSO}$ & \vline & $LU(10^{-3}\Delta t, 10\Delta t)$ \\
-$\sigma_{\rm micro}$ & \vline & $LU(10^{-3}, 10)$ \\
-$L_{\rm micro}$ & \vline & $LU(10^{-3}\Delta t, 10\Delta t)$ \\
-$\alpha_{\rm micro}$ & \vline & $U(1, 2)$ \\
-$\kappa$ & \vline & $LU(1, 10)$
-\end{tabular}
-\end{center}
-\caption{Prior probability distributions for the parameters. $U(a, b)$ denotes
-a uniform prior between $a$ and $b$, and $LU(a, b)$ denotes a log-uniform
-(that is, uniform in the logarithm of the parameter) prior $\propto 1/x$
-between $a$ and $b$. $\Delta t$ is the duration of the data set. All priors
-are independent.\label{tab:prior}}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\appendix
+\onecolumn
+
+\section{GP Covariance Function}
+Recall that the noise-free light curves of the two images are given by:
+\begin{eqnarray}
+y_1(t) &=& m_1 + q(t - \tau_1) + \mu_1(t) \\
+y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t)
+\end{eqnarray}
+where $\tau_1 = 0$.
+Using this, and the model assumptions discussed above, we can obtain the
+covariance of $y_1$ with itself at a different time, as well as $y_2$ with itself
+at a different time, and $y_1$ and $y_2$ at any time lag. For distinct images
+$i$ and $j$ the covariance is:
+\begin{eqnarray}
+\textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right]
+&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_j(t_2) - m_j\right)\right]\\
+&=& \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
+\left(q(t_2 - \tau_j) + \mu_j(t_2)\right)\right]\\
+&=& \mathds{E}\left[q(t_1 - \tau_i)q(t_2 - \tau_j) + q(t_1 - \tau_i)\mu_j(t_2)
++ \mu_i(t_1)q(t_2 - \tau_j) + \mu_i(t_1)\mu_j(t_2)
+\right]\\
+&=& S_q\exp
+\left[-\frac{|(t_2 -\tau_j)- (t_1-\tau_i)|}{L_q}\right].
+\end{eqnarray}
+For a single image $i$, the covariance is:
+\begin{eqnarray}
+\textnormal{Cov}\left[y_i(t_1), y_i(t_2)\right]
+&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_i(t_2) - m_i\right)\right]\\
+&=& \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
+\left(q(t_2 - \tau_i) + \mu_i(t_2)\right)\right]\\
+&=& \mathds{E}\left[q(t_1 - \tau_i)q(t_2 - \tau_i) + q(t_1 - \tau_i)\mu_i(t_2)
++ \mu_i(t_1)q(t_2 - \tau_i) + \mu_i(t_1)\mu_i(t_2)
+\right]\\
+&=& S_q\exp
+\left[-\frac{|(t_2 -\tau_j)- (t_1-\tau_i)|}{L_q}\right] +
+S_i\exp\left[
+-\left(\frac{|t_2 - t_1|
+}{\ell_i}\right)^\alpha
+\right]
+.
+\end{eqnarray}
+
 \end{document}

### 2013-08-24 doc/Paper1/ms.tex
Noise term

@@
 \end{eqnarray}
+The covariance between a data point $Y_i$ and another data point $Y_j$ is
+given by the covariance of the underlying noise-free light curves at times $t_i$
+and $t_j$, plus a noise term:
+\begin{eqnarray}
+\textnormal{Cov}\left(
+Y_i, Y_j
+\right) = \textnormal{Cov}\left(y_{k_i}(t_i), y_{k_j}(t_j)\right)
++ \delta_{ij}\left[\left[\kappa_1 + (\kappa_2 - \kappa_1)\mathds{1}(B_i = 1)\right]\sigma_i\right]^2
+\end{eqnarray}
+where $\sigma_i$ is the reported error bar given with the data. The coefficient
+in front of $\sigma_i$ implies that the error bar should be multiplied by
+$\kappa_1$ if this is a ``good'' data point, or by $\kappa_2$ if it is a ``bad''
+data point.
 

### 2013-08-24 doc/Paper1/ms.tex
Wording

@@
 where $\tau_1 = 0$.
-Using this, and the model assumptions discussed above, we can obtain the
-covariance of $y_1$ with itself at a different time, as well as $y_2$ with itself
-at a different time, and $y_1$ and $y_2$ at any time lag. For distinct images
-$i$ and $j$ the covariance is:
+Using this, and the model assumptions discussed in the paper, we can obtain the
+covariance of image $i$ at time $t_1$ with image $j$ at time $t_2$, where
+$i \neq j$:
 \begin{eqnarray}
@@
 \end{eqnarray}
-For a single image $i$, the covariance is:
+We can also obtain the covariance of image $i$ measured at two times $t_1$ and
+$t_2$:
 \begin{eqnarray}

### 2013-08-25 doc/Paper1/ms.tex
Different notation for error bar boosts

@@
 $S_{q}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
-$\{B_i\}_{i=1}^N$	&	Badness flag for each data point & $\mathds{1}\left(b_i \leq f_{\rm bad}\right)$\\
+$\{a_i\}_{i=1}^N$	&	Errorbar multiplier for each data point & $\left\{\begin{array}{lr}\kappa_1, & b_i > f_{\rm bad}\\ \kappa_2, & b_i \leq f_{\rm bad}\end{array}\right.$\\
 \hline
@@
 This is another mechanism for accounting for outliers.
-
-In summary, the parameter vector is
-\begin{equation}
-\left\{ \{\tau\}, \{m\}, \sigma_{\rm QSO}, L_{\rm QSO}, \{\sigma_{\rm micro}\}
-\{L_{\rm micro}\}, \alpha_{\rm micro}, \kappa \right\}
-\end{equation}
-where parameters in braces denote a set (one for each QSO image).  For a
-double image lens, the total number of parameters in the model is~12. For a
-quad, this number is~20. A set of uninformative prior
-distributions for these parameters are listed in Table~\ref{priors}; we assign
-these priors in the analysis described in the first part of this paper. Later,
-we will show how a more informative set of priors can be derived, and used to
-improve the precision of the time delay inferences {\bf (WE HOPE)}.
 
@@
 covariance of image $i$ at time $t_1$ with image $j$ at time $t_2$, where
-$i \neq j$:
+$i \neq j$. Here, $t_1$ and $t_2$ stand for any two times.
 \begin{eqnarray}
@@
 \right) = \textnormal{Cov}\left(y_{k_i}(t_i), y_{k_j}(t_j)\right)
-+ \delta_{ij}\left[\left[\kappa_1 + (\kappa_2 - \kappa_1)\mathds{1}(B_i = 1)\right]\sigma_i\right]^2
++ \delta_{ij}\left(a_i\sigma_i\right)^2
 \end{eqnarray}
 where $\sigma_i$ is the reported error bar given with the data. The coefficient
-in front of $\sigma_i$ implies that the error bar should be multiplied by
-$\kappa_1$ if this is a ``good'' data point, or by $\kappa_2$ if it is a ``bad''
-data point.
+$a_i$
+in front of $\sigma_i$ is the robustness feature that depends on $\kappa_1$,
+$\kappa_2$, and the badness variable for data point $i$.
 

### 2013-08-27 doc/Paper1/ms.tex
changed kappa notation

@@
 $\kappa_1$	&	Error-bar boost for ``good'' data & $\frac{1}{2}\delta(\kappa_1 - 1) + \frac{1}{2}$LogUniform$(1,100)$\\
-$\kappa_2$	&	Error-bar boost for ``bad'' data & LogUniform$(\kappa_1, 100\kappa_1)$\\
+$\kappa_2$	&	Extra error-bar boost for ``bad'' data & LogUniform$(1, 100)$\\
 $\{b_i\}_{i=1}^N$ &	Badness hidden variables for each data point & Uniform$(0,1)$\\
@@
 $S_{q}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
-$\{a_i\}_{i=1}^N$	&	Errorbar multiplier for each data point & $\left\{\begin{array}{lr}\kappa_1, & b_i > f_{\rm bad}\\ \kappa_2, & b_i \leq f_{\rm bad}\end{array}\right.$\\
+$\{a_i\}_{i=1}^N$	&	Errorbar multiplier for each data point & $\left\{\begin{array}{lr}\kappa_1, & b_i > f_{\rm bad}\\ \kappa_1\kappa_2, & b_i \leq f_{\rm bad}\end{array}\right.$\\
 \hline
@@
 than it is by a time-correlated microlensing term. Thus, to improve the
-robustness of our method, we include a ``noise boost'' parameter $\kappa$
+robustness of our method, we include ``noise boost'' parameters $\kappa_1$,
+$\kappa_2$, and $f_{\rm bad}$,
 describing the factor by which the reported error bars should be multiplied.
@@
 \right) = \textnormal{Cov}\left(y_{k_i}(t_i), y_{k_j}(t_j)\right)
-+ \delta_{ij}\left(a_i\sigma_i\right)^2
++ \delta_{ij}\left(a_i\sigma_i\right)^2\label{eq:covariance}
 \end{eqnarray}
@@
 in front of $\sigma_i$ is the robustness feature that depends on $\kappa_1$,
-$\kappa_2$, and the badness variable for data point $i$.
+$\kappa_2$, and the badness variable for data point $i$. Equation~\ref{eq:covariance}
+provides the elements of the covariance matrix in the multivariate normal
+sampling distribution.
 

### 2013-08-28 src/DNest3/suyu.py
Compare to Sherry

@@ -1,0 +1,50 @@
+from pylab import *
+
+#This system is special
+#in that the three relative time delays between the four images were measured accurately with errors of only a few
+#percent: tAB = 31.5
+#+2.0
+#-1.0 days, tCB = 36.0
+#+1.5
+#-1.5 days,
+#and tDB = 77.0
+#+2.0
+#-1.0 days (Fassnacht et al. 1999, 2002).
+
+# IMAGE ID 0 = A
+# IMAGE ID 1 = B
+# IMAGE ID 2 = C
+# IMAGE ID 3 = D
+
+posterior_sample = loadtxt('posterior_sample.txt')
+
+tAB = posterior_sample[:,4] - posterior_sample[:,5]
+tCB = posterior_sample[:,6] - posterior_sample[:,5]
+tDB = posterior_sample[:,7] - posterior_sample[:,5]
+
+hist(tAB, 50, alpha=0.2, color='b')
+hist(tCB, 50, alpha=0.2, color='r')
+hist(tDB, 50, alpha=0.2, color='g')
+xlabel('Time Delay $\\tau$ (days)')
+show()
+
+print(mean(tAB), std(tAB))
+print(mean(tCB), std(tCB))
+print(mean(tDB), std(tDB))
+
+
+# Find 68% credible intervals
+def summary(theta):
+	s = sort(theta)
+	a = s[int(0.16*len(theta))]
+	b = s[int(0.5*len(theta))]
+	c = s[int(0.84*len(theta))]
+	return (b - a, b, c - b)
+
+print(summary(tAB))
+print(summary(tCB))
+print(summary(tDB))
+
+
+
+

### 2013-08-28 src/DNest3/suyu.py
PLot our credible intervals and Fassnacht's

@@ -25,13 +25,6 @@ tDB = posterior_sample[:,7] - posterior_sample[:,5]
 hist(tAB, 50, alpha=0.2, color='b')
 hist(tCB, 50, alpha=0.2, color='r')
 hist(tDB, 50, alpha=0.2, color='g')
-xlabel('Time Delay $\\tau$ (days)')
-show()
-
-print(mean(tAB), std(tAB))
-print(mean(tCB), std(tCB))
-print(mean(tDB), std(tDB))
-
 
 # Find 68% credible intervals
 def summary(theta):
@@ -41,9 +34,33 @@ def summary(theta):
 	c = s[int(0.84*len(theta))]
 	return (b - a, b, c - b)
 
-print(summary(tAB))
-print(summary(tCB))
-print(summary(tDB))
+summary1 = summary(tAB)
+summary2 = summary(tCB)
+summary3 = summary(tDB)
+
+axhline(0, linewidth=2, color='k')
+
+# Plot summaries
+errorbar([summary1[1]], [-0.3], xerr=[[summary1[0]], [summary1[2]]], fmt='bo', linewidth=2)
+errorbar([summary2[1]], [-0.2], xerr=[[summary2[0]], [summary2[2]]], fmt='ro', linewidth=2)
+errorbar([summary3[1]], [-0.1], xerr=[[summary3[0]], [summary3[2]]], fmt='go', linewidth=2)
+
+# Plot Fassnacht's time delays
+errorbar([31.5], [-0.5], xerr=[[1.], [2.]], fmt='bo', linewidth=2)
+errorbar([36.], [-0.5], xerr=[[1.5], [1.5]], fmt='ro', linewidth=2)
+errorbar([77.], [-0.5], xerr=[[1.], [2.]], fmt='go', linewidth=2)
+
+ylim(-1)
+xlabel('Time Delay $\\tau$ (days)')
+show()
+
+print(mean(tAB), std(tAB))
+print(mean(tCB), std(tCB))
+print(mean(tDB), std(tDB))
+
+print(summary1)
+print(summary2)
+print(summary3)
 
 
 

### 2013-08-28 src/DNest3/suyu.py
Ratio of credible interval lengths

@@ -62,6 +62,9 @@ print(summary1)
 print(summary2)
 print(summary3)
 
+print((summary1[2] + summary1[0])/3.)
+print((summary2[2] + summary2[0])/3.)
+print((summary3[2] + summary3[0])/3.)
 
 
 

### 2013-08-29 src/DNest3/Limits.cpp
Parameterise magnitudes by m0 and then others relative to that, with cauchy(0,1) priors on the latter.

@@ -12,15 +12,9 @@ Limits::Limits()
 
 void Limits::set(const Data& data)
 {
-	mag_min.resize(data.get_numImages());
-	mag_max.resize(data.get_numImages());
-	mag_range.resize(data.get_numImages());
-	for(int i=0; i<data.get_numImages(); i++)
-	{
-		mag_min[i] = data.get_yMean()[i] - 30.*data.get_yStDev()[i]; 
-		mag_max[i] = data.get_yMean()[i] + 30.*data.get_yStDev()[i];
-		mag_range[i] = mag_max[i] - mag_min[i];
-	}
+	mag_min = data.get_yMean()[0] - 30.*data.get_yStDev()[0]; 
+	mag_max = data.get_yMean()[0] + 30.*data.get_yStDev()[0];
+	mag_range = mag_max - mag_min;
 
 	tau_min = -data.get_tRange();
 	tau_max =  data.get_tRange();

### 2013-08-29 src/DNest3/Limits.h

@@ -14,8 +14,8 @@ class Limits
 	private:
 		// Naming convention: varName_min, varName_max, varName_range
 
-		// Mean magnitudes
-		std::vector<double> mag_min, mag_max, mag_range;
+		// Magnitude for image 0
+		double mag_min, mag_max, mag_range;
 
 		// Time delays (first is zero by definition)
 		double tau_min, tau_max, tau_range;

### 2013-08-29 src/DNest3/TDModel.cpp

@@ -44,7 +44,10 @@ void TDModel::fromPrior()
 {
 	for(int i=0; i<numImages; i++)
 	{
-		mag[i] = limits.mag_min[i] + limits.mag_range[i]*randomU();
+		if(i==0)
+			mag[i] = limits.mag_min + limits.mag_range*randomU();
+		else
+			mag[i] = tan(M_PI*(randomU() - 0.5));
 
 		if(i == 0)
 			tau[i] = 0.;
@@ -85,11 +88,22 @@ double TDModel::perturb1()
 {
 	int which = randInt(numImages);
 
-	mag[which] += limits.mag_range[which]
-			*pow(10., 1.5 - 6.*randomU())*randn();
-	mag[which] = mod(mag[which] - limits.mag_min[which]
-				,limits.mag_range[which])
-				+ limits.mag_min[which];
+	if(which == 0)
+	{
+		mag[which] += limits.mag_range
+				*pow(10., 1.5 - 6.*randomU())*randn();
+		mag[which] = mod(mag[which] - limits.mag_min
+					,limits.mag_range)
+					+ limits.mag_min;
+	}
+	else
+	{
+		mag[which] = atan(mag[which])/M_PI + 0.5;
+		mag[which] += pow(10., 1.5 - 6.*randomU())*randn();
+		mag[which] = mod(mag[which], 1.);
+		mag[which] = tan(M_PI*(mag[which] - 0.5));
+	}
+
 	return 0.;
 }
 
@@ -278,7 +292,12 @@ void TDModel::formCovarianceMatrix()
 void TDModel::formMeanVector()
 {
 	for(int i=0; i<numPoints; i++)
-		meanVector(i) = mag[Data::get_instance().get_ID()[i]];
+	{
+		if(Data::get_instance().get_ID()[i] == 0)
+			meanVector(i) = mag[0];
+		else
+			meanVector(i) = mag[0] + mag[Data::get_instance().get_ID()[i]];
+	}
 }
 
 double TDModel::covariance(double t1, double t2, int ID1, int ID2)

### 2013-08-29 src/DNest3/TDModel.h

@@ -18,7 +18,8 @@ class TDModel:public DNest3::Model
 		// Useful integers
 		int numPoints, numImages;
 
-		// Mean magnitudes
+		// Mean magnitude of image 0, followed by magnitudes of
+		// other images relative to image 0
 		std::vector<double> mag;
 
 		// Time delays (first is zero by definition)

### 2013-09-02 src/Microlensing/microlensing.py
Save the rays

@@ -12,7 +12,7 @@ microlensing light curves for battle testing Flotsam.
 rng.seed(123)
 
 # Number of rays per dimension
-N = 1000
+N = 4000
 
 # The rectangle in the lens plane
 x_min, x_max = -10., 10.
@@ -28,7 +28,7 @@ y = y[::-1, :]
 
 # Convergence, external shear, number of stars
 # Note: "gamma" DOES NOTHING
-rho, gamma, num_stars = 0.3, 0., 100
+rho, gamma, num_stars = 0.5, 0., 500
 
 # Generate the stars, in a circle
 radius = np.min([0.5*(x_max - x_min), 0.5*(y_max - y_min)])
@@ -48,6 +48,7 @@ for i in xrange(0, num_stars):
 		rsq = x_stars[i]**2 + y_stars[i]**2
 
 plt.plot(x_stars, y_stars, 'r*')
+plt.title('The Stars')
 plt.axis('scaled')
 plt.show()
 
@@ -63,12 +64,25 @@ for i in xrange(0, num_stars):
 	xs -= mass*(x - x_stars[i])/(np.pi*rsq)
 	ys -= mass*(y - y_stars[i])/(np.pi*rsq)
 
-	if (i+1)%1 == 0:
-		counts = scipy.histogram2d(xs.flatten(), ys.flatten(), bins=1000,\
+	print('{i}/{n}'.format(i=(i+1), n=num_stars))
+
+	if (i+1)%50 == 0:
+		counts = scipy.histogram2d(xs.flatten(), ys.flatten(), bins=250,\
 					range=[[x_min, x_max], [y_min, y_max]])
 		plt.imshow(counts[0])
 		plt.title('{a}/{b}'.format(a=(i+1), b=num_stars))
 		plt.draw()
+
+rays = np.empty((xs.size, 2))
+rays[:,0] = xs.flatten()
+rays[:,1] = ys.flatten()
+
+# Save the rays
+import cPickle as pickle
+print('Saving...')
+pickle.dump(rays, open('rays.pickle', 'wb'))
+print('Done.')
+
 plt.ioff()
 plt.show()
 

### 2013-09-02 src/Microlensing/simulate_data.py
Make simulated data

@@ -1,0 +1,44 @@
+from pylab import *
+
+def ar1(n=10000, a=0.999, b=1.):
+	y = zeros(n)
+	for i in xrange(1, n):
+		y[i] = a*y[i-1] + b*randn()
+	return y
+
+seed(123)
+
+qso = ar1()/20.
+
+# Observation times
+y1 = qso[8100:9100:10].copy()
+y2 = qso[8000:9000:10].copy()
+N = y1.size
+
+# Offset
+y2 += 2.
+
+# Invent timestamps
+t = arange(0, N)
+
+# Add noise
+y1 += 0.3*randn(N)
+y2 += 0.3*randn(N)
+
+# Make Flotsam data
+data = empty((2*N, 4))
+
+data[0:N, 0] = t
+data[0:N, 1] = y1
+data[0:N, 2] = 0.3
+data[0:N, 3] = 0
+data[N:, 0] = t
+data[N:, 1] = y2
+data[N:, 2] = 0.3
+data[N:, 3] = 1
+
+errorbar(data[0:N, 0], data[0:N, 1], yerr=data[0:N, 2], fmt='bo')
+errorbar(data[N:, 0], data[N:, 1], yerr=data[N:, 2], fmt='ro')
+show()
+
+savetxt('easy_data.txt', data)

### 2013-09-02 src/Microlensing/simulate_data.py
Make hard data

@@ -42,3 +42,71 @@ errorbar(data[N:, 0], data[N:, 1], yerr=data[N:, 2], fmt='ro')
 show()
 
 savetxt('easy_data.txt', data)
+
+# Add microlensing to make hard data
+# Load the rays
+import cPickle as pickle
+print('Loading...')
+rays = pickle.load(open('rays.pickle', 'rb'))
+print('Done.')
+
+# FIRST IMAGE
+x0 = -1.
+y0 = -1.
+vx = 0.01
+vy = 0.01
+r = 0.05
+
+x = x0 + vx*arange(0, N)
+y = y0 + vy*arange(0, N)
+
+plot(rays[0::37,0], rays[0::37,1], 'b.', markersize=1)
+plot(x, y, 'ro-', markersize=10)
+axis([-4., 4., -4., 4.])
+show()
+
+mag = zeros(N)
+ion()
+hold(False)
+for i in xrange(0, N):
+	mag[i] = 2.5*log10(sum((rays[:,0] - x[i])**2 + (rays[:,1] - x[i])**2 <= r**2))
+	plot(mag[0:(i+1)])
+	draw()
+ioff()
+show()
+
+data[0:N, 1] += mag - mean(mag)
+
+# SECOND IMAGE
+x0 = 1.
+y0 = 1.
+vx = 0.01
+vy = -0.01
+r = 0.05
+
+x = x0 + vx*arange(0, N)
+y = y0 + vy*arange(0, N)
+
+hold(True)
+plot(rays[0::37,0], rays[0::37,1], 'b.', markersize=1)
+plot(x, y, 'ro-', markersize=10)
+axis([-4., 4., -4., 4.])
+show()
+
+mag = zeros(N)
+ion()
+hold(False)
+for i in xrange(0, N):
+	mag[i] = 2.5*log10(sum((rays[:,0] - x[i])**2 + (rays[:,1] - x[i])**2 <= r**2))
+	plot(mag[0:(i+1)])
+	draw()
+ioff()
+show()
+data[N:, 1] += mag - mean(mag)
+
+savetxt('hard_data.txt', data)
+
+hold(True)
+errorbar(data[0:N, 0], data[0:N, 1], yerr=data[0:N, 2], fmt='bo')
+errorbar(data[N:, 0], data[N:, 1], yerr=data[N:, 2], fmt='ro')
+show()

### 2013-09-03 src/DNest3/suyu.py
Better plotting

@@ -22,9 +22,10 @@ tAB = posterior_sample[:,4] - posterior_sample[:,5]
 tCB = posterior_sample[:,6] - posterior_sample[:,5]
 tDB = posterior_sample[:,7] - posterior_sample[:,5]
 
-hist(tAB, 50, alpha=0.2, color='b')
-hist(tCB, 50, alpha=0.2, color='r')
-hist(tDB, 50, alpha=0.2, color='g')
+h1 = hist(tAB, 50, alpha=0.2, color='b')
+h2 = hist(tCB, 50, alpha=0.2, color='r')
+h3 = hist(tDB, 50, alpha=0.2, color='g')
+height = max((h1[0].max(), h2[0].max(), h3[0].max()))
 
 # Find 68% credible intervals
 def summary(theta):
@@ -41,16 +42,16 @@ summary3 = summary(tDB)
 axhline(0, linewidth=2, color='k')
 
 # Plot summaries
-errorbar([summary1[1]], [-0.3], xerr=[[summary1[0]], [summary1[2]]], fmt='bo', linewidth=2)
-errorbar([summary2[1]], [-0.2], xerr=[[summary2[0]], [summary2[2]]], fmt='ro', linewidth=2)
-errorbar([summary3[1]], [-0.1], xerr=[[summary3[0]], [summary3[2]]], fmt='go', linewidth=2)
+errorbar([summary1[1]], [-0.03*height], xerr=[[summary1[0]], [summary1[2]]], fmt='bo', linewidth=2)
+errorbar([summary2[1]], [-0.05*height], xerr=[[summary2[0]], [summary2[2]]], fmt='ro', linewidth=2)
+errorbar([summary3[1]], [-0.05*height], xerr=[[summary3[0]], [summary3[2]]], fmt='go', linewidth=2)
 
 # Plot Fassnacht's time delays
-errorbar([31.5], [-0.5], xerr=[[1.], [2.]], fmt='bo', linewidth=2)
-errorbar([36.], [-0.5], xerr=[[1.5], [1.5]], fmt='ro', linewidth=2)
-errorbar([77.], [-0.5], xerr=[[1.], [2.]], fmt='go', linewidth=2)
+errorbar([31.5], [-0.1*height], xerr=[[1.], [2.]], fmt='bo', linewidth=2, alpha=0.2)
[…]

### 2013-09-03 doc/Paper1/ms.tex
Simulated data figure

@@
 \hline
+$\tau_1$	&	Time delay of image 1	&	$\delta(\tau_1 - 0)$\\
 $\tau_2$	&	Time delay of image 2	&	Cauchy$\left(0, 0.1\tau_{\rm range}\right)T\left(\tau_{\rm min},\tau_{\rm max}\right)$\\
-$\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform$(m_{\rm min}, m_{\rm max})$\\
+$m_1$	&	Mean magnitude for image 1	& Uniform$(m_{\rm min}, m_{\rm max})$\\
+$m'_2$		&	Magnitude of image 2 relative to image 1	& Cauchy$(0,1)$\\
 $\beta$		&	Size of QSO variations		& LogUniform$(\beta_{\rm min}, \beta_{\rm max})$\\
@@
 \hline
+$m_2$	&	Magnitude of image 2	& $m_1 + m'_2$\\
 $S_{q}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
@@
 affected by observing strategy?
+
+\section{Demonstration on Simulated Data}
+In this section we demonstrate the inference methods on simulated data. We
+created an ``easy'' data set consisting solely of quasar variability with no
+microlensing. Then, a ``hard'' data set was created by adding microlensing
+variability. The data consisted of 100 measurements of two images over a time
+span of 100 days. The true time delay was $\tau_2 = 10$ days. The two data
+sets are plotted in Figure~\ref{fig:sim_data}.
+
+\begin{figure*}
+\includegraphics[scale=0.4]{Figures/sim_data.eps}
+\caption{\label{fig:sim_data}}
+\end{figure*}
 

### 2013-09-03 doc/Paper1/Figures/sim_data.py

@@ -1,0 +1,46 @@
+from pylab import *
+
+rc("font", size=20, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
+data = loadtxt('../../../src/Microlensing/easy_data.txt')
+N = data.shape[0]//2
+
+figure(figsize=(20, 6))
+hold(True)
+
+subplot(1,3,1)
+errorbar(data[0:N, 0], data[0:N, 1], yerr=data[0:N, 2], fmt='bo', label='Image 1')
+errorbar(data[N:, 0], data[N:, 1], yerr=data[N:, 2], fmt='ro', label='Image 2')
+xlabel('$t$ (days)')
+ylabel('$Y$ (magnitudes)')
+title('Easy Data')
+axis([-3, 103, -3, 4])
+legend(loc='upper left', numpoints=1)
+
+easy = data
+data = loadtxt('../../../src/Microlensing/hard_data.txt')
+
+subplot(1,3,2)
+errorbar(data[0:N, 0], data[0:N, 1], yerr=data[0:N, 2], fmt='bo', label='Image 1')
+errorbar(data[N:, 0], data[N:, 1], yerr=data[N:, 2], fmt='ro', label='Image 2')
+xlabel('$t$ (days)')
+ylabel('$Y$ (magnitudes)')
+title('Hard Data')
+axis([-3, 103, -3, 4])
+legend(loc='upper left', numpoints=1)
+
+subplot(1,3,3)
+plot(data[0:N, 0], data[0:N, 1] - easy[0:N, 1], 'b', linewidth=2, label='Image 1')
+plot(data[N:, 0], data[N:, 1] - easy[N:, 1], 'r', linewidth=2, label='Image 2')
+xlabel('$t$ (days)')
+ylabel('$\\mu(t)$ (magnitudes)')
+title('Microlensing in Hard Data')
+axis([-3, 103, -3, 4])
+legend(loc='upper left', numpoints=1)
+
+savefig('sim_data.pdf', bbox_inches='tight')
+import os
+os.system('pdftops -eps sim_data.pdf')
+show()
+

### 2013-09-03 doc/Paper1/Figures/b1608_data.py
Plot B1608 data

@@ -1,0 +1,29 @@
+from pylab import *
+
+rc("font", size=20, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
+data = loadtxt('../../../src/DNest3/1608_all_mags.txt')
+
+data[:,0] -= data[:,0].min()
+
+figure(figsize=(16, 8))
+colors = ['b', 'r', 'g', 'y']
+shapes = ['o', 's', 'p', '*']
+sizes  = [6, 5, 6, 7]
+names  = ['A', 'B', 'C', 'D']
+for i in xrange(0, 4):
+	which = data[:,3] == i
+	plot(data[which, 0], data[which, 1], shapes[i], color=colors[i], markersize=sizes[i], label='Image {I}'.format(I=names[i]))
+legend(loc='upper left', numpoints=1)
+axis([-50, 1250, -4, 0])
+title('B1608+656 Data')
+
+xlabel('$t$ (days)')
+ylabel('$Y$ (magnitudes)')
+
+savefig('b1608_data.pdf', bbox_inches='tight')
+import os
+os.system('pdftops -eps b1608_data.pdf')
+show()
+

### 2013-09-21 src/DNest3/TDModel.cpp
Significant speedups

@@ -31,6 +31,8 @@ TDModel::TDModel()
 	tau.resize(numImages);
 	logSig_ml.resize(numImages);
 	logTau_ml.resize(numImages);
+	sig_ml.resize(numImages);
+	tau_ml.resize(numImages);
 
 	meanVector = Vector(numPoints);
 	covarianceMatrix = Matrix(numPoints, numPoints);
@@ -257,6 +259,15 @@ double TDModel::perturb()
 
 void TDModel::formCovarianceMatrix()
 {
+	sig_qso = exp(logSig_qso);
+	tau_qso = exp(logTau_qso);
+	for(int i=0; i<numImages; i++)
+	{
+		sig_ml[i] = exp(logSig_ml[i]);
+		tau_ml[i] = exp(logTau_ml[i]);
+	}
+	coeff = pow(sig_qso, 2)*tau_qso;
+
 	// Fill covariance matrix
 	// with covariance function evaluations
 	for(int i=0; i<numPoints; i++)
@@ -302,26 +313,11 @@ void TDModel::formMeanVector()
 
 double TDModel::covariance(double t1, double t2, int ID1, int ID2)
 {
-	double sig_qso = exp(logSig_qso);
-	double tau_qso = exp(logTau_qso);
-
-	double exponent = abs((t1 - tau[ID1]) - (t2 - tau[ID2]))/tau_qso;
-	double C = pow(sig_qso*sqrt(tau_qso), 2)
-				*exp(-exponent);
-
-	vector<double> sig_ml(numImages);
-	vector<double> tau_ml(numImages);
-	for(int i=0; i<numImages; i++)
-	{
-		sig_ml[i] = exp(logSig_ml[i]);
-		tau_ml[i] = exp(logTau_ml[i]);
-	}
+	double C = coeff*exp(-abs((t1 - tau[ID1]) - (t2 - tau[ID2]))/tau_qso);
 
 	if(ID1 == ID2)
-	{
-		exponent = pow(abs(t1 - t2)/tau_ml[ID1], alpha);
-		C += pow(sig_ml[ID1], 2)*exp(-exponent);
-	}
+		C += pow(sig_ml[ID1], 2)*exp(-pow(abs(t1 - t2)/tau_ml[ID1], alpha));
+
 	return C;
 }
 

### 2013-09-21 src/DNest3/TDModel.h

@@ -45,6 +45,11 @@ class TDModel:public DNest3::Model
 		double latent_boost1, boost1, boost2;
 		std::vector<double> bad_uniforms; // U(0,1) prior
 
+
+		// Temporary stuff
+		std::vector<double> sig_ml, tau_ml;
+		double sig_qso, tau_qso, coeff;
+
 		// Mean vector
 		Vector meanVector;
 

### 2013-09-21 src/Brute_Force/MyModel.cpp
Magnitude priors

@@ -20,23 +20,42 @@
 #include "MyModel.h"
 #include "RandomNumberGenerator.h"
 #include "Utils.h"
+#include "Data.h"
 #include <cmath>
 
 using namespace std;
 using namespace DNest3;
 
 MyModel::MyModel()
+:delta_mag(Data::get_instance().get_numImages())
 {
 
 }
 
 void MyModel::fromPrior()
 {
-
+	mag0 = -100. + 200.*randomU();
+	for(size_t i=0; i<delta_mag.size(); i++)
+		delta_mag[i] = tan(M_PI*(randomU() - 0.5));
 }
 
 double MyModel::perturb()
 {
+	int which = randInt(2);
+	if(which == 0)
+	{
+		mag0 += 200.*pow(10., 1.5 - 6.*randomU())*randn();
+		mag0 = mod(mag0 + 100., 200.) - 100.;
+	}
+	else
+	{
+		int which2 = randInt(delta_mag.size());
+		double u = 0.5 + atan(delta_mag[which2])/M_PI;
+		u += pow(10., 1.5 - 6.*randomU())*randn();
+		u = mod(u, 1.);
+		delta_mag[which2] = tan(M_PI*(u - 0.5));
+	}
+
 	return 0.;
 }
 
@@ -47,7 +66,9 @@ double MyModel::logLikelihood() const
 
 void MyModel::print(std::ostream& out) const
 {
-
+	out<<mag0<<' ';
+	for(size_t i=0; i<delta_mag.size(); i++)
+		out<<delta_mag[i]<<' ';
 }
 
 string MyModel::description() const

### 2013-09-21 src/Brute_Force/MyModel.h

@@ -26,6 +26,8 @@
 class MyModel:public DNest3::Model
 {
 	private:
+		double mag0;
+		std::vector<double> delta_mag;
 
 	public:
 		MyModel();

### 2013-09-21 src/Brute_Force/main.cpp

@@ -20,18 +20,22 @@
 #include <iostream>
 #include "Start.h"
 #include "MyModel.h"
+#include "Data.h"
 
 using namespace std;
 using namespace DNest3;
 
 int main(int argc, char** argv)
 {
-	#ifndef DNest3_No_Boost
-	MTSampler<MyModel> sampler = setup_mt<MyModel>(argc, argv);
-	#else
-	Sampler<MyModel> sampler = setup<MyModel>(argc, argv);
-	#endif
+	// Process command line options and load data
+	CommandLineOptions options(argc, argv);
+	string dataFile = options.get_dataFile();
+	Data::get_instance().load(dataFile.c_str());
 
+	// Initialise the sampler
+	MTSampler<MyModel> sampler = setup_mt<MyModel>(options);
+
+	// Go!
 	sampler.run();
 	return 0;
 }

### 2013-09-21 src/Brute_Force/MyModel.h
Comments

@@ -26,7 +26,10 @@
 class MyModel:public DNest3::Model
 {
 	private:
+		// Mean magnitude of image 0
 		double mag0;
+
+		// Magnitudes of others relative to image 0 (ignore delta_mag[0])
 		std::vector<double> delta_mag;
 
 	public:

### 2013-09-21 src/Brute_Force/MyModel.cpp
QSO variability parameters

@@ -28,6 +28,7 @@ using namespace DNest3;
 
 MyModel::MyModel()
 :delta_mag(Data::get_instance().get_numImages())
+,n_qso(1000)
 {
 
 }
@@ -37,17 +38,24 @@ void MyModel::fromPrior()
 	mag0 = -100. + 200.*randomU();
 	for(size_t i=0; i<delta_mag.size(); i++)
 		delta_mag[i] = tan(M_PI*(randomU() - 0.5));
+
+	tau_qso = exp(log(1.) + log(1E6)*randomU());
+	beta_qso = exp(log(1E-3) + log(1E6)*randomU());
+	for(size_t i=0; i<n_qso.size(); i++)
+		n_qso[i] = randn();
 }
 
 double MyModel::perturb()
 {
-	int which = randInt(2);
+	double logH = 0.;
+	int which = randInt(4);
+
 	if(which == 0)
 	{
 		mag0 += 200.*pow(10., 1.5 - 6.*randomU())*randn();
 		mag0 = mod(mag0 + 100., 200.) - 100.;
 	}
-	else
+	else if(which == 1)
 	{
 		int which2 = randInt(delta_mag.size());
 		double u = 0.5 + atan(delta_mag[which2])/M_PI;
@@ -55,8 +63,44 @@ double MyModel::perturb()
 		u = mod(u, 1.);
 		delta_mag[which2] = tan(M_PI*(u - 0.5));
 	}
+	else if(which == 2)
+	{
+		tau_qso = log(tau_qso);
+		tau_qso += log(1E6)*pow(10., 1.5 - 6.*randomU())*randn();
+		tau_qso = mod(tau_qso - log(1.), log(1E6)) + log(1.);
+		tau_qso = exp(tau_qso);
+	}
+	else if(which == 3)
+	{
+		beta_qso = log(beta_qso);
+		beta_qso += log(1E6)*pow(10., 1.5 - 6.*randomU())*randn();
+		beta_qso = mod(beta_qso - log(1E-3), log(1E6)) + log(1E-3);
+		beta_qso = exp(beta_qso);
+	}
+	else if(which == 4)
+	{
+		double chance = pow(10., 0.5 - 4.*randomU());
+		double scale = pow(10., 1.5 - 6.*randomU());
+		bool full = randomU() <= 0.3;
+		for(size_t i=0; i<n_qso.size(); i++)
+		{
+			if(randomU() <= chance)
+			{
+				if(full)
[…]

### 2013-09-21 src/Brute_Force/MyModel.h

@@ -32,6 +32,10 @@ class MyModel:public DNest3::Model
 		// Magnitudes of others relative to image 0 (ignore delta_mag[0])
 		std::vector<double> delta_mag;
 
+		// QSO variability parameters
+		double tau_qso, beta_qso;
+		std::vector<double> n_qso;
+
 	public:
 		MyModel();
 

### 2013-09-21 src/Brute_Force/MyModel.cpp
QSO time series

@@ -29,6 +29,7 @@ using namespace DNest3;
 MyModel::MyModel()
 :delta_mag(Data::get_instance().get_numImages())
 ,n_qso(1000)
+,y_qso(1000)
 {
 
 }
@@ -43,6 +44,8 @@ void MyModel::fromPrior()
 	beta_qso = exp(log(1E-3) + log(1E6)*randomU());
 	for(size_t i=0; i<n_qso.size(); i++)
 		n_qso[i] = randn();
+
+	assemble();
 }
 
 double MyModel::perturb()
@@ -100,7 +103,17 @@ double MyModel::perturb()
 
 	}
 
+	assemble();
+
 	return logH;
+}
+
+void MyModel::assemble()
+{
+	double alpha = exp(-1./tau_qso);
+	y_qso[0] = mag0 + beta_qso/sqrt(1. - alpha*alpha)*n_qso[0];
+	for(size_t i=1; i<y_qso.size(); i++)
+		y_qso[i] = mag0 + alpha*(y_qso[i-1] - mag0) + beta_qso*n_qso[i];
 }
 
 double MyModel::logLikelihood() const
@@ -114,6 +127,8 @@ void MyModel::print(std::ostream& out) const
 	for(size_t i=0; i<delta_mag.size(); i++)
 		out<<delta_mag[i]<<' ';
 	out<<tau_qso<<' '<<beta_qso<<' ';
+	for(size_t i=0; i<y_qso.size(); i++)
+		out<<y_qso[i]<<' ';
 }
 
 string MyModel::description() const

### 2013-09-21 src/Brute_Force/MyModel.h

@@ -36,6 +36,17 @@ class MyModel:public DNest3::Model
 		double tau_qso, beta_qso;
 		std::vector<double> n_qso;
 
+
+
+
+
+
+		// DERIVED STUFF
+		std::vector<double> y_qso;
+
+		// Assemble function
+		void assemble();
+
 	public:
 		MyModel();
 

### 2013-09-23 src/Brute_Force/MyModel.cpp
Tau parameters

@@ -29,6 +29,7 @@ using namespace DNest3;
 MyModel::MyModel()
 :delta_mag(Data::get_instance().get_numImages())
 ,n_qso(1000)
+,tau(Data::get_instance().get_numImages())
 ,y_qso(1000)
 {
 
@@ -37,7 +38,9 @@ MyModel::MyModel()
 void MyModel::fromPrior()
 {
 	mag0 = -100. + 200.*randomU();
-	for(size_t i=0; i<delta_mag.size(); i++)
+
+	delta_mag[0] = 0.;
+	for(size_t i=1; i<delta_mag.size(); i++)
 		delta_mag[i] = tan(M_PI*(randomU() - 0.5));
 
 	tau_qso = exp(log(1.) + log(1E6)*randomU());
@@ -45,13 +48,17 @@ void MyModel::fromPrior()
 	for(size_t i=0; i<n_qso.size(); i++)
 		n_qso[i] = randn();
 
+	tau[0] = 0.;
+	for(size_t i=1; i<tau.size(); i++)
+		tau[i] = 10.*tan(M_PI*(randomU() - 0.5));
+
 	assemble();
 }
 
 double MyModel::perturb()
 {
 	double logH = 0.;
-	int which = randInt(4);
+	int which = randInt(6);
 
 	if(which == 0)
 	{
@@ -60,7 +67,7 @@ double MyModel::perturb()
 	}
 	else if(which == 1)
 	{
-		int which2 = randInt(delta_mag.size());
+		int which2 = 1 + randInt(delta_mag.size() - 1);
 		double u = 0.5 + atan(delta_mag[which2])/M_PI;
 		u += pow(10., 1.5 - 6.*randomU())*randn();
 		u = mod(u, 1.);
@@ -81,6 +88,14 @@ double MyModel::perturb()
 		beta_qso = exp(beta_qso);
 	}
 	else if(which == 4)
+	{
+		int which2 = 1 + randInt(tau.size() - 1);
+		double u = 0.5 + atan(tau[which2]/10.)/M_PI;
+		u += pow(10., 1.5 - 6.*randomU())*randn();
+		u = mod(u, 1.);
+		tau[which2] = 10.*tan(M_PI*(u - 0.5));
+	}
+	else if(which == 5)
 	{
 		double chance = pow(10., 0.5 - 4.*randomU());
 		double scale = pow(10., 1.5 - 6.*randomU());
[…]

### 2013-09-23 src/Brute_Force/MyModel.h

@@ -36,10 +36,8 @@ class MyModel:public DNest3::Model
 		double tau_qso, beta_qso;
 		std::vector<double> n_qso;
 
-
-
-
-
+		// Time delays
+		std::vector<double> tau;
 
 		// DERIVED STUFF
 		std::vector<double> y_qso;

### 2013-09-23 src/Brute_Force/MyModel.cpp
Likelihood!

@@ -28,11 +28,13 @@ using namespace DNest3;
 
 MyModel::MyModel()
 :delta_mag(Data::get_instance().get_numImages())
-,n_qso(1000)
+,n_qso(10000)
 ,tau(Data::get_instance().get_numImages())
-,y_qso(1000)
+,y_qso(10000)
 {
-
+	t_min = Data::get_instance().get_tMin() -   Data::get_instance().get_tRange();
+	t_max = Data::get_instance().get_tMin() + 2*Data::get_instance().get_tRange();
+	dt = (t_max - t_min)/(static_cast<int>(y_qso.size()) - 1);
 }
 
 void MyModel::fromPrior()
@@ -123,6 +125,14 @@ double MyModel::perturb()
 	return logH;
 }
 
+double MyModel::evaluate_y_qso(double t) const
+{
+	int i = static_cast<int>((t - t_min)/t_range);
+	if(i >= 0 && i < static_cast<int>(y_qso.size()))
+		return y_qso[i];
+	return 0.;
+}
+
 void MyModel::assemble()
 {
 	double alpha = exp(-1./tau_qso);
@@ -131,9 +141,23 @@ void MyModel::assemble()
 		y_qso[i] = mag0 + alpha*(y_qso[i-1] - mag0) + beta_qso*n_qso[i];
 }
 
+
+
 double MyModel::logLikelihood() const
 {
 	double logL = 0.;
+
+	const vector<double>& t = Data::get_instance().get_t();
+	const vector<double>& y = Data::get_instance().get_y();
+	const vector<double>& sig = Data::get_instance().get_sig();
+	const vector<int>& id = Data::get_instance().get_ID();
+
+	double mu;
+	for(size_t i=0; i<y.size(); i++)
+	{
+		mu = mag0 + delta_mag[id[i]] + evaluate_y_qso(t[i] - tau[id[i]]);
+		logL += -0.5*pow((y[i] - mu)/sig[i], 2);
+	}
 
 	return logL;
 }

### 2013-09-23 src/Brute_Force/MyModel.h

@@ -41,6 +41,10 @@ class MyModel:public DNest3::Model
 
 		// DERIVED STUFF
 		std::vector<double> y_qso;
+		double t_min, t_max, t_range, dt;
+
+		// Evaluate y_qso function
+		double evaluate_y_qso(double t) const;
 
 		// Assemble function
 		void assemble();

### 2013-09-23 src/Brute_Force/MyModel.cpp
Various changes, I'm not convinced

@@ -31,6 +31,7 @@ MyModel::MyModel()
 ,n_qso(10000)
 ,tau(Data::get_instance().get_numImages())
 ,y_qso(10000)
+,mu(Data::get_instance().get_numPoints())
 {
 	t_min = Data::get_instance().get_tMin() -   Data::get_instance().get_tRange();
 	t_max = Data::get_instance().get_tMin() + 2*Data::get_instance().get_tRange();
@@ -60,6 +61,7 @@ void MyModel::fromPrior()
 double MyModel::perturb()
 {
 	double logH = 0.;
+
 	int which = randInt(6);
 
 	if(which == 0)
@@ -127,9 +129,10 @@ double MyModel::perturb()
 
 double MyModel::evaluate_y_qso(double t) const
 {
-	int i = static_cast<int>((t - t_min)/t_range);
-	if(i >= 0 && i < static_cast<int>(y_qso.size()))
-		return y_qso[i];
+	int i = static_cast<int>((t - t_min)/dt);
+	double w = (t - (t_min + i*dt))/dt;
+	if(i >= 0 && i < static_cast<int>(y_qso.size()) - 1)
+		return (1. - w)*y_qso[i] + w*y_qso[i+1];
 	return 0.;
 }
 
@@ -139,6 +142,12 @@ void MyModel::assemble()
 	y_qso[0] = mag0 + beta_qso/sqrt(1. - alpha*alpha)*n_qso[0];
 	for(size_t i=1; i<y_qso.size(); i++)
 		y_qso[i] = mag0 + alpha*(y_qso[i-1] - mag0) + beta_qso*n_qso[i];
+
+	const vector<double>& t = Data::get_instance().get_t();
+	const vector<int>& id = Data::get_instance().get_ID();
+
+	for(size_t i=0; i<t.size(); i++)	
+		mu[i] = mag0 + delta_mag[id[i]] + evaluate_y_qso(t[i] - tau[id[i]]);
 }
 
 
@@ -147,17 +156,11 @@ double MyModel::logLikelihood() const
 {
 	double logL = 0.;
 
-	const vector<double>& t = Data::get_instance().get_t();
 	const vector<double>& y = Data::get_instance().get_y();
 	const vector<double>& sig = Data::get_instance().get_sig();
-	const vector<int>& id = Data::get_instance().get_ID();
 
-	double mu;
[…]

### 2013-09-23 src/Brute_Force/MyModel.h

@@ -41,7 +41,9 @@ class MyModel:public DNest3::Model
 
 		// DERIVED STUFF
 		std::vector<double> y_qso;
-		double t_min, t_max, t_range, dt;
+		double t_min, t_max, dt;
+
+		std::vector<double> mu;
 
 		// Evaluate y_qso function
 		double evaluate_y_qso(double t) const;
